\section{Prayers at the Foot of the Altar}
\fancyhead[RO,LE]{\textit{Foot of the Altar}}
\begin{secrubric}
    A Hymn may be sung during the \blackrubric{Prayers at the Foot of the Altar,} except in a Low Mass.
\end{secrubric}
\begin{secrubric}
    After the Rite of Aspersion (or immediately upon processing to the Foot of the Altar if the Rite not occur), the Priest---having vested---approaches the Altar, makes the due reverence, signs himself with the sign of the Cross from forehead to breast, and says in the clear voice:
\end{secrubric}
\elcol{℣. In the Name of the Father, {\ding{64}} and of the Son, and of the Holy Ghost.}{℣. In nómine Patris, {\ding{64}} et Fílii, et Spíritus Sancti.}
\elcol{℟. Amen.}{℟. Amen.}
\begin{rubric}
    Then, with hands joined before his breast, the Priest begins the Antiphon:
\end{rubric}
\elcol{℣. I will go unto the altar of God.}{℣. Introíbo ad altáre Dei.}
\elcol{℟. Even unto the God of my joy and gladness.}{℟. Ad Deum, qui l{\ae}tíficat juventútem meam.}
\begin{rubric}
    In Masses for the Dead, and in Masses of the Season from Passion Sunday to Holy Saturday exclusive, the following Psalm \blackrubric{Give sentence} is omitted together with \blackrubric{Glory be} and the repetition of the Antiphon. And the Priest then moves immediately to \blackrubric{Our help.}
\end{rubric}

\vspace{-0.45\baselineskip}
\par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}

	%MANUAL ADJUSTMENT:
\vspace{0.65\baselineskip}

\elcol{℣. Give sentence with me, O God, and defend my cause against the ungodly people: O deliver me from the deceitful and wicked man.

℟. For thou art the God of my strength; why hast thou put me from thee: and why go I so heavily, while the enemy oppresseth me?

℣. O send out thy light and thy truth, that they may lead me: and bring me unto thy holy hill, and to thy dwelling.

℟. And that I may go unto the altar of God, even the God of my joy and gladness: and upon the harp will I give thanks unto thee, O God, my God.

℣. Why art thou so heavy, O my soul: and why art thou so disquieted within me?

℟. O put thy trust in God: for I will yet give him thanks, which is the help of my countenance, and my God.

℣. Glory be to the Father, and to the Son: and to the Holy Ghost.

℟. As it was in the beginning, is now, and ever shall be: world without end. Amen.

℣. I will go unto the altar of God:

℟. Even unto the God of my joy and gladness.}
{℣. Júdica me, Deus, et discérne causam meam de gente non sancta: ab hómine iníquo et dolóso érue me.

℟. Quia tu es, Deus, fortitúdo mea: quare me repulísti, et quare tristis incédo, dum afflígit me inimícus?

℣. Emítte lucem tuam et veritátem tuam: ipsa me deduxérunt, et adduxérunt in montem sanctum tuum et in tabernácula tua.

℟. Et introíbo ad altáre Dei: ad Deum, qui l{\ae}tíficat juventútem meam.

℣. Confitébor tibi in cíthara, Deus, Deus meus: quare tristis es, ánima mea, et quare contúrbas me?

℟. Spera in Deo, quóniam adhuc confitébor illi: salutáre vultus mei, et Deus meus.

℣. Glória Patri, et Fílio, et Spirítui Sancto.

℟. Sicut erat in princípio, et nunc, et semper: et in sǽcula s{\ae}culórum. Amen.

℣. Introíbo ad altáre Dei.

℟. Ad Deum, qui l{\ae}tíficat juventútem meam.}
    \par\noindent
	\centerline{\rule{0.5\textwidth}{0.4pt}}\\
	\par\noindent
\elcol{℣. Our help is in the Name {\ding{64}} of the Lord:

℟. Who hath made heaven and earth.}{℣. Adjutórium nostrum {\ding{64}} in nómine Dómini.

℟. Qui fecit c{\ae}lum et terram.}
\begin{rubric}
    Then, with hands joined, bowing profoundly, the Priest says,
\end{rubric}
\elcol{\lett{I}{confess} to Almighty God, to Blessed Mary ever Virgin, to Blessed Michael the Archangel, to Blessed John Baptist, to the Holy Apostles Peter and Paul, to all the Saints, and to you, brethren: that I have sinned exceedingly in thought, word, and deed: \inrub{Strike breast thrice.} by my fault, my own fault, my own most grievous fault. Wherefore I beg Blessed Mary ever Virgin, Blessed Michael the Archangel, Blessed John Baptist, the Holy Apostles Peter and Paul, all the Saints, and you, brethren, to pray for me to the Lord our God.}{\lett{C}{onf\smash{í}teor} Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et vobis, fratres: quia peccávi nimis cogitatióne, verbo et ópere: \inrub{Strike breast thrice.} mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et vos, fratres, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℟. Almighty God have mercy upon thee, forgive thee thy sins, and bring thee to everlasting life.}{℟. Misereátur tui omnípotens Deus, et, dimíssis peccátis tuis, perdúcat te ad vitam {\ae}térnam.}
\elcol{℟. Amen.\margcolrub{1}{The Priest rises. The Ministers bow.}}{℟. Amen.\marglatincolrub{1}{The Priest rises. The Ministers bow.}}
\clearpage
\begin{rubric}
	The Ministers, bowing, say,
\end{rubric}
\elcol{\lett{I}{confess} to Almighty God, to Blessed Mary ever Virgin, to Blessed Michael the Archangel, to Blessed John Baptist, to the Holy Apostles Peter and Paul, to all the Saints, and to thee, Father: that I have sinned exceedingly in thought, word, and deed: \inrub{Strike breast thrice.} by my fault, my own fault, my own most grievous fault. Wherefore I beg Blessed Mary ever Virgin, Blessed Michael the Archangel, Blessed John Baptist, the Holy Apostles Peter and Paul, all the Saints, and thee, Father, to pray for me to the Lord our God.}{\lett{C}{onf\smash{í}teor} Deo omnipoténti, beát{\ae} Marí{\ae} semper Vírgini, beáto Michaéli Archángelo, beáto Joánni Baptíst{\ae}, sanctis Apóstolis Petro et Paulo, ómnibus Sanctis, et tibi, pater: quia peccávi nimis cogitatióne, verbo et ópere: \inrub{Strike breast thrice.} mea culpa, mea culpa, mea máxima culpa. Ideo precor beátam Maríam semper Vírginem, beátum Michaélem Archángelum, beátum Joánnem Baptístam, sanctos Apóstolos Petrum et Paulum, omnes Sanctos, et te, pater, oráre pro me ad Dóminum, Deum nostrum.}
\elcol{℣. Almighty God have mercy upon you, forgive you your sins, and bring you to everlasting life.}{℣. Misereátur vestri omnípotens Deus, et, dimíssis peccátis vestris, perdúcat vos ad vitam {\ae}térnam.}
\elcol{℟. Amen.}{℟. Amen.}
\elcol{℣. The Almighty and Merciful Lord grant unto us pardon, {\ding{64}} absolution, and remission of our sins.

℟. Amen.\margcolrub{2}{Ministers rise.}

℣. Wilt thou not turn again and quicken us, O God?

℟. That thy people may rejoice in thee.

℣. O Lord, show thy mercy upon us.

℟. And grant us thy salvation.

℣. O Lord, hear my prayer.

℟. And let my cry come unto thee.

℣. The Lord be with you.

℟. And with thy spirit.}
{℣. Indulgéntiam, {\ding{64}} absolutiónem et remissiónem peccatórum nostrórum tríbuat nobis omnípotens et miséricors Dóminus.

℟. Amen.\marglatincolrub{2}{Ministers rise.}

℣. Deus, tu convérsus vivificábis nos?

℟. Et plebs tua l{\ae}tábitur in te.

℣. Osténde nobis, Dómine, misericórdiam tuam.

℟. Et salutáre tuum da nobis.

℣. Dómine, exáudi oratiónem meam.

℟. Et clamor meus ad te véniat.

℣. Dóminus vobíscum.

℟. Et cum spíritu tuo.}

\begin{rubric}
    The Priest, extending and joining his hands, says in a loud voice:
\end{rubric}

\elcol{℣. Let us pray.}{℣. Orémus.}