\thispagestyle{copyrightpage}
\noindent
The Saint Botolph's Missal ($\copyright$ 2026 Apologia Anglicana, LLC) is licensed under CC BY-SA 4.0. To view a copy of this license, visit \url{https://creativecommons.org/licenses/by-sa/4.0/}\\
\par\noindent
Edited by J. P. `Augustine' Watson (Augustine@ApologiaAnglicana.org), with the irreplaceable assistance of Ryan May.\\

ISBN: 978-1-969461-07-1 (Personal Size Hardcover)

ISBN: 978-1-969461-08-8 (Altar Size Hardcover)\\

Library of Congress Control Number: 2025919720\\

\par\noindent
We are thankful to have such great giants upon whose shoulders we stand. These works are used in this Book according to fair use, where their content is not already in the public domain. What was used from the book, or in what way it was consulted, is described under each citation.\\

\textit{The American Missal: Revised \& Expanded}. Glendale: Lancelot Andrewes Press, 2010.\par\noindent
\textit{The propers for Our Lady of Walsingham, St. Raphael of Brooklyn, St. Tikhon of Moscow, and the Old Testament Patriarchs \& Prophets.}\\

\textit{The English Missal for the Laity}. London: W. Knott \& Sons Limited, 1933.\par\noindent
\textit{Mass Rubrics. Parts of the Mass not found in the Prayer Book Tradition. Propers not found in the Prayer Book Tradition nor used from other books.}\\
\par\noindent
The following artwork has been used, according to the terms of its license.\\

Fr James Bradley (\url{https://www.flickr.com/photos/jamesbradley/}). \textit{IMG\_2936 Saint Botolph} (\url{https://www.flickr.com/photos/jamesbradley/28884698391/}). Creative Commons BY 2.0 (\url{https://creativecommons.org/licenses/by/2.0/deed.en}).
\par\noindent
\textit{Used in the cover of the Personal Edition.}\\

Madboy74 (\url{https://commons.wikimedia.org/wiki/User:Madboy74}). \textit{Coa Illustration Cross Canterbury.svg} (\url{https://commons.wikimedia.org/wiki/File:Coa_Illustration_Cross_Canterbury.svg}). Creative Commons BY-SA 4.0 (\url{https://creativecommons.org/licenses/by-sa/4.0/deed.en}).\par\noindent
\textit{Used in the cover of the Altar Edition.}

\clearpage