\fancyhead[C]{{\LARGE Kalendar of the Church Year}}\label{kalendar}
\phantomsection
\addcontentsline{toc}{chapter}{Kalendar of the Church Year}

\begin{longtblr}{rowsep = 0pt,
colspec = {l|l},
}
	\SetCell[c=2]{l} \textsc{January}\\
	1&{\color{RubricRed}Circumcision of Our Lord,} \textit{\nth{2} Double}\\
	&(Octave Day of the Nativity)\\
	2&{\color{RubricRed}Most Holy Name of Jesus,} \textit{\nth{2} Double}\\
	&Octave Day of St. Stephen, Protomartyr, \textit{Simple}\\
	3&Octave Day of St. John, Apostle \& Evangelist, \textit{Simple}\\
	4&Octave Day of the Holy Innocents, \textit{Simple}\\
	&St. Titus, Bishop \& Confessor, \textit{Memorial}\\
	5&Vigil of the Epiphany, \textit{\nth{2} Vigil}\\
	&Pope St. Telesphorus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	6&{\color{RubricRed}Epiphany of Our Lord,} \textit{\nth{1} Double,} \nth{2} Octave\\
	7&Day II within the Octave of the Epiphany, \textit{Semidouble}\\
	8&Day III within the Octave of the Epiphany, \textit{Semidouble}\\
	9&Day IV within the Octave of the Epiphany, \textit{Semidouble}\\
	10&Day V within the Octave of the Epiphany, \textit{Semidouble}\\
	&St. Paul the First Hermit, Confessor \textit{Memorial}\\
	11&Day VI within the Octave of the Epiphany, \textit{Semidouble}\\
	&Pope St. Hyginus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	12&Day VII within the Octave of the Epiphany, \textit{Semidouble}\\
	&St. Benedict Biscop, Abbot, \textit{Memorial}\\
	%Emphasis is here laid on the Baptism to be consistent with other Octave Days
	13&{\color{RubricRed}Octave Day of the Epiphany,} \textit{Greater Double}\\
	&(Baptism of Our Lord)\\
	14&St. Hilary of Poitiers, Bishop, Confessor, \& Doctor, \textit{Double}\\
	&St. Felix of Nola, Martyr, \textit{Memorial}\\
	15&St. Maurus, Abbot, \textit{Greater Double}\\
	16&Pope St. Marcellus I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	17&St. Anthony, Abbot, \textit{Double}\\
	18&Chair of St. Peter in Rome, \textit{Greater Double}\\
	&Commemoration of St. Paul, \textit{Commemoration}\\
	&St. Prisca of Rome, Virgin Martyress, \textit{Memorial}\\
	19&Sts. Marius, Martha, Audifax, \& Abachum, Martyrs, \textit{Memorial}\\
	&St. Mark of Ephesus, Bishop \& Confessor, \textit{Memorial}\\
	20&Sts. Pope Fabian of Rome (Bishop) \& Sebastian, Martyrs, \textit{Double}\\
	21&St. Agnes of Rome, Virgin Martyress, \textit{Greater Double}\\
	22&Sts. Vincent \& Anastasius, Martyrs, \textit{Memorial}\\
	23&St. Emerentiana, Virgin Martyress, \textit{Memorial}\\
	24&St. Timothy, Bishop \& Martyr, \textit{Double}\\
	25&{\color{RubricRed}Conversion of St. Paul the Apostle,} \textit{Greater Double}\\
	&{Commemoration of St. Peter,} \textit{Commemoration}\\
	26&St. Polycarp of Smyrna, Bishop \& Martyr, \textit{Memorial}\\
	27&St. John Chrysostom, Bishop, Confessor, \& Doctor, \textit{Double}\\
	28&Pope St. Cyril of Alexandria, Bishop, Confessor, \& Doctor, \textit{Double}\\
	&St. Agnes of Rome, Virgin Martyress, \textit{Memorial}\\
	29&\\
	30&St. Martina, Virgin Martyress, \textit{Memorial}\\
	31&(Earliest Date of Septuagesima)\\
	%
	\SetCell[c=2]{l} \textsc{February}\\
	1&St. Ignatius of Antioch, Bishop \& Martyr, \textit{Double}\\
	&St. Bridget of Ireland, Virgin, \textit{Memorial}\\
	2&{\color{RubricRed}Purification of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	3&St. Blaise of Sebaste, Bishop \& Martyr, \textit{Memorial}\\
	4&St. Joseph of Aleppo, Martyr, \textit{Memorial}\\
	&New Martyrs of Russia, \textit{Memorial}\\
	5&St. Agatha of Sicily, Virgin Martyress, \textit{Greater Double}\\
	6&St. Dorothea of Caesarea, Virgin Martyress, \textit{Memorial}\\
	&St. Photius I of Constantinople, Bishop \& Confessor, \textit{Memorial}\\
	7&St. Romuald, Abbot, \textit{Double}\\
	8&\\
	9&St. Apollonia of Alexandria, Virgin Martyress, \textit{Memorial}\\
	%Some make this a second double:
	10&St. Scholastica, Virgin, \textit{Greater Double}\\
	11&Pope St. Gregory II, Bishop \& Confessor, \textit{Memorial}\\
	12&\\
	13&\\
	14&St. Valentine, Martyr, \textit{Memorial}\\
	15&Sts. Faustinus \& Jovita, Martyrs, \textit{Memorial}\\
	16&\\
	17&(Earliest Date of Lent)\\
	18&St. Simeon of Jerusalem, Bishop \& Martyr, \textit{Memorial}\\
	19&\\
	20&\\
	21&\\
	22&{\color{RubricRed}Chair of St. Peter at Antioch,} \textit{\nth{2} Double}\\
	&Commemoration of St. Paul, \textit{Commemoration}\\
	23&Vigil of St. Matthias, \textit{Vigil}\\
	24&{\color{RubricRed}St. Matthias, Apostle,} \textit{\nth{2} Double}\\
	25&St. Walburga of Heidenheim, Virgin, \textit{Memorial}\\
	26&Pope St. Alexander of Alexandria, Bishop \& Confessor, \textit{Memorial}\\
	27&St. Raphael of Brooklyn, Bishop \& Confessor, \textit{Greater Double}\\
	28&\\
	%
	\SetCell[c=2]{l} \textsc{March}\\
	1&St. David of Wales, Bishop \& Confessor, \textit{Memorial}\\
	2&St. Chad, Bishop \& Confessor, \textit{Memorial}\\
	3&\\
	4&Pope St. Lucius I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	5&\\
	6&Sts. Perpetua \& Felicitas, Martyresses, \textit{Double}\\
	&(Latest Date of Septuagesima)\\
	7&\\
	8&\\
	9&St. Gregory of Nyssa, Bishop, Confessor, \& Doctor, \textit{Memorial}\\
	10&Forty Holy Martyrs, \textit{Memorial}\\
	11&\\
	12&{\color{RubricRed}Pope St. Gregory of Rome, Bishop, Confessor, \& Doctor,} \textit{\nth{2} Double}\\
	13&\\
	14&\\
	15&\\
	16&\\
	17&St. Patrick, Bishop \& Confessor, \textit{Double}\\
	&St. Joseph of Arimathea, Confessor, \textit{Memorial}\\
	18&St. Cyril of Jerusalem, Bishop, Confessor, \& Doctor, \textit{Double}\\
	&King St. Edward, Martyr, \textit{Memorial}\\
	19&{\color{RubricRed}St. Joseph, Spouse of the Blessed Virgin Mary, Confessor,} \textit{\nth{2} Double}\\
	20&St. Cuthbert, Bishop \& Confessor, \textit{Double}\\
	21&{\color{RubricRed}St. Benedict, Abbot,} \textit{\nth{2} Double}\\
	22&\\
	23&(Latest Date of Lent)\\
	24&St. Gabriel the Archangel, \textit{Greater Double}\\
	25&{\color{RubricRed}Annunciation of the Blessed Virgin Mary,} \textit{\nth{1} Double}\\
	26&\\
	27&St. John Damascene, Confessor \& Doctor, \textit{Double}\\
	28&\\
	29&\\
	30&St. John Climacus, Abbot, \textit{Memorial}\\
	31&St. Innocent of Alaska, Bishop \& Confessor, \textit{Memorial}\\
	%
	\SetCell[c=2]{l} \textsc{April}\\
	1&\\
	2&\\
	3&\\
	4&St. Isidore of Seville, Bishop, Confessor, \& Doctor, \textit{Double}\\
	&(Earliest Date of Easter)\\
	5&\\
	6&\\
	7&{\color{RubricRed}St. Tikhon of Moscow, Bishop \& Confessor,} \textit{\nth{1} Double,} Common Octave\\
	8&\\
	9&\\
	10&\\
	11&Pope St. Leo the Great, Bishop, Confessor, \& Doctor, \textit{Double}\\
	12&\\
	13&St. Hermengild, Martyr, \textit{Memorial}\\
	14&St. Justin, Martyr, \textit{Memorial}\\
	&Sts. Tiburtius, Valerian, \& Maximus, Martyrs, \textit{Memorial}\\
	15&\\
	16&\\
	17&Pope St. Anicetus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	18&\\
	19&\\
	20&\\
	21&\\
	22&Pope Sts. Soter \& Caius of Rome, Bishops \& Martyrs, \textit{Memorial}\\
	23&{\color{RubricRed}St. George, Martyr,} \textit{\nth{1} Double,} Common Octave\\
	24&Day II within the Octave of St. George, \textit{Semidouble}\\
	25&{\color{RubricRed}St. Mark, Evangelist,} \textit{\nth{2} Double}\\
	&Day III within the Octave of St. George, \textit{Semidouble}\\
	26&Day IV within the Octave of St. George, \textit{Semidouble}\\
	&Sts. Cletus \& Marcellinus, Martyrs, \textit{Memorial}\\
	27&Day V within the Octave of St. George, \textit{Semidouble}\\
	28&Day VI within the Octave of St. George, \textit{Semidouble}\\
	&St. Vitalis, Martyr, \textit{Memorial}\\
	29&Day VII within the Octave of St. George, \textit{Semidouble}\\
	30&Octave Day of St. George, \textit{Greater Double}\\
	%
	\SetCell[c=2]{l} \textsc{May}\\
	1&{\color{RubricRed}Sts. Philip \& James, Apostles,} \textit{\nth{2} Double}\\
	2&Pope St. Athanasius of Alexandria, Bishop, Confessor, \& Doctor, \textit{Double}\\
	3&{\color{RubricRed}Invention of the Holy Cross,} \textit{\nth{2} Double}\\
	&Sts. Pope Alexander of Rome (Bishop), Eventius, Theodulus (Martyrs),\\
	&\hspace{1em} \& Juvenalis (Bishop \& Confessor), \textit{Memorial}\\
	4&St. Monica, Widow, \textit{Memorial}\\
	5&\\
	6&St. John before the Latin Gate, \textit{Greater Double}\\
	7&St. Alexis Toth, Confessor, \textit{Memorial}\\
	8&Apparition of St. Michael the Archangel, \textit{Greater Double}\\
	&Pope St. Boniface IV of Rome, Bishop \& Confessor, \textit{Memorial}\\
	&(Latest Date of Easter)\\
	9&St. Gregory Nazianzen, Bishop, Confessor, \& Doctor, \textit{Double}\\
	10&Sts. Gordian \& Epimachus, Martyrs, \textit{Memorial}\\
	11&\\
	12&Sts. Nereus, Achilles, Domitilla, \& Pancras, Martyrs, \textit{Memorial}\\
	13&\\
	14&St. Boniface of Tarsus, Martyr, \textit{Memorial}\\
	15&\\
	16&\\
	17&\\
	18&St. Venantius, Martyr, \textit{Double}\\
	19&St. Pudentiana, Virgin, \textit{Memorial}\\
	20&\\
	21&\\
	22&St. Romanus, Abbot, \textit{Memorial}\\
	23&\\
	24&St. Vincent of Lerins, Confessor, \textit{Memorial}\\
	25&Pope St. Urban of Rome, Bishop \& Martyr, \textit{Memorial}\\
	26&St. Augustine of Canterbury, Bishop \& Confessor, \textit{Double}\\
	&Pope St. Eleutherius of Rome, Bishop \& Martyr, \textit{Memorial}\\
	27&St. Bede the Venerable, Confessor \& Doctor, \textit{Double}\\
	&Pope St. John I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	28&\\
	29&\\
	30&Pope St. Felix I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	31&St. Petronilla, Virgin, \textit{Memorial}\\
	%
	\SetCell[c=2]{l} \textsc{June}\\
	1&\\
	2&St. Marcellinus, Peter, \& Erasmus, Martyrs, \textit{Memorial}\\
	3&\\
	4&\\
	5&St. Boniface, Bishop \& Martyr, \textit{Double}\\
	6&\\
	7&\\
	8&\\
	9&St. Columba of Iona, Abbot, \textit{Double}\\
	&Sts. Primus \& Felician, Martyrs, \textit{Memorial}\\
	10&Queen St. Margaret of Scotland, Matron, \textit{Memorial}\\
	11&{\color{RubricRed}St. Barnabas, Apostle,} \textit{Greater Double}\\
	12&Sts. Basilides, Cyrinus, Nabor, \& Nazarius, Martyrs, \textit{Memorial}\\
	13&\\
	14&St. Basil the Great, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
	15&Sts. Vitus, Modestus, \& Crescentia, Martyrs, \textit{Memorial}\\
	16&\\
	17&\\
	18&St. Ephrem the Syrian, Confessor \& Doctor, \textit{Double}\\
	&Sts. Marcus \& Marcellianus, Martyrs, \textit{Memorial}\\
	19&Sts. Gervase \& Protase, Martyrs, \textit{Memorial}\\
	20&Pope St. Silverius of Rome, Bishop \& Martyr, \textit{Memorial}\\
	21&\\
	22&St. Alban, Protomartyr of England, \textit{Double}\\
	&St. Paulinus, Bishop \& Confessor, \textit{Memorial}\\
	23&Vigil of St. John Baptist, \textit{Vigil}\\
	&Queen St. Etheldreda, Virgin, \textit{Memorial}\\
	24&{\color{RubricRed}St. John Baptist,} \textit{\nth{1} Double,} Common Octave\\
	&(Nativity of St. John Baptist)\\
	25&Day II within the Octave of St. John Baptist, \textit{Semidouble}\\
	26&Sts. John \& Paul, Martyrs, \textit{Double}\\
	&Day III within the Octave of St. John Baptist, \textit{Semidouble}\\
	27& Day IV within the Octave of St. John Baptist, \textit{Semidouble}\\
	28& Day V within the Octave of St. John Baptist, \textit{Semidouble}\\
	&Vigil of Sts. Peter \& Paul, \textit{Vigil}\\
	&Pope St. Leo II of Rome, Bishop \& Confessor, \textit{Memorial}\\
	29&{\color{RubricRed}St. Peter, Apostle,} \textit{\nth{1} Double,} Common Octave\\
	&(Sts. Peter \& Paul, Apostles)\\
	&{Day VI within the Octave of St. John Baptist,} \textit{Semidouble}\\
	30&Commemoration of St. Paul, Apostle, \textit{Greater Double}\\
	&Commemoration of St. Peter, Apostle, \textit{Commemoration}\\
	&Day VII within the Octave of St. John Baptist, \textit{Semidouble}\\
	%
	\SetCell[c=2]{l} \textsc{July}\\
	1&{\color{RubricRed}Most Precious Blood of Our Lord Jesus Christ,} \textit{\nth{2} Double}\\
	&{Octave Day of St. John Baptist,} \textit{Greater Double}\\
	&Day III within the Octave of Sts. Peter \& Paul, \textit{Semidouble}\\
	2&{\color{RubricRed}Visitation of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	&St. John of San Francisco, Bishop \& Confessor, \textit{Memorial}\\
	&Sts. Processus \& Martinian, Martyrs, \textit{Memorial}\\
	&Day IV within the Octave of Sts. Peter \& Paul, \textit{Semidouble}\\
	3&St. Iren{\ae}us of Lyon, Bishop \& Martyr, \textit{Double}\\
	&Day V within the Octave of Sts. Peter \& Paul, \textit{Semidouble}\\
	4&Day VI within the Octave of Sts. Peter \& Paul, \textit{Semidouble}\\
	5&Day VII within the Octave of Sts. Peter \& Paul, \textit{Semidouble}\\
	6&Octave Day of Sts. Peter \& Paul, \textit{Greater Double}\\
	7&Sts. Cyril \& Methodius, Bishops \& Confessors, \textit{Double}\\
	8&\\
	9&\\
	10&Seven Holy Brothers, with Sts. Rufina \& Secunda, Martyrs, \textit{Memorial}\\
	&Sts. Joseph of Damascus \& Companions, Martyrs, \textit{Memorial}\\
	11&Pope St. Pius I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	&[Solemnity of St. Benedict (Monastics \& Oblates only)], \textit{Greater Double}\\
	12&Sts. Nabor \& Felix, Martyrs, \textit{Memorial}\\
	13&Pope St. Anacletus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	14&\\
	15&King St. Vladimir of Kiev, Confessor, \textit{Double}\\
	16&Sts. Nicholas \& Habib Khasha, Martyrs, \textit{Memorial}\\
	&[Our Lady of Einsiedeln (Monastics \& Oblates only)], \textit{Greater Double}\\
	17&St. Alexius, Confessor, \textit{Memorial}\\
	18&Translation of St. Raphael of Brooklyn, \textit{Double}\\
	&Sts. Symphorosa \& Her Seven Sons, Martyresses, \textit{Memorial}\\
	&St. Sergius of Radonezh, Abbot, \textit{Memorial}\\
	19&St. Seraphim of Sarov, Confessor, \textit{Memorial}\\
	20&St. Elias the Prophet, Confessor, \textit{Double}\\
	&St. Margaret of Antioch, Virgin Martyress, \textit{Memorial}\\
	21&St. Praxedes, Virgin, \textit{Memorial}\\
	22&{\color{RubricRed}St. Mary Magdalene, Penitent,} \textit{Greater Double}\\
	23&St. John Cassian, Abbot, \textit{Memorial}\\
	&St. Apollinaris of Ravenna, Bishop \& Martyr, \textit{Memorial}\\
	&St. Liborius, Bishop \& Confessor, \textit{Memorial}\\
	24&Vigil of St. James, \textit{Vigil}\\
	&St. Christina, Virgin Martyress, \textit{Memorial}\\
	25&{\color{RubricRed}St. James, Apostle,} \textit{\nth{2} Double}\\
	&St. Christopher, Martyr, \textit{Memorial}\\
	26&{\color{RubricRed}St. Anne, Mother of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	27&St. Pantaleon, Martyr, \textit{Memorial}\\
	28&Sts. Nazarius, Celsus, Martyrs; Pope St. Victor of Rome, Bishop \& Martyr;\\
	&\hspace{1em} \& Pope St. Innocent of Rome, Bishop \& Confessor, \textit{Memorial}\\
	29&St. Martha of Bethany, Virgin, \textit{Double}\\
	&Sts. Pope Felix II of Rome (Bishop), Simplicius, Faustinus,\\
	&\hspace{1em} \& Beatrice, Martyrs, \textit{Memorial}\\
	30&Sts. Abdon \& Sennen, Martyrs, \textit{Memorial}\\
	31&\\
	%
	\SetCell[c=2]{l} \textsc{August}\\
	1&Chains of St. Peter, Apostle, \textit{Greater Double}\\
	&Commemoration of St. Paul, Apostle, \textit{Commemoration}\\
	&Holy Maccabees, Martyrs, \textit{Memorial}\\
	2&Pope St. Stephen of Rome, Bishop \& Martyr, \textit{Memorial}\\
	3&Invention of St. Stephen, Protomartyr, \textit{Double}\\
	4&\\
	5&Dedication of Our Lady of the Snows, \textit{Greater Double}\\
	&King St. Oswald, Martyr, \textit{Memorial}\\
	6&{\color{RubricRed}Transfiguration of Our Lord Jesus Christ,} \textit{\nth{2} Double}\\
	&Sts. Pope Sixtus II of Rome (Bishop), Felicissimus,\\
	&\hspace{1em} \& Agapitus, Martyrs, \textit{Memorial}\\
	7&{\color{RubricRed}Most Holy Name of Jesus,} \textit{\nth{2} Double}\\
	&St. Donatus, Bishop \& Martyr, \textit{Memorial}\\
	8&Sts. Cyriacus, Largus, \& Smaragdus, Martyrs, \textit{Memorial}\\
	9&Vigil of St. Lawrence, \textit{Vigil}\\
	&St. Romanus, Martyr, \textit{Memorial}\\
	10&{\color{RubricRed}St. Lawrence, Martyr,} \textit{\nth{2} Double,} Simple Octave\\
	11&Sts. Tiburtius \& Susanna, Martyrs, \textit{Memorial}\\
	12&\\
	13&St. Maximus of Constantinople, Confessor, \textit{Double}\\
	&Sts. Hippolytus \& Cassian, Martyrs, \textit{Memorial}\\
	14&Vigil of the Assumption, \textit{Vigil}\\
	&St. Eusebius, Confessor, \textit{Memorial}\\
	15&{\color{RubricRed}Assumption of the Blessed Virgin Mary,} \textit{\nth{1} Double,} Common Octave\\
	16&{\color{RubricRed}St. Joachim, Father of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	17&Day III within the Octave of the Assumption, \textit{Semidouble}\\
	&Octave Day of St. Lawrence, \textit{Simple}\\
	18&Day IV within the Octave of the Assumption, \textit{Semidouble}\\
	&St. Helen, Empress, \textit{Memorial}\\
	&St. Agapitus, Martyr, \textit{Memorial}\\
	19&Day V within the Octave of the Assumption, \textit{Semidouble}\\
	20&Day VI within the Octave of the Assumption, \textit{Semidouble}\\
	21&Day VII within the Octave of the Assumption, \textit{Semidouble}\\
	22&Octave Day of the Assumption, \textit{Greater Double}\\
	&Sts. Timothy (Bishop), Hippolytus, \& Symphorian, Martyrs, \textit{Memorial}\\
	23&Vigil of St. Bartholomew, \textit{Vigil}\\
	24&{\color{RubricRed}St. Bartholomew, Apostle,} \textit{\nth{2} Double}\\
	25&\\
	26&Pope St. Zephyrinus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	27&\\
	28&St. Augustine of Hippo, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
	&St. Hermes, Martyr, \textit{Memorial}\\
	29&Decollation of St. John Baptist, \textit{Greater Double}\\
	&St. Sabina, Martyress, \textit{Memorial}\\
	30&Sts. Felix \& Adauctus, Martyrs, \textit{Memorial}\\
	31&St. Aidan of Lindisfarne, Bishop \& Confessor, \textit{Memorial}\\
	%
	\SetCell[c=2]{l} \textsc{September}\\
	1&St. Giles, Abbot, \textit{Memorial}\\
	&Twelve Holy Brothers, Martyrs, \textit{Memorial}\\
	2&King St. Stephen of Hungary, Confessor, \textit{Memorial}\\
	3&\\
	4&St. Gorazde of Prague, Bishop \& Martyr, \textit{Double}\\
	5&\\
	6&\\
	7&\\
	8&{\color{RubricRed}Nativity of the Blessed Virgin Mary,} \textit{\nth{2} Double,} Simple Octave\\
	&St. Hadrian, Martyr, \textit{Memorial}\\
	9&St. Gorgonius, Martyr, \textit{Memorial}\\
	10&\\
	11&Sts. Protus \& Hyacinth, Martyrs, \textit{Memorial}\\
	12&Most Holy Name of Mary, \textit{Greater Double}\\
	13&\\
	14&Exaltation of the Holy Cross, \textit{Greater Double}\\
	15&{\color{RubricRed}Seven Sorrows of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	&St. Nicomedes, Martyr, \textit{Memorial}\\
	16&Sts. Pope Cornelius of Rome \& Cyprian, Bishops and Martyrs, \textit{Double}\\
	&Sts. Euphemia, Lucy, \& Geminian, Martyrs, \textit{Memorial}\\
	17&\\
	18&\\
	19&Sts. Januarius \& Companions, Martyrs, \textit{Double}\\
	&St. Theodore of Canterbury, Bishop \& Confessor, \textit{Memorial}\\
	20&Sts. Eustace \& Companions, Martyrs, \textit{Double}\\
	&Vigil of St. Matthew, \textit{Vigil}\\
	21&{\color{RubricRed}St. Matthew, Apostle \& Evangelist,} \textit{\nth{2} Double}\\
	22&Sts. Maurice \& Companions, Martyrs, \textit{Memorial}\\
	23&Pope St. Linus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	&St. Thecla, Virgin Martyress, \textit{Memorial}\\
	24&\\
	25&\\
	26&Sts. Cyprian \& Justina, Martyrs, \textit{Memorial}\\
	27&Sts. Cosmas \& Damian, Martyrs, \textit{Memorial}\\
	28&Duke St. Wenceslaus, Martyr, \textit{Memorial}\\
	29&{\color{RubricRed}St. Michael \& All Angels,} \textit{\nth{1} Double}\\
	&(Dedication of St. Michael the Archangel)\\
	30&St. Jerome, Confessor \& Doctor, \textit{Greater Double}\\
	&St. Gregory of Armenia, Bishop \& Confessor, \textit{Memorial}\\
	%
	\SetCell[c=2]{l} \textsc{October}\\
	1&St. Remigius, Bishop \& Confessor, \textit{Memorial}\\
	2&Holy Guardian Angels, \textit{Greater Double}\\
	3&\\
	4&\\
	5&Sts. Placidus \& Companions, Martyrs, \textit{Double}\\
	6&\\
	7&Holy Rosary of the Blessed Virgin Mary, \textit{Greater Double}\\
	&Pope St. Mark of Rome, Bishop \& Confessor, \textit{Memorial}\\
	&Sts. Sergius, Bacchus, Marcellus, \& Apuleius, Martyrs, \textit{Memorial}\\
	8&\\
	9&Sts. Denys, Rusticus, \& Eleutherius, Martyrs, \textit{Memorial}\\
	10&\\
	11&{\color{RubricRed}Motherhood of the Blessed Virgin Mary,} \textit{\nth{2} Double}\\
	12&St. Wilfrid, Bishop \& Confessor, \textit{Memorial}\\
	13&King St. Edward, Confessor, \textit{Memorial}\\
	14&Pope St. Callistus of Rome, Bishop \& Martyr, \textit{Double}\\
	15&{\color{RubricRed}Our Lady of Walsingham,} \textit{Greater Double}\\
	16&\\
	17&\\
	18&{\color{RubricRed}St. Luke, Evangelist,} \textit{\nth{2} Double}\\
	19&St. Frideswide, Virgin, \textit{Memorial}\\
	20&\\
	21&St. Hilarion, Abbot, \textit{Memorial}\\
	&Sts. Ursula \& Companions, Virgin Martyresses, \textit{Memorial}\\
	22&\\
	23&\\
	24&St. Raphael the Archangel, \textit{Greater Double}\\
	25&Sts. Chrysanthus \& Daria, Martyrs, \textit{Memorial}\\
	26&Pope St. Evaristus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	27&Vigil of Sts. Simon \& Jude, \textit{Vigil}\\
	28&{\color{RubricRed}Sts. Simon \& Jude, Apostles,} \textit{\nth{2} Double}\\
	29&\\
	30&\\
	31&Vigil of All Hallows, \textit{Vigil}\\
	%
	\SetCell[c=2]{l} \textsc{November}\\
	1&{\color{RubricRed}All Hallows,} \textit{\nth{1} Double,} Common Octave\\
	2&{\color{RubricRed}All Souls,} \textit{Double}\\
	&{Day II within the Octave of All Hallows,} \textit{Semidouble}\\
	3&Day III within the Octave of All Hallows, \textit{Semidouble}\\
	&St. Winifred, Virgin Martyress, \textit{Memorial}\\
	4&Day IV within the Octave of All Hallows, \textit{Semidouble}\\
	&Sts. Vitalis \& Agricola, Martyrs, \textit{Memorial}\\
	5&Day V within the Octave of All Hallows, \textit{Semidouble}\\
	&St. Elizabeth, Mother of St. John Baptist, \textit{Memorial}\\
	6&Day VI within the Octave of All Hallows, \textit{Semidouble}\\
	7&Day VII within the Octave of All Hallows, \textit{Semidouble}\\
	&St. Willibrord, Bishop \& Confessor, \textit{Memorial}\\
	8&Octave Day of All Hallows, \textit{Greater Double}\\
	&Four Crowned Martyrs, \textit{Memorial}\\
	9&Dedication of the Basilica of St. Saviour, \textit{Greater Double}\\
	&St. Theodore Tyro, Martyr, \textit{Memorial}\\
	10&Sts. Tryphon, Respicius, \& Nympha, Martyrs, \textit{Memorial}\\
	11&St. Martin of Tours, Bishop \& Confessor, \textit{Greater Double}\\
	&St. Mennas, Martyr, \textit{Memorial}\\
	12&Pope St. Martin I of Rome, Bishop \& Martyr, \textit{Memorial}\\
	13&St. Britius of Tours, Bishop \& Confessor, \textit{Memorial}\\
	14&St. Gregory Palamas, Bishop \& Confessor, \textit{Memorial}\\
	15&\\
	16&\\
	%MANUAL ADJUSTMENT:
	&\\
	17&St. Gregory the Wonder-worker, Bishop \& Confessor, \textit{Memorial}\\
	&St. Hilda of Whitby, Virgin, \textit{Memorial}\\
	18&Dedication of the Basilica of the Holy Apostles Peter \& Paul, \textit{Greater Double}\\
	19&Pope St. Pontianus of Rome, Bishop \& Martyr, \textit{Memorial}\\
	20&King St. Edmund, Martyr, \textit{Double}\\
	21&Presentation of the Blessed Virgin Mary, \textit{Greater Double}\\
	&St. Gelasius, Bishop \& Confessor, \textit{Memorial}\\
	&St. Columbanus, Abbot, \textit{Memorial}\\
	22&St. Cecilia, Virgin Martyress, \textit{Double}\\
	23&Pope St. Clement I of Rome, Bishop \& Martyr, \textit{Double}\\
	&St. Felicitas, Martyress, \textit{Memorial}\\
	24&St. Chrysogonus, Martyr \textit{Memorial}\\
	25&St. Catherine of Alexandria, Virgin Martyress, \textit{Double}\\
	26&Pope St. Peter of Alexandria, Bishop \& Martyr, \textit{Memorial}\\
	27&\\
	28&\\
	29&Vigil of St. Andrew, \textit{Vigil}\\
	&St. Saturninus of Toulouse, Bishop \& Martyr, \textit{Memorial}\\
	30&{\color{RubricRed}St. Andrew, Apostle,} \textit{\nth{2} Double}\\
	%
	\SetCell[c=2]{l} \textsc{December}\\
	1&\\
	2&St. Peter Chrysologus, Bishop, Confessor, \& Doctor, \textit{Double}\\
	&St. Bibiana, Virgin Martyress, \textit{Memorial}\\
	3&\\
	4&St. Barbara, Virgin Martyress, \textit{Memorial}\\
	5&St. Sabbas of Jud{\ae}a, Abbot	, \textit{Memorial}\\
	6&St. Nicholas of Myra, Bishop \& Confessor, \textit{Double}\\
	7&St. Ambrose of Milan, Bishop, Confessor, \& Doctor, \textit{Greater Double}\\
	8&{\color{RubricRed}Conception of the Blessed Virgin Mary,} \textit{\nth{2} Double,} Common Octave\\
	9&Day II within the Octave of the Conception, \textit{Semidouble}\\
	10&Day III within the Octave of the Conception, \textit{Semidouble}\\
	&Pope St. Melchiades of Rome, Bishop \& Martyr, \textit{Memorial}\\
	11&Day IV within the Octave of the Conception, \textit{Semidouble}\\
	&Pope St. Damasus I of Rome, Bishop \& Confessor, \textit{Memorial}\\
	12&Day V within the Octave of the Conception, \textit{Semidouble}\\
	13&St. Lucy, Virgin Martyress, \textit{Greater Double}\\
	&Day VI within the Octave of the Conception, \textit{Semidouble}\\
	&St. Herman of Alaska, Confessor, \textit{Memorial}\\
	14&Day VII within the Octave of the Conception, \textit{Semidouble}\\
	15&Octave Day of the Conception of the Blessed Virgin Mary, \textit{Greater Double}\\
	16&St. Eusebius of Vercelli, Bishop \& Martyr, \textit{Memorial}\\
	17&\\
	18&Expectation of the Blessed Virgin Mary, \textit{Greater Double}\\
	19&\\
	20&Vigil of St. Thomas, \textit{Vigil}\\
	21&{\color{RubricRed}St. Thomas, Apostle,} \textit{\nth{2} Double}\\
	22&\\
	23&\\
	24&{\color{RubricRed}Vigil of the Nativity of Our Lord,} \textit{\nth{1} Vigil}\\
	25&{\color{RubricRed}Nativity of Our Lord,} \textit{\nth{1} Double,} \nth{3} Octave\\
	&St. Anastasia, Virgin Martyress, \textit{Commemoration}\\
	26&{\color{RubricRed}St. Stephen, Protomartyr,} \textit{\nth{2} Double,} Simple Octave\\
	&Day II within the Nativity Octave, \textit{Semidouble}\\
	27&{\color{RubricRed}St. John, Apostle \& Evangelist,} \textit{\nth{2} Double,} Simple Octave\\
	&Day III within the Nativity Octave, \textit{Semidouble}\\
	28&{\color{RubricRed}Holy Innocents, Martyrs,} \textit{\nth{2} Double,} Simple Octave\\
	&Day IV within the Nativity Octave, \textit{Semidouble}\\
	29&Day V within the Nativity Octave, \textit{Semidouble}\\
	30&Day VI within the Nativity Octave, \textit{Semidouble}\\
	31&Pope St. Sylvester of Rome, Bishop \& Confessor, \textit{Double}\\
	&Day VII within the Nativity Octave, \textit{Semidouble}\\
\end{longtblr}