\fancyhead[RO,LE]{\textit{}}
\fancyhead[RE,LO]{}
\fancyhead[C]{\LARGE Easter Even}
    
\section*{Blessing of the Paschal Fire}
\fancyhead[RO,LE]{\textit{Paschal Fire}}
\begin{rubric}
	At a suitable time the Altars are covered with linen cloths, but the candles remain unlit until the beginning of Mass. Meanwhile fire is struck from a flint outside the church, and from it coals are kindled. After Mattins (or, None) has been said in Choir, the Priest, vested in Amice, Alb, Girdle, Stole, and violet Cope, or without a Chasuble, standing with the Ministers, with the Cross, holy water, and incense, before the door of the Church, if it can conveniently be done, or in the entry of the church itself, blesses the new fire, saying:
\end{rubric}

℣. The Lord be with you.

℟. And with thy spirit.

\letuspray
\lett{O}{God,} who through thy Son, the true corner-stone, hast bestowed upon thy faithful people the fire of thy brightness: sanct~{\ding{64}}~ify this new fire, now struck from the flint-stone, to be profitable to our service: and grant unto us that by this paschal feast we may be so inflamed with heavenly desires; that we may with pure hearts attain unto the feast of thy eternal brightness. Through the same Christ, our Lord. \textit{Amen.}

\letuspray
\lett{O}{Lord} God, Father almighty, who art light unfailing, and the author of all lights: bl~{\ding{64}}~ess this light which hath here been sanctified and blessed by thee, who hast enlightened the whole world: that we may be enkindled by that light, and enlightened with the fire of thy brightness: and like as thou didst enlighten Moses going forth from Egypt, so do thou enlighten our hearts and understandings; that we may be found worthy to attain unto life and light everlasting. Through Christ, our Lord. \textit{Amen.}

\letuspray
\lett{O}{Lord} holy, Father almighty, everlasting God: vouchsafe to assist us who bless this fire in thy name, and in the name of thy only-begotten Son, Jesus Christ, our God and Lord, and of the Holy Ghost; and assist us against the fiery darts of the enemy, and enlighten us with thy heavenly grace: Who livest and reignest with the same thy only-begotten Son and the Holy Ghost, God: throughout all ages world without end. \textit{Amen.}

%MANUAL ADJUSTMENT:
\clearpage
\begin{rubric}
	He then blesses the five grains of incense, to be placed in the Candle, saying at once this Collect:
\end{rubric}
\lett{W}{e} beseech thee, almighty God, that the abundance of thy bles~{\ding{64}}~sing may come down upon this incense; and as thou, thyself unseen, art the regenerator of all things, so kindle the brightness of this night; that not only may our sacrifice, which is offered on this night, shine with the secret mingling of thy light; but that in all places wheresoever aught shall be carried from this mystery here sanctified, the wickedness of the crafts of the devil may be driven forth, and the power of thy majesty ever be present. Through Christ, our Lord. \textit{Amen.}

\section*{Procession of the Light of Christ}
\fancyhead[RO,LE]{\textit{Light of Christ}}

\begin{rubric}
	Whilst be blesses the grains of incense, an Acolyte takes some of the blessed coals, and sets them in the thurible: and the Priest, when he has finished the last Prayer, puts incense from the boat into the thurible, giving the usual blessing. Then he sprinkles the aforesaid grains of incense and the fire with holy water, saying: \blackrubric{Thou shalt purge me, O Lord,} without note and without the Psalm, and censes them thrice.

	Meanwhile all the lights in the church are extinguished, to be relighted later from the blessed fire.
	
	Then the Deacon, vested in a white Dalmatic, takes the reed upon the top of which are three candles in the shape of a triangle. The Thurifer goes first with an Acolyte bearing the five grains of incense: the Subdeacon follow's with the Cross, the Clergy in their order: then the Deacon with the reed, after him the Celebrant.
	
	When the Deacon enters the church, he lowers the reed, and an Acolyte, carrying a candle which has been lighted from the new fire, lights one of the three candles placed upon it, and the Deacon, elevating the reed, genuflects, as do all others with him except the Subdeacon bearing the Cross, and sings alone: %\blackrubric{The Light of Christ}.
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/LumenChristi.gabc}

%℟. Thanks be to God.

\begin{rubric}
	And proceeding to the middle of the church, another candle is lighted: and having again genuflected, he sings in a higher tone: \blackrubric{The Light of Christ.}
\end{rubric}

℟. Thanks be to God.

\begin{rubric}
	For the third time he proceeds before the Altar, where the third candle is lighted: and again, having genuflected as before, he sings still higher: \blackrubric{The Light of Christ.}
\end{rubric}

℟. Thanks be to God.

\begin{rubric}
	The Celebrant then ascends to the Epistle corner of the Altar, and the Deacon, banding the reed with the triple candle to an Acolyte, receives the book, and asks a blessing from the Celebrant as at the Gospel, the Priest saying:
\end{rubric}
\par\noindent
The Lord be in thy heart, and on thy lips: that thou mayest worthily and fitly proclaim his paschal praise: In the name of the Father, and of the Son, {\ding{64}} and of the Holy Ghost. Amen.

\begin{rubric}
	The Deacon then goes to the desk and placing the book upon it, censes it. On the right hand of the Deacon stand the Subdeacon with the Cross, and the Thurifer: on his left an Acolyte holding the reed, and another the five grains of blessed incense to be fixed into the Candle. Then, all rising and standing, as at the Gospel, the Deacon sings:
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/AngelicHost.gabc}

%\lett{N}{ow} let the Angelic host of heaven rejoice: let the divine mysteries rejoice: and for the victory of the mighty King let the trumpet of salvation sound forth. Let the earth also be glad, illumined by the rays of this great brightness: and enlightened by the splendour of the eternal King, let her know that she hath put away the darkness of the whole world. Let our mother the Church also rejoice, adorned with the brightness of so great a light: and let this temple resound with the triumphant voices of the peoples. Wherefore, dearly beloved brethren, who are here present in the wondrous clearness of this holy light, join with me, I beseech you, in calling upon the mercy of almighty God. That he who hath been pleased, for no merit of mine, to admit me into the number of his Levites: may pour on me the brightness of his light, and make me meet to proclaim the praises of this Candle. Through Jesus Christ his Son our Lord: Who liveth and reigneth with him in the unity of the Holy Ghost, God. Throughout all ages, world without end. \textit{Amen.}\\

\gregorioscore{resources/gabc/Propers/EasterEven/SursumCorda.gabc}

\gregorioscore{resources/gabc/Propers/EasterEven/CandlePreface.gabc}

%℣. The Lord be with you.
%
%℟. And with thy spirit.
%
%℣. Lift up your hearts.
%
%℟. We lift them up unto the Lord.
%
%℣. Let us give thanks unto our Lord God.
%
%℟. It is meet and right so to do.\\
%\par\noindent
%It is very meet, right, and our bounden duty, that we should, with the whole affection of our heart and mind, and with the service of our lips, give praise unto the invisible God, the Father almighty, and unto his only begotten Son, Jesus Christ our Lord. Who paid for us to the eternal Father the debt of Adam's transgression: and with his dear blood wiped away the reproach of our former offences. For this is the paschal feast wherein the very Lamb is slain, by whose blood the door-posts of the faithful are made holy. This is the night, wherein of old thou didst lead forth our fathers, the children of Israel, out of Egypt, and didst make them to pass as on dry land through the Red Sea. Yea, this is the night, that with the brightness of the pillar hath purged away the darkness of iniquity. This is the night, which throughout the whole world today doth separate them that believe in Christ from the wickedness of the age and from the darkness of transgressions, reneweth them unto grace, restoreth them to holiness. This is the night wherein, breaking the chains of death, Christ ascendeth from hell in triumph. For nought indeed had it profited us to be born, if it had not profited us to be redeemed. O how wonderful the condescension of thy loving kindness! O how inestimable the goodness of thy love: who to redeem a slave didst deliver up thy Son! O truly necessary sin of Adam, which by the death of Christ was done away! O happy fault, which was counted worthy to have such and so great a Redeemer! O night truly blessed, which alone was worthy to know the time and the hour wherein Christ rose again from hell! This is the night whereof it is written: And the night is as clear as the day: And, then shall my night be turned into day. The sanctifying power therefore of this night putteth to flight the deeds of wickedness, washeth away sins: restoreth innocence to the fallen, and joy to them that mourn: casteth out enmities, prepareth concord, and boweth down principalities.

\begin{rubric}
	Here the Deacon fixes the five grains of incense in the Candle, placing them in the form of a cross, in this order:
\end{rubric}

\begin{center}
	1
	
	4 2 5
	
	3
\end{center}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/WorkOfBees.gabc}
%Therefore in this night of grace accept, O holy Father, the evening sacrifice of this incense: which, by the hands of thy ministers, holy Church doth lay before thee, in the solemn offering of this Candle, made from the work of bees. But we already know the excellency of this pillar, which for the honour of God the sparkling fire doth kindle.

\begin{rubric}
	Here the Deacon lights the Candle with one of the three candles set on the reed.
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/NoLoss.gabc}
%Which, though it be divided into parts, suffereth not loss by the borrowing of its light. For it is fed by melting wax, which the bee the mother hath wrought into the substance of this precious candle.

\begin{rubric}
	Here the lamps are lighted.
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/BlissfulNight.gabc}
%O night truly blessed, which did spoil the Egyptians, and made rich the Hebrews! O night wherein heavenly things are joined unto earthly, things human unto things divine. We therefore pray thee, O Lord: that this Candle, consecrated to the honour of thy name, may continue without ceasing to vanquish the darkness of this night. That, being accepted for a savour of sweetness, it may be mingled with the lights of heaven. May the morning star find it burning. That morning star, I say, which knoweth not his going down. That star, which, rising again from hell, steadfastly giveth light to all mankind. We therefore pray thee, O Lord: that thou wouldest vouchsafe to rule, govern and preserve with thy continual protection us thy servants, the whole clergy and thy most faithful people: together with our most blessed Father \textit{N.} (and our Bishop \textit{N.}), granting us peaceful times in this our paschal joy. Through the same Jesus Christ thy Son our Lord: Who liveth and reigneth with thee in the unity of the Holy Ghost God: throughout all ages, world without end. \textit{Amen.}

%MANUAL ADJUSTMENT:
\clearpage
\section*{The Prophecies}
\fancyhead[RO,LE]{\textit{Prophecies}}

%MANUAL ADJUSTMENT:
\vspace{-0.3\baselineskip}

\begin{rubric}
	The blessing of the Candle ended, the Deacon lays aside his white vestments, and puts on violet, and goes to the Celebrant: who, laying aside his Cope, puts on a Maniple and violet Chasuble. Then the Prophecies are read without title, and at the end the response \blackrubric{Thanks be to God} is not made, and the Celebrant reads them in a low voice at the Epistle corner of the Altar.
	
%	At the end of the Prophecies are said the Prayers as given below.
	
	Before, or during the reading of the Prophecies, the Priests catechise any Catechumens about to be baptised, and prepare them for Baptism.
\end{rubric}

%MANUAL ADJUSTMENT:
\vspace{-0.5\baselineskip}

\HWreadingcitation{Prophecy I}{Genesis 1:1}

%MANUAL ADJUSTMENT:
\vspace{-0.25\baselineskip}

\lett{I}{n} the beginning God created the heaven and the earth. And the earth was without form, and void; and darkness was upon the face of the deep. And the Spirit of God moved upon the face of the waters. And God said, Let there be light: and there was light. And God saw the light, that it was good: and God divided the light from the darkness. And God called the light Day, and the darkness he called Night. And the evening and the morning were the first day.

And God said, Let there be a firmament in the midst of the waters, and let it divide the waters from the waters. And God made the firmament, and divided the waters which were under the firmament from the waters which were above the firmament: and it was so. And God called the firmament Heaven. And the evening and the morning were the second day.

And God said, Let the waters under the heaven be gathered together unto one place, and let the dry land appear: and it was so. And God called the dry land Earth; and the gathering together of the waters called he Seas: and God saw that it was good. And God said, Let the earth bring forth grass, the herb yielding seed, and the fruit tree yielding fruit after his kind, whose seed is in itself, upon the earth: and it was so. And the earth brought forth grass, and herb yielding seed after his kind, and the tree yielding fruit, whose seed was in itself, after his kind: and God saw that it was good. And the evening and the morning were the third day.

And God said, Let there be lights in the firmament of the heaven to divide the day from the night; and let them be for signs, and for seasons, and for days, and years: and let them be for lights in the firmament of the heaven to give light upon the earth: and it was so. And God made two great lights; the greater light to rule the day, and the lesser light to rule the night: he made the stars also. And God set them in the firmament of the heaven to give light upon the earth, and to rule over the day and over the night, and to divide the light from the darkness: and God saw that it was good. And the evening and the morning were the fourth day.

And God said, Let the waters bring forth abundantly the moving creature that hath life, and fowl that may fly above the earth in the open firmament of heaven. And God created great whales, and every living creature that moveth, which the waters brought forth abundantly, after their kind, and every winged fowl after his kind: and God saw that it was good. And God blessed them, saying, Be fruitful, and multiply, and fill the waters in the seas, and let fowl multiply in the earth. And the evening and the morning were the fifth day.

And God said, Let the earth bring forth the living creature after his kind, cattle, and creeping thing, and beast of the earth after his kind: and it was so. And God made the beast of the earth after his kind, and cattle after their kind, and every thing that creepeth upon the earth after his kind: and God saw that it was good.

And God said, Let us make man in our image, after our likeness: and let them have dominion over the fish of the sea, and over the fowl of the air, and over the cattle, and over all the earth, and over every creeping thing that creepeth upon the earth. So God created man in his own image, in the image of God created he him; male and female created he them. And God blessed them, and God said unto them, Be fruitful, and multiply, and replenish the earth, and subdue it: and have dominion over the fish of the sea, and over the fowl of the air, and over every living thing that moveth upon the earth.

And God said, Behold, I have given you every herb bearing seed, which is upon the face of all the earth, and every tree, in the which is the fruit of a tree yielding seed; to you it shall be for meat. And to every beast of the earth, and to every fowl of the air, and to every thing that creepeth upon the earth, wherein there is life, I have given every green herb for meat: and it was so. And God saw every thing that he had made, and, behold, it was very good. And the evening and the morning were the sixth day.

Thus the heavens and the earth were finished, and all the host of them. And on the seventh day God ended his work which he had made; and he rested on the seventh day from all his work which he had made.

\begin{rubric}
	\blackrubric{Thanks be to God} is not said either here or after the other Prophecies.
\end{rubric}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who didst wonderfully create man, and hast more wonderfully redeemed him: grant us, we beseech thee, such strength of mind to withstand the enticements of sin; that we may be found worthy to attain to everlasting joys. Through Jesus Christ thy Son our Lord. Who liveth.

\HWreadingcitation{Prophecy II}{Genesis 5:32}
\lett{A}{nd} Noah was five hundred years old: and Noah begat Shem, Ham, and Japheth. And it came to pass, when men began to multiply on the face of the earth, and daughters were born unto them, that the sons of God saw the daughters of men that they were fair; and they took them wives of all which they chose. And the Lord said, My spirit shall not always strive with man, for that he also is flesh: yet his days shall be an hundred and twenty years. There were giants in the earth in those days; and also after that, when the sons of God came in unto the daughters of men, and they bare children to them, the same became mighty men which were of old, men of renown.

And God saw that the wickedness of man was great in the earth, and that every imagination of the thoughts of his heart was only evil continually. And it repented the Lord that he had made man on the earth, and it grieved him at his heart. And the Lord said, I will destroy man whom I have created from the face of the earth; both man, and beast, and the creeping thing, and the fowls of the air; for it repenteth me that I have made them. But Noah found grace in the eyes of the Lord.

These are the generations of Noah: Noah was a just man and perfect in his generations, and Noah walked with God. And Noah begat three sons, Shem, Ham, and Japheth. The earth also was corrupt before God, and the earth was filled with violence. And God looked upon the earth, and, behold, it was corrupt; for all flesh had corrupted his way upon the earth. And God said unto Noah, The end of all flesh is come before me; for the earth is filled with violence through them; and, behold, I will destroy them with the earth. Make thee an ark of gopher wood; rooms shalt thou make in the ark, and shalt pitch it within and without with pitch. And this is the fashion which thou shalt make it of: The length of the ark shall be three hundred cubits, the breadth of it fifty cubits, and the height of it thirty cubits. A window shalt thou make to the ark, and in a cubit shalt thou finish it above; and the door of the ark shalt thou set in the side thereof; with lower, second, and third stories shalt thou make it. And, behold, I, even I, do bring a flood of waters upon the earth, to destroy all flesh, wherein is the breath of life, from under heaven; and every thing that is in the earth shall die. But with thee will I establish my covenant; and thou shalt come into the ark, thou, and thy sons, and thy wife, and thy sons' wives with thee. And of every living thing of all flesh, two of every sort shalt thou bring into the ark, to keep them alive with thee; they shall be male and female. Of fowls after their kind, and of cattle after their kind, of every creeping thing of the earth after his kind, two of every sort shall come unto thee, to keep them alive. And take thou unto thee of all food that is eaten, and thou shalt gather it to thee; and it shall be for food for thee, and for them. Thus did Noah; according to all that God commanded him, so did he. 

And Noah was six hundred years old when the flood of waters was upon the earth. The same day were all the fountains of the great deep broken up, and the windows of heaven were opened. And the rain was upon the earth forty days and forty nights. In the selfsame day entered Noah, and Shem, and Ham, and Japheth, the sons of Noah, and Noah's wife, and the three wives of his sons with them, into the ark; they, and every beast after his kind, and all the cattle after their kind, and every creeping thing that creepeth upon the earth after his kind, and every fowl after his kind, every bird of every sort. And the ark went upon the face of the waters. And the waters prevailed exceedingly upon the earth; and all the high hills, that were under the whole heaven, were covered. Fifteen cubits upward did the waters prevail; and the mountains were covered. And all flesh died that moved upon the earth, both of fowl, and of cattle, and of beast, and of every creeping thing that creepeth upon the earth. And Noah only remained alive, and they that were with him in the ark. And the waters prevailed upon the earth an hundred and fifty days.

And God remembered Noah, and every living thing, and all the cattle that was with him in the ark: and God made a wind to pass over the earth, and the waters asswaged; the fountains also of the deep and the windows of heaven were stopped, and the rain from heaven was restrained; and the waters returned from off the earth continually: and after the end of the hundred and fifty days the waters were abated.

And it came to pass at the end of forty days, that Noah opened the window of the ark which he had made: and he sent forth a raven, which went forth to and fro, until the waters were dried up from off the earth. Also he sent forth a dove from him, to see if the waters were abated from off the face of the ground; but the dove found no rest for the sole of her foot, and she returned unto him into the ark, for the waters were on the face of the whole earth: then he put forth his hand, and took her, and pulled her in unto him into the ark. And he stayed yet other seven days; and again he sent forth the dove out of the ark; and the dove came in to him in the evening; and, lo, in her mouth was an olive leaf pluckt off: so Noah knew that the waters were abated from off the earth. And he stayed yet other seven days; and sent forth the dove; which returned not again unto him any more.

And God spake unto Noah, saying, Go forth of the ark, thou, and thy wife, and thy sons, and thy sons' wives with thee. Bring forth with thee every living thing that is with thee, of all flesh, both of fowl, and of cattle, and of every creeping thing that creepeth upon the earth; that they may breed abundantly in the earth, and be fruitful, and multiply upon the earth. And Noah went forth, and his sons, and his wife, and his sons' wives with him: every beast, every creeping thing, and every fowl, and whatsoever creepeth upon the earth, after their kinds, went forth out of the ark.

And Noah builded an altar unto the Lord; and took of every clean beast, and of every clean fowl, and offered burnt offerings on the altar. And the Lord smelled a sweet savour.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who art strength unchangeable and light eternal: look down in mercy on the wondrous mystery of thy whole Church, and by the operation of thy continual providence accomplish in tranquillity the work of man's salvation; and let the whole world perceive and know that things cast down are being raised up, things grown old are being made new, and all things are returning to their perfection, through him from whom they took their beginning: Jesus Christ thy Son our Lord: Who liveth and reigneth with thee.

\HWreadingcitation{Prophecy III}{Genesis 22:1}
\lett{I}{n those days:} God did tempt Abraham, and said unto him, Abraham: and he said, Behold, here I am. And he said, Take now thy son, thine only son Isaac, whom thou lovest, and get thee into the land of Moriah; and offer him there for a burnt offering upon one of the mountains which I will tell thee of.

And Abraham rose up early in the morning, and saddled his ass, and took two of his young men with him, and Isaac his son, and clave the wood for the burnt offering, and rose up, and went unto the place of which God had told him. Then on the third day Abraham lifted up his eyes, and saw the place afar off. And Abraham said unto his young men, Abide ye here with the ass; and I and the lad will go yonder and worship, and come again to you. And Abraham took the wood of the burnt offering, and laid it upon Isaac his son; and he took the fire in his hand, and a knife; and they went both of them together. And Isaac spake unto Abraham his father, and said, My father: and he said, Here am I, my son. And he said, Behold the fire and the wood: but where is the lamb for a burnt offering? And Abraham said, My son, God will provide himself a lamb for a burnt offering: so they went both of them together. And they came to the place which God had told him of; and Abraham built an altar there, and laid the wood in order, and bound Isaac his son, and laid him on the altar upon the wood. And Abraham stretched forth his hand, and took the knife to slay his son. And the angel of the Lord called unto him out of heaven, and said, Abraham, Abraham: and he said, Here am I. And he said, Lay not thine hand upon the lad, neither do thou any thing unto him: for now I know that thou fearest God, seeing thou hast not withheld thy son, thine only son from me. And Abraham lifted up his eyes, and looked, and behold behind him a ram caught in a thicket by his horns: and Abraham went and took the ram, and offered him up for a burnt offering in the stead of his son. And Abraham called the name of that place Jehovah-jireh: as it is said to this day, In the mount of the Lord it shall be seen.

And the angel of the Lord called unto Abraham out of heaven the second time, and said, By myself have I sworn, saith the Lord, for because thou hast done this thing, and hast not withheld thy son, thine only son: that in blessing I will bless thee, and in multiplying I will multiply thy seed as the stars of the heaven, and as the sand which is upon the sea shore; and thy seed shall possess the gate of his enemies; and in thy seed shall all the nations of the earth be blessed; because thou hast obeyed my voice. So Abraham returned unto his young men, and they rose up and went together to Beer-sheba; and Abraham dwelt at Beer-sheba.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} the sovereign Father of the faithful, who dost spread abroad throughout all the world the grace of adoption, to the increasing of the sons of thy promise: and, according as thou didst swear unto thy servant Abraham, dost through this paschal sacrament make him the father of all nations; grant that thy peoples may worthily enter upon the grace of thy calling. Through.

\HWreadingcitation{Prophecy IV}{Exodus 14:24}
\lett{I}{n those days:} It came to pass, that in the morning watch the Lord looked unto the host of the Egyptians through the pillar of fire and of the cloud, and troubled the host of the Egyptians, and took off their chariot wheels, that they drave them heavily: so that the Egyptians said, Let us flee from the face of Israel; for the Lord fighteth for them against the Egyptians.

 And the Lord said unto Moses, Stretch out thine hand over the sea, that the waters may come again upon the Egyptians, upon their chariots, and upon their horsemen. And Moses stretched forth his hand over the sea, and the sea returned to his strength when the morning appeared; and the Egyptians fled against it; and the Lord overthrew the Egyptians in the midst of the sea. And the waters returned, and covered the chariots, and the horsemen, and all the host of Pharaoh that came into the sea after them; there remained not so much as one of them. But the children of Israel walked upon dry land in the midst of the sea; and the waters were a wall unto them on their right hand, and on their left. Thus the Lord saved Israel that day out of the hand of the Egyptians; and Israel saw the Egyptians dead upon the sea shore. And Israel saw that great work which the Lord did upon the Egyptians: and the people feared the Lord, and believed the Lord, and his servant Moses.

Then sang Moses and the children of Israel this song unto the Lord, and spake, saying,

\tract{We will sing unto the Lord: for he hath triumphed gloriously: the horse and his rider hath he thrown into the sea: the Lord is my strength and song, and he is become my salvation. ℣. He is my God, and I will praise him: my father's God, and I will exalt him. ℣. The Lord is a man of war: the Lord is his name.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} whose miracles of old we perceive to shine forth even in our times: who by the water of regeneration dost now work for the salvation of the nations, as thou didst deliver one people from the pursuit of the Egyptians by the power of thy right hand: vouchsafe; that the fulness of the whole world may be numbered among the sons of Abraham and made partakers of the dignity of Israel. Through.

\HWreadingcitation{Prophecy V}{Isaiah 54:17}
\lett{T}{his} is the heritage of the servants of the \divineName{Lord}: and their righteousness is of me, saith the Lord.

\secondline{Ho, every one that thirsteth, come ye to the waters,
and he that hath no money; come ye, buy, and eat;
yea, come, buy wine and milk without money and without price.
Wherefore do ye spend money for that which is not bread?
and your labour for that which satisfieth not?
hearken diligently unto me, and eat ye that which is good,
and let your soul delight itself in fatness.
Incline your ear, and come unto me:
hear, and your soul shall live;
and I will make an everlasting covenant with you,
even the sure mercies of David.
Behold, I have given him for a witness to the people,
a leader and commander to the people.
Behold, thou shalt call a nation that thou knowest not,
and nations that knew not thee shall run unto thee
because of the Lord thy God, and for the Holy One of Israel;
for he hath glorified thee.}

Seek ye the Lord while he may be found,
call ye upon him while he is near:
let the wicked forsake his way,
and the unrighteous man his thoughts:
and let him return unto the Lord,
and he will have mercy upon him;
and to our God, for he will abundantly pardon.

For my thoughts are not your thoughts,
neither are your ways my ways, saith the Lord.
For as the heavens are higher than the earth,
so are my ways higher than your ways,
and my thoughts than your thoughts.
For as the rain cometh down, and the snow from heaven,
and returneth not thither, but watereth the earth,
and maketh it bring forth and bud,
that it may give seed to the sower, and bread to the eater:
so shall my word be that goeth forth out of my mouth:
it shall not return unto me void,
but it shall accomplish that which I please,
and it shall prosper in the thing whereto I sent it, saith the Lord almighty.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{A}{lmighty} and everlasting God, multiply to the honour of thy name what thou didst swear to the faith of the fathers: and spread abroad the sons of promise by holy adoption; that as the Saints of old doubted not thine oath, so thy Church may now see in abundant measure the fulfilment of the same. Through.

\HWreadingcitation{Prophecy VI}{Baruch 3:9}

\lett{H}{ear,} O Israel, the commandments of life: give ear to understand wisdom. How happeneth it, O Israel, that thou art in thine enemies' land, that thou art waxen old in a strange country, that thou art defiled with the dead, that thou art counted with them that go down into the grave? Thou hast forsaken the fountain of wisdom. For if thou hadst walked in the way of God, thou shouldest have dwelled in peace for ever. Learn where is wisdom, where is strength, where is understanding; that thou mayest know also where is length of days, and life, where is the light of the eyes, and peace.

Who hath found out her place? and who hath come into her treasuries? Where are the princes of the heathen, and such as ruled the beasts that are upon the earth; they that had their pastime with the fowls of the air, and they that hoarded up silver and gold, wherein men trust; and of whose getting there is no end? For they that wrought in silver, and were so careful, and whose works are past finding out, they are vanished and gone down to the grave, and others are come up in their steads. Younger men have seen the light, and dwelt upon the earth: but the way of knowledge have they not known, neither understood they the paths thereof: neither have their children laid hold of it: they are far off from their way. It hath not been heard of in Canaan, neither hath it been seen in Teman. The sons also of Agar that seek understanding, which are in the land, the merchants of Merran and Teman, and the authors of fables, and the searchers out of understanding; none of these have known the way of wisdom, or remembered her paths.

O Israel, how great is the house of God! and how large is the place of his possession! great, and hath none end; high, and unmeasurable. There were the giants born that were famous of old, great of stature, and expert in war. These did not God choose, neither gave he the way of knowledge unto them: so they perished, because they had no wisdom, they perished through their own foolishness. Who hath gone up into heaven, and taken her, and brought her down from the clouds? Who hath gone over the sea, and found her, and will bring her for choice gold? There is none that knoweth her way, nor any that comprehendeth her path.

But he that knoweth all things knoweth her, he found her out with his understanding: he that prepared the earth for evermore hath filled it with fourfooted beasts: he that sendeth forth the light, and it goeth; he called it, and it obeyed him with fear: and the stars shined in their watches, and were glad: when he called them, they said, Here we be; they shined with gladness unto him that made them. This is our God, and there shall none other be accounted of in comparison of him. He hath found out all the way of knowledge, and hath given it unto Jacob his servant, and to Israel that is beloved of him. Afterward did she appear upon earth, and was conversant with men. 

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who by the calling of the nations dost ever multiply thy Church: mercifully grant; that they whom thou dost cleanse by the waters of baptism may be defended by thy continual protection. Through.

\HWreadingcitation{Prophecy VII}{Ezekiel 37:1}
\lett{I}{n those days:} The hand of the \divineName{Lord} was upon me, and carried me out in the spirit of the Lord, and set me down in the midst of the valley which was full of bones, and caused me to pass by them round about: and, behold, there were very many in the open valley; and, lo, they were very dry. And he said unto me, Son of man, can these bones live? And I answered, O Lord God, thou knowest.

Again he said unto me, Prophesy upon these bones, and say unto them, O ye dry bones, hear the word of the Lord. Thus saith the Lord God unto these bones; Behold, I will cause breath to enter into you, and ye shall live: and I will lay sinews upon you, and will bring up flesh upon you, and cover you with skin, and put breath in you, and ye shall live; and ye shall know that I am the Lord.

So I prophesied as I was commanded: and as I prophesied, there was a noise, and behold a shaking, and the bones came together, bone to his bone. And when I beheld, lo, the sinews and the flesh came up upon them, and the skin covered them above: but there was no breath in them. Then said he unto me, Prophesy unto the wind, prophesy, son of man, and say to the wind, Thus saith the Lord God; Come from the four winds, O breath, and breathe upon these slain, that they may live. So I prophesied as he commanded me, and the breath came into them, and they lived, and stood up upon their feet, an exceeding great army.

Then he said unto me, Son of man, these bones are the whole house of Israel: behold, they say, Our bones are dried, and our hope is lost: we are cut off for our parts. Therefore prophesy and say unto them, Thus saith the Lord God; Behold, O my people, I will open your graves, and cause you to come up out of your graves, and bring you into the land of Israel. And ye shall know that I am the Lord, when I have opened your graves, O my people, and brought you up out of your graves, and shall put my spirit in you, and ye shall live, and I shall place you in your own land, faith the Lord almighty.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who dost instruct us in the pages of either Testament in the celebrating of the paschal sacrament: grant unto us so to understand thy mercy; that, through the receiving of these present gifts, we may be established in the expectation of those which are to come. Through.

\HWreadingcitation{Prophecy VIII}{Isaiah 4:1}
\lett{I}{n that day} seven women shall take hold of one man, saying: We will eat our own bread, and wear our own apparel: only let us be called by thy name, to take away our reproach. In that day shall the branch of the Lord be beautiful and glorious, and the fruit of the earth shall be excellent and comely for them that are escaped of Israel. And it shall come to pass, that he that is left in Zion, and he that remaineth in Jerusalem, shall be called holy, even every one that is written among the living in Jerusalem: when the Lord shall have washed away the filth of the daughters of Zion, and shall have purged the blood of Jerusalem from the midst thereof by the spirit of judgment, and by the spirit of burning. And the Lord will create upon every dwelling place of mount Zion, and upon her assemblies, a cloud and smoke by day, and the shining of a flaming fire by night: for upon all the glory shall be a defence. And there shall be a tabernacle for a shadow in the daytime from the heat, and for a place of refuge, and for a covert from storm and from rain.

\tract{My well-beloved hath a vineyard in a very fruitful hill. ℣. And he fenced it, and gathered out the stones thereof: and planted it with the choicest vine, and built a tower in the midst of it. ℣. And also made a wine-press therein: for the vineyard of the Lord of hosts is the house of Israel.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who by the mouth of thy holy prophets hast manifested thyself in all the children of thy Church, to be in all places of thy dominion the sower of good seed and the husbandman of thine elect branches: grant unto thy peoples, who are named thy vineyard and thy harvest field; that being purged from the filthiness of thorns and briars, they may be made to bring forth worthy fruit in abundance. Through.

\HWreadingcitation{Prophecy IX}{Exodus 12:1}
\lett{I}{n those days:} The \divineName{Lord} spake unto Moses and Aaron in the land of Egypt, saying, This month shall be unto you the beginning of months: it shall be the first month of the year to you.

Speak ye unto all the congregation of Israel, saying, In the tenth day of this month they shall take to them every man a lamb, according to the house of their fathers, a lamb for an house: and if the household be too little for the lamb, let him and his neighbour next unto his house take it according to the number of the souls; every man according to his eating shall make your count for the lamb. Your lamb shall be without blemish, a male of the first year: ye shall take it out from the sheep, or from the goats: and ye shall keep it up until the fourteenth day of the same month: and the whole assembly of the congregation of Israel shall kill it in the evening. And they shall take of the blood, and strike it on the two side posts and on the upper door post of the houses, wherein they shall eat it. And they shall eat the flesh in that night, roast with fire, and unleavened bread; and with bitter herbs they shall eat it. Eat not of it raw, nor sodden at all with water, but roast with fire; his head with his legs, and with the purtenance thereof. And ye shall let nothing of it remain until the morning; and that which remaineth of it until the morning ye shall burn with fire.

And thus shall ye eat it; with your loins girded, your shoes on your feet, and your staff in your hand; and ye shall eat it in haste: it is the Lord’s passover.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{A}{lmighty} and everlasting God, who art wondrous in the dispensation of all thy works: let thy redeemed understand that the creation of the world in the beginning was not a work more excellent than the sacrifice, in the end of time, of Christ our Passover: Who liveth and reigneth with thee.

\HWreadingcitation{Prophecy X}{Jonah 3:1}
\lett{I}{n those days:} The word of the Lord came unto Jonah the Prophet the second time, saying: Arise, go unto Nineveh, that great city, and preach unto it the preaching that I bid thee. So Jonah arose, and went unto Nineveh, according to the word of the Lord. Now Nineveh was an exceeding great city of three days’ journey. And Jonah began to enter into the city a day’s journey, and he cried, and said, Yet forty days, and Nineveh shall be overthrown.

So the people of Nineveh believed God, and proclaimed a fast, and put on sackcloth, from the greatest of them even to the least of them. For word came unto the king of Nineveh, and he arose from his throne, and he laid his robe from him, and covered him with sackcloth, and sat in ashes. And he caused it to be proclaimed and published through Nineveh by the decree of the king and his nobles, saying, Let neither man nor beast, herd nor flock, taste any thing: let them not feed, nor drink water: but let man and beast be covered with sackcloth, and cry mightily unto God: yea, let them turn every one from his evil way, and from the violence that is in their hands. Who can tell if God will turn and repent, and turn away from his fierce anger, that we perish not?

And God saw their works, that they turned from their evil way; and he had mercy on his people, even he the Lord, our God.

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} who hast united the diversity of nations in the confession of thy name: grant us both to will, and to have power to do the things which thou commandest; that the people whom thou hast called to everlasting life may agree in unity of faith and in godliness of conversation. Through.

\HWreadingcitation{Prophecy XI}{Deuteronomy 31:22}
\lett{I}{n those days:} Moses wrote this song, and taught it the children of Israel. And he gave Joshua the son of Nun a charge, and said, Be strong and of a good courage: for thou shalt bring the children of Israel into the land which I sware unto them: and I will be with thee.

And it came to pass, when Moses had made an end of writing the words of this law in a book, until they were finished, that Moses commanded the Levites, which bare the ark of the covenant of the Lord, saying, Take this book of the law, and put it in the side of the ark of the covenant of the Lord your God, that it may be there for a witness against thee. For I know thy rebellion, and thy stiff neck: behold, while I am yet alive with you this day, ye have been rebellious against the Lord; and how much more after my death?

Gather unto me all the elders of your tribes, and your officers, that I may speak these words in their ears, and call heaven and earth to record against them. For I know that after my death ye will utterly corrupt yourselves, and turn aside from the way which I have commanded you; and evil will befall you in the latter days; because ye will do evil in the sight of the Lord, to provoke him to anger through the work of your hands.

And Moses spake in the ears of all the congregation of Israel the words of this song, until they were ended.

\tract{Give ear, O ye heavens, and I will speak: and hear, O earth, the words of my mouth. ℣. My doctrine shall drop as the rain: my speech shall distil as the dew. ℣. As the small rain upon the tender herb, and as the showers upon the grass: because I will publish the name of the Lord: ℣. Ascribe ye greatness unto our God: he is the rock, his work is perfect, for all his ways are judgment. ℣. A God of truth and without iniquity: just and right is the Lord.}

\textsc{Priest.} Let us pray.

\textsc{Deacon.} Let us bow the knee.

\textsc{Subdeacon.} Arise.

\lett{O}{God,} the exaltation of the humble and the strength of them that do stand, who by thy holy servant Moses didst vouchsafe so to instruct thy people in the chanting of thy sacred song, that the repeating of the law by them might likewise be for our instruction: stir up thy power among all the fulness of the nations whom thou hast justified, and grant us gladness, assuaging fear; that the sins of all may be blotted out by thy forgiveness, and that which was denounced in vengeance may turn to our salvation. Through.

\HWreadingcitation{Prophecy XII}{Daniel 3:1}
\lett{I}{n those days:} Nebuchadnezzar the king made an image of gold, whose height was threescore cubits, and the breadth thereof six cubits: he set it up in the plain of Dura, in the province of Babylon. Then Nebuchadnezzar the king sent to gather together the princes, the governors, and the captains, the judges, the treasurers, the counsellors, the sheriffs, and all the rulers of the provinces, to come to the dedication of the image which Nebuchadnezzar the king had set up. Then the princes, the governors, and captains, the judges, the treasurers, the counsellors, the sheriffs, and all the rulers of the provinces, were gathered together unto the dedication of the image that Nebuchadnezzar the king had set up; and they stood before the image that Nebuchadnezzar had set up. Then an herald cried aloud, To you it is commanded, O people, nations, and languages, that at what time ye hear the sound of the cornet, flute, harp, sackbut, psaltery, dulcimer, and all kinds of musick, ye fall down and worship the golden image that Nebuchadnezzar the king hath set up: and whoso falleth not down and worshippeth shall the same hour be cast into the midst of a burning fiery furnace. Therefore at that time, when all the people heard the sound of the cornet, flute, harp, sackbut, psaltery, and all kinds of musick, all the people, the nations, and the languages, fell down and worshipped the golden image that Nebuchadnezzar the king had set up.

Wherefore at that time certain Chaldeans came near, and accused the Jews. They spake and said to the king Nebuchadnezzar, O king, live for ever. Thou, O king, hast made a decree, that every man that shall hear the sound of the cornet, flute, harp, sackbut, psaltery, and dulcimer, and all kinds of musick, shall fall down and worship the golden image: and whoso falleth not down and worshippeth, that he should be cast into the midst of a burning fiery furnace. There are certain Jews whom thou hast set over the affairs of the province of Babylon, Shadrach, Meshach, and Abed-nego; these men, O king, have not regarded thee: they serve not thy gods, nor worship the golden image which thou hast set up.

Then Nebuchadnezzar in his rage and fury commanded to bring Shadrach, Meshach, and Abed-nego. Then they brought these men before the king. Nebuchadnezzar spake and said unto them, Is it true, O Shadrach, Meshach, and Abed-nego, do not ye serve my gods, nor worship the golden image which I have set up? Now if ye be ready that at what time ye hear the sound of the cornet, flute, harp, sackbut, psaltery, and dulcimer, and all kinds of musick, ye fall down and worship the image which I have made; well: but if ye worship not, ye shall be cast the same hour into the midst of a burning fiery furnace; and who is that God that shall deliver you out of my hands? Shadrach, Meshach, and Abed-nego, answered and said to the king, O Nebuchadnezzar, we are not careful to answer thee in this matter. If it be so, our God whom we serve is able to deliver us from the burning fiery furnace, and he will deliver us out of thine hand, O king. But if not, be it known unto thee, O king, that we will not serve thy gods, nor worship the golden image which thou hast set up.

Then was Nebuchadnezzar full of fury, and the form of his visage was changed against Shadrach, Meshach, and Abed-nego: therefore he spake, and commanded that they should heat the furnace one seven times more than it was wont to be heated. And he commanded the most mighty men that were in his army to bind Shadrach, Meshach, and Abed-nego, and to cast them into the burning fiery furnace. Then these men were bound in their coats, their hosen, and their hats, and their other garments, and were cast into the midst of the burning fiery furnace. Therefore because the king’s commandment was urgent, and the furnace exceeding hot, the flame of the fire slew those men that took up Shadrach, Meshach, and Abed-nego. And these three men, Shadrach, Meshach, and Abed-nego, fell down bound into the midst of the burning fiery furnace. And they walked in the midst of the fire, praising God, and blessing the Lord.

\letuspray
\lett{A}{lmighty} and everlasting God, the only hope of the world, who, by the preaching of thy Prophets hast declared the mysteries of these present times: mercifully increase the devotion of thy people; forasmuch as without thine inspiration thy faithful can in no wise advance in virtue. Through.

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent

\section*{Blessing of the Baptismal Font}\label{EasterFontBlessing}
\fancyhead[RO,LE]{\textit{Baptismal Font}}

\begin{rubric}
	These things being finished, if the church has a baptismal Font, the Priest who is to bless the Font puts on a violet Cope, and preceded by the Cross with lights and the blessed Candle lighted, descends with the Clergy and Ministers, vested, to the Font: and meanwhile the following is sung:
\end{rubric}

\tract{Like as the hart desireth the water-brooks: so longeth my soul after thee, O God. ℣. My soul is athirst for God, yea, even for the living God: when shall I come to appear before the presence of God? ℣. My tears have been my meat day and night, while they daily say unto me: Where is now thy God?}

%MANUAL ADJUSTMENT:
\clearpage
\begin{rubric}
	Then the Priest, before he enters for the blessing of the Font, says this Prayer near by the Font:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{A}{lmighty} and everlasting God, mercifully look upon the devotion of this people called to a new birth, who, like the hart, seek the fountain of thy waters: and mercifully grant; that the thirst of their faith may, by the mystery of Baptism, sanctify them in body and soul. Through.

\begin{rubric}
	Then he proceeds to the blessing of the Font, saying:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{A}{lmighty} and everlasting God, be present at the mysteries, be present at the sacraments of thy great goodness: and send forth the spirit of adoption for the regenerating of the new peoples whom the font of baptism doth bring forth unto thee; that what is to be done by our humble ministry, may be effectually fulfilled by thy power. Through Jesus Christ, thy Son, our Lord: Who liveth and reigneth with thee in the unity of the Holy Ghost, God.

\begin{rubric}
	Then raising his voice to the tone of a Preface, he proceeds with joined hands:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/PostProphecySursumCorda.gabc}

%Throughout all ages, world without end.
%
%℟. Amen.
%
%℣. The Lord be with you.
%
%℟. And with thy spirit.
%
%℣. Lift up your hearts.
%
%℟. We lift them up unto the Lord.
%
%℣. Let us give thanks unto our Lord God.
%
%℟. It is meet and right so to do.
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/BaptismPreface.gabc}
%It is very meet, right, and our bounden duty, that we should at all times and in all places give thanks unto thee: O Lord holy Father almighty, everlasting God: Who by thy invisible power dost wondrously give effect to the work of thy sacraments: And though we be unworthy to perform mysteries so great: Yet thou dost not leave us destitute of the gifts of thy grace, but mercifully inclinest thine ears even unto these our supplications. O God, whose Spirit, in the first beginnings of the world, moved over the waters: that even then the nature of water might conceive the virtue of sanctification. O God, who didst wash away in the waters the iniquities of a sinful world, and didst even in the outpouring of the flood shew forth in a figure our regeneration: that by the mystery of this same element, there should be an end to sin and likewise a beginning of virtue. Look, O Lord, upon the face of thy Church, and multiply in her the power of thy regeneration, for thou dost by the streams of thine abundant grace make glad thy city : and openest the fount of baptism throughout the whole world for the renewing of the nations: that, by the command of thy majesty, she may receive from the Holy Ghost the grace of thine only begotten Son.

\begin{rubric}
	Here the Priest divides the water in the form of a cross with his extended hand, which wipes at once with a linen cloth, saying:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/SecretMingling.gabc}
%May he, by the secret mingling of his divine power, make fruitful this water prepared for the regeneration of mankind: that having received sanctification, and being born again a new creature, there may come forth an offspring of heaven from the spotless womb of this divine fountain: And that all, whether sex divide them in body, or age in time, may alike be brought forth into one childhood by grace, their mother. Far hence, O Lord, at thy command let every unclean spirit depart: far hence be all the wickedness of the craft of the devil. Let no power of the enemy here have place or portion: let it not ensnare and compass us around: let it not waylay us in secret: let it not taint us with corruption.

\begin{rubric}
	He touches the water with his hand.
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/Undefiled.gabc}
%May this holy and undefiled creature be free from every assault of the enemy, and purified by the departure of all iniquity. May it be a living fountain, water that doth regenerate, a purifying stream: that all who shall be washed in this laver of salvation may, by the operation of the Holy Ghost within them, obtain the grace of perfect purification.

\begin{rubric}
	He makes three crosses over the Font, saying:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/WaterBlessing.gabc}
%Wherefore I bless thee, O creature of water, by the living {\ding{64}} God, by the true {\ding{64}} God, by the holy {\ding{64}} God: by God who in the beginning through his word divided thee from the dry land: whose spirit moved upon thee.

\begin{rubric}
	Here he divides the water with his hand, and scatters it towards the four quarters of the world, saying:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/DividingWaters.gabc}
%Who made thee to flow from the fountain of paradise, and commanded thee to water the whole earth with thy four rivers. Who in the desert bestowed upon thee sweetness when thou wast bitter, that men might drink, and brought thee forth from the rock for the thirsting people. I ble {\ding{64}} ss thee also by Jesus Christ his only Son, our Lord: who in Cana of Galilee by a wondrous miracle did change thee through his power into wine. Who walked upon thee with his feet: and was baptized in thee by John in Jordan. Who brought thee forth together with blood from his side: and commanded his disciples, that believing they should be baptized in thee, saying: Go ye, teach all nations, baptizing them in the name of the Father, and of the Son, and of the Holy Ghost.

\begin{rubric}
	He changes his voice, and proceeds in the tone of the Lesson:
\end{rubric}
\par\noindent
Do thou, Almighty God, mercifully assist us, who keep these thy commandments: do thou graciously breathe upon us.

\begin{rubric}
	He breathes thrice upon the water in the form of a cross, saying:
\end{rubric}
\par\noindent
Do thou with thy mouth bless this element of water: that as by nature it hath power to cleanse and wash the body, so also it may be effectual for the purifying of the soul.

\begin{rubric}
	Here the Priest lowers the Candle into the water a little way: and resuming the tone of the Preface, says:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/Fulness.gabc}
%May the power of the Holy Ghost descend upon the fulness of this font.

\begin{rubric}
	Then withdrawing the Candle from the water, he again dips it more deeply, and repeats in a somewhat higher voice, \blackrubric{May the power.} Then he again withdraws the Candle from the water, and the third time dips it to the bottom, repeating in a still higher voice: \blackrubric{May the power,} as above.
	
	Then breathing three times on the water in this form \textsc{Ψ} he proceeds:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/WholeSubstance.gabc}
%And make the whole substance of this water to be fruitful unto regeneration.

%MANUAL ADJUSTMENT:
\clearpage
\begin{rubric}
	Here the Candle is taken out of the water, and he proceeds:
\end{rubric}
\par\noindent
\gregorioscore{resources/gabc/Propers/EasterEven/TrueInnocence.gabc}
%Here may the stains of every sin be blotted out: here may nature, created in thine image, be restored to the honour of its first estate, and cleansed from all the defilement of its old condition: that every man who cometh to this sacrament of regeneration may be born again unto the new childhood of true innocence.

\begin{rubric}
	He says that which follows in the reading voice:
\end{rubric}
\par\noindent
Through Jesus Christ thy Son, our Lord: Who shall come to judge the quick and the dead, and the world by fire. \textit{Amen.}

\begin{rubric}
	Then some of the water that has been blessed is sprinkled by the assisting Priests over the People.
	
	And meantime one of the ministers of the church takes some of the water in a vessel, for sprinkling in houses and other places. Which done, the Priest who is blessing the Font, pours some of the Oil of Catechumens into the water in the form of a cross, saying in an audible voice:
\end{rubric}
\par\noindent
May this font be sanctified and made fruitful by the Oil of salvation, for such as shall be born again therefrom, unto life everlasting. \textit{Amen.}

\begin{rubric}
	Then, in the same manner, he pours in some of the Chrism, saying:
\end{rubric}
\par\noindent
May this inpouring of the Chrism of our Lord Jesus Christ, and of the Holy Ghost the Paraclete, be wrought in the name of the holy Trinity. \textit{Amen.}

\begin{rubric}
	Lastly he takes both phials of the said holy Oil and the Chrism and pours in some of both together in the form of a cross, saying:
\end{rubric}
\par\noindent
May this commingling of the Chrism of sanctification, and of the Oil of unction and of the Water of baptism be likewise wrought in the name of the Fa {\ding{64}} ther, and of the S {\ding{64}} on, and of the Holy {\ding{64}} Ghost. \textit{Amen.}

\begin{rubric}
	Then he mingles the Oil itself with the water, and with his hand scatters it over the whole Font. If there be any to be baptized, he baptizes them in the usual manner.
\end{rubric}

\begin{rubric}
	Then as the Priest and Ministers return to the Altar, the Litany is sung by two Cantors, and the Choir repeat it also together, as is directed below.
\end{rubric}

    \par\noindent
	\centerline{\rule{0.25\textwidth}{0.4pt}}
	\par\noindent

\section*{Paschal Litany}\label{PaschalLitany}
\fancyhead[RO,LE]{\textit{Paschal Litany}}

\begin{rubric}
	But where there is no baptismal Font, when the last Prophecy with its Collect is ended, the Celebrant lays aside his Chasuble, and with the Ministers lies prostrate before the Altar: and all kneeling, the Litany is chanted by two Cantors in the midst of the Choir, and both sides of the Choir respond together.
	
	But when they come to the \blackrubric{℣. We sinners,} the Priest and his Ministers rise, and proceed to the sacristy, where they put on white vestments for the solemn Mass, and meanwhile the candles on the Altar are lighted.
	
	At the end of the Litany, \blackrubric{Kyrie eleison} is sung solemnly, and repeated, as is usual at Mass.
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/PaschalKyrie.gabc}

\gregorioscore{resources/gabc/Propers/EasterEven/DivineLaudations.gabc}

\gregorioscore{resources/gabc/Propers/EasterEven/MarianPetition.gabc}

%\elcoldent{Lord, have mercy upon us.}{Kyrie, eléison.}
%\elcol{℟. Christ, have mercy upon us.}{℟. Christe, eléison.}
%\elcoldent{Lord, have mercy upon us.}{Kyrie, eléison.}
%\elcoldent{O Christ hear us.}{Christe, audi nos.}
%\elcol{℟. O Christ graciously hear us.}{℟. Christe, exáudi nos.}
%\elcoldent{O God the Father of heaven,}{Pater de c{\ae}lis, Deus,}
%\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
%\elcoldent{O God the Son, Redeemer of the world,}{Fili, Redémptor mundi, Deus,}
%\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
%\elcoldent{O God the Holy Ghost,}{Spíritus Sancte, Deus,}
%\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
%\elcoldent{Holy Trinity one God,}{Sancta Trínitas, unus Deus,}
%\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
%\elcoldent{Holy Mary,}{Sancta María,}
%\elcol{℟. Pray for us.}{℟. Ora pro nobis.}

\elcoldent{Holy Mother of God,}{Sancta Dei Génitrix,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Virgin of virgins,}{Sancta Virgo vírginum,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Michael,}{Sancte Micha{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Gabriel,}{Sancte Gábri{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy Raphael,}{Sancte Rápha{\"e}l,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Angels and Archangels,}{Omnes sancti Angeli et Archángeli,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy orders of blessed Spirits,}{Omnes sancti beatórum Spírituum órdines,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. John Baptist,}{Sancte Joánnes Baptísta,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Joseph,}{Sancte Josephe,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Patriarchs and Prophets,}{Omnes sancti Patriárch{\ae} et Prophét{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Peter,}{Sancte Petre,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Paul,}{Sancte Paule,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Andrew,}{Sancte Andréa,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. John,}{Sancte Joánnes,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Apostles and Evangelists,}{Omnes sancti Apóstoli et Evangelíst{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Disciples of the Lord,}{Omnes sancti Discípuli Dómini,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Stephen,}{Sancte Stéphane,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Lawrence,}{Sancte Laurénti,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Vincent,}{Sancte Vincénti,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Martyrs,}{Omnes sancti Mártyres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Sylvester,}{Sancte Silvéster,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Gregory,}{Sancte Gregóri,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Augustine,}{Sancte Augustíne,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Bishops and Confessors,}{Omnes sancti Pontífices et Confessóres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Doctors,}{Omnes sancti Doctóres,}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{Holy St. Anthony,}{Sancte Antóni,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{Holy St. Benedict,}{Sancte Benedícte,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Priests and Levites,}{Omnes sancti Sacerdótes et Levít{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye holy Monks and Hermits,}{Omnes sancti Mónachi et Eremít{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{St. Mary Magdalene,}{Sancta María Magdaléna,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Agnes,}{Sancta Agnes,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Cecilia,}{Sancta C{\ae}cília,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Agatha,}{Sancta Agatha,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{St. Anastasia,}{Sancta Anastásia,}
\elcol{℟. Pray for us.}{℟. Ora pro nobis.}
\elcoldent{All ye holy Virgins and Widows,}{Omnes sanct{\ae} Vírgines et Vídu{\ae},}
\elcol{℟. Pray for us.}{℟. Oráte pro nobis.}
\elcoldent{All ye Saints of God,}{Omnes Sancti et Sanct{\ae} Dei,}
\elcol{℟. Intercede for us.}{℟. Intercédite pro nobis.}

\gregorioscore{resources/gabc/Propers/EasterEven/BeThouMerciful.gabc}

%\elcoldent{Be thou merciful,}{Propítius esto,}
%\elcol{℟. Spare us, O Lord.}{℟. Parce nobis, Dómine.}
\elcoldent{Be thou merciful,}{Propítius esto,}
\elcol{℟. Hear us, O Lord.}{℟. Exáudi nos, Dómine.}
\elcoldent{From all evil,}{Ab omni malo,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{From all sin,}{Ab omni peccáto,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{From everlasting death,}{A morte perpétua,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By the mystery of thy holy Incarnation,}{Per mystérium sanct{\ae} Incarnatiónis tu{\ae},}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thine advent,}{Per Advéntum tuum,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy nativity,}{Per Nativitátem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy baptism and holy fasting,}{Per Baptísmum et sanctum Jejúnium tuum,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy cross and passion,}{Per Crucem et Passiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy death and burial,}{Per Mortem et sepultúram tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy holy resurrection,}{Per sanctam Resurrectiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By thy wonderful ascension,}{Per admirábilem Ascensiónem tuam,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{By the coming of the Holy Ghost the Paraclete,}{Per advéntum Spíritus Sancti Parácliti,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}
\elcoldent{In the day of judgment,}{In die judícii,}
\elcol{℟. Deliver us, O Lord.}{℟. Líbera nos, Dómine.}

\gregorioscore{resources/gabc/Propers/EasterEven/WeSinners.gabc}

\gregorioscore{resources/gabc/Propers/EasterEven/WeBeseechThee.gabc}

%\elcoldent{We sinners,}{Peccatóres,}
%\elcol{℟. Beseech thee, hear us.}{℟. Te rogámus, audi nos.}
%\elcoldent{That thou wouldest spare us,}{Ut nobis parcas,}
%\elcol{℟. We beseech thee, hear us.}{℟. Te rogámus, audi nos.}

\elcoldent{That thou wouldest vouchsafe to govern and preserve thy holy Church,}{Ut Ecclésiam tuam sanctam régere et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to preserve our Apostolic lord and all orders of the Church in holy religion,}{Ut Domnum Apostólicum, et omnes ecclesiásticos órdines in sancta religióne conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to humble the enemies of holy Church,}{Ut inimícos sanct{\ae} Ecclési{\ae} humiliáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to give unto Christian kings and rulers peace and true concord,}{Ut régibus et princípibus christiánis pacem et veram concórdiam donáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to strengthen, and preserve us in thy holy service,}{Ut nosmetípsos in tuo sancto servítio confortáre et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to reward all our benefactors with everlasting blessings,}{Ut ómnibus benefactóribus nostris sempitérna bona retríbuas,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to give and preserve the fruits of the earth,}{Ut fructus terr{\ae} dare et conserváre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe to grant unto all the faithful departed rest eternal,}{Ut ómnibus fidélibus defúnctis réquiem {\ae}térnam donáre dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}
\elcoldent{That thou wouldest vouchsafe graciously to hear us,}{Ut nos exáudire dignéris,}
\elcol{℟. We beseech thee, hear us.}{℟. Te rogamus, audi nos.}

\gregorioscore{resources/gabc/Propers/EasterEven/Agnus.gabc}

%\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
%\elcol{℟. Spare us, O Lord, hear us.}{℟. Parce nobis, Dómine.}
%\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
%\elcol{℟. Graciously hear us, O Lord.}{℟. Exáudi nos, Dómine.}
%\elcoldent{O Lamb of God, who takest away the sins of the world,}{Agnus Dei, qui tollis peccáta mundi,}
%\elcol{℟. Have mercy upon us.}{℟. Miserére nobis.}
%\elcoldent{O Christ, hear us.}{Christe, audi nos.}
%\elcol{℟. O Christ, graciously hear us.}{℟. Christe, exáudi nos.}


\clearpage
\section*{Holy Mass}
\fancyhead[RO,LE]{\textit{Holy Mass}}

\begin{rubric}
	Here the Cantors solemnly intone \blackrubric{Kyrie, eléison,} \blackrubric{Christe, eléison,} \blackrubric{Kyrie, eléison,} repeating each invocation thrice.
	
	Meanwhile the Priest with the Ministers in white vestments approaches the Altar: and having said the Psalm \blackrubric{Give sentence with me, O God,} with \blackrubric{Glory be,} he makes the Confession: then going up, he kisses the Altar, and censes it in the usual manner: and after the Choir has finished \blackrubric{Kyrie, eléison,} he solemnly intones \blackrubric{Gloria in excelsis,} and the bells are rung. Then the Priest says:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

\HWcollect
\lett{O}{God,} who dost enlighten this most sacred night with the glory of the Resurrection of the Lord: preserve in the new offspring of thy family the spirit of adoption, which thou hast given; that they being renewed in body and soul may render thee pure service. Through the same.

\HWreadingcitation{Epistle}{Colossians 3:1}
\lett{B}{rethren:} If ye be risen with Christ, seek those things which are above, where Christ sitteth on the right hand of God. Set your affection on things above, not on things on the earth. For ye are dead, and your life is hid with Christ in God. When Christ, who is our life, shall appear, then shall ye also appear with him in glory.

\begin{rubric}
	The Epistle ended, the Celebrant begins: \blackrubric{Alleluia.}
	
	And be sings it through thrice, raising his voice each time: the Choir repeats it each time in the same tone.
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/PaschalAlleluia.gabc}

\begin{rubric}
	Afterwards the Choir continues:
\end{rubric}
\par\noindent
℣. O give thanks unto the Lord, for he is gracious: because his mercy endureth for ever.

\tract{O praise the Lord, all ye heathen: praise him, all ye nations. ℣. For his merciful kindness is ever more and more towards us: and the truth of the Lord endureth for ever.}

\begin{rubric}
	At the Gospel, lights are not carried, but only incense. A blessing is asked, and other things are done as usual.
\end{rubric}

\HWreadingcitation{Gospel}{Matthew 28:1}

\lett{I}{n} the end of the sabbath, as it began to dawn toward the first day of the week, came Mary Magdalene and the other Mary to see the sepulchre. And, behold, there was a great earthquake: for the angel of the Lord descended from heaven, and came and rolled back the stone from the door, and sat upon it. His countenance was like lightning, and his raiment white as snow: and for fear of him the keepers did shake, and became as dead men. And the angel answered and said unto the women, Fear not ye: for I know that ye seek Jesus, which was crucified. He is not here: for he is risen, as he said. Come, see the place where the Lord lay. And go quickly, and tell his disciples that he is risen from the dead; and, behold, he goeth before you into Galilee; there shall ye see him: lo, I have told you.

\begin{rubric}
	The Creed is not said, but, the Gospel ended, the Priest says: \blackrubric{The Lord be with you,} then: \blackrubric{Let us pray.} The Offertory Verse is not said. At the Lavabo \blackrubric{Gloria Patri} is said.
\end{rubric}

\secret
\lett{R}{eceive,} we beseech thee O Lord, the prayers of thy people together with the offering of these sacrifices: that those things which we have begun in these paschal mysteries may by thine operation be profitable for our healing in eternity. Through.

\begin{rubric}
    	\blackrubric{The Peace of the Lord} is said, but the Kiss of Peace is not given. \blackrubric{Agnus Dei} is not said, nor the Postcommunion, but the accustomed Prayers before the Communion are said.
\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
\section*{Vespers}

\begin{secrubric}
	After the receiving of the Sacraments, Vespers is sung in choir.
\end{secrubric}


\subsection{Psalm 117. \textit{Laudate Dominum}}

\antiphon{}{Alleluia, alleluia, alleluia.}

\elcol{\lett{O}{praise} the \textsc{Lord}, all ye heathen : praise him, all ye nations.}{\lett{L}{aud\smash{á}te} Dóminum, omnes gentes: * laudáte eum, omnes pópuli:}
\elcol{For his merciful kindness is ever more and more toward us : and the truth of the \textsc{Lord} endureth for ever. Praise the \textsc{Lord}.}{Quóniam confirmáta est super nos misericórdia ejus: * et véritas Dómini manet in {\ae}térnum.}
\elcol{Glory be to the Father, and to the Son, *
    and to the Holy Ghost:}{Glória Patri, et Fílio, * et Spirítui Sancto.}
\elcol{As it was in the beginning, is now, and ever shall be, *
    world without end. Amen.}{Sicut erat in princípio, et nunc, et semper, * et in sǽcula s{\ae}culórum. Amen.}
    
\antiphon{}{Alleluia, alleluia, alleluia.}

\begin{rubric}
	The Chapter, Hymn, and Verse are not said, but the Celebrant begins at once to sing the Antiphon to the Magnificat:
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/PaschalAntiphon.gabc}

%\lett{I}{n the end of the Sabbath.}

\begin{rubric}
	And the Choir continues:
\end{rubric}

As it began to dawn toward the first day of the week: came Mary Magdalene and the other Mary to see the sepulchre, alleluia.

\begin{rubric}
	Then the Magnificat is sung with \blackrubric{Gloria Patri} at the end, and the censing is done, as at other times at Evensong.
\end{rubric}

\begin{rubric}
	The Antiphon having been repeated, the Celebrant says:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

%MANUAL ADJUSTMENT:
\clearpage
%\letuspray
\lett{P}{our} forth upon us, O Lord, the Spirit of thy charity: that as thou hast fulfilled us with these paschal sacraments, so of thy goodness thou wouldest make us to be of one heart and mind. Through ... in the unity of the same.

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\begin{rubric}
	And the Deacon, turning himself to the people, sings:
\end{rubric}

\gregorioscore{resources/gabc/Propers/EasterEven/IteMissaEst.gabc}

    %\elcol{℣. Depart in peace, alleluia, alleluia.}{℣. Ite, Missa est, allelúja, allelúja.}
    \elcol{℟. Thanks be to God, alleluia, alleluia.}{℟. Deo grátias, allelúja, allelúja.}

\begin{rubric}
	And the Priest, having said \blackrubric{Let this my bounden duty,} gives the Blessing in the usual manner.
\end{rubric}