\section{Holy Week}

\subsection{Palm Sunday}\label{PalmSunday}
\feastday{Palm Sunday}

\subsubsection{Blessing of Palms}
\begin{rubric}
In Choir, Mattins (or, Terce) having been said, the sprinkling of water is performed in the usual manner. Then the Priest, vested in a violet Cope, or without the Chasuble, with the Ministers vested in like manner, proceeds to bless branches of palm and olive or other trees placed in the midst before the Altar, or at the Epistle corner. And first the Choir sing the Antiphon:
\end{rubric}
\antiphon{}{Hosanna to the Son of David: blessed is he that cometh in the name of the Lord. O King of Israel: Hosanna in the Highest.}
\begin{rubric}
    Then the Priest standing at the Epistle corner, without turning toward the people, sings with joined hands in the tone of the Ferial Collect:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\letuspray
\lett{O}{God,} whose love and service is our justice, multiply upon us the gifts of thy unspeakable grace: and like as by the death of thy Son thou hast given unto us the hope of our faith; so grant that by his resurrection we may be brought unto the country which we seek. Who liveth and reigneth with thee.

\begin{rubric}
    Then the Subdeacon in the accustomed place sings the following Lesson in the tone of the Epistle, and at the end kisses the hand of the Priest.
\end{rubric}

\readingcitation{Lesson}{Exodus 15:27}
\lett{I}{n those days:} The children of Israel came to Elim, where were twelve wells of water, and threescore and ten palm trees: and they encamped there by the waters. And they took their journey from Elim, and all the congregation of the children of Israel came unto the wilderness of Sin, which is between Elim and Sinai, on the fifteenth day of the second month after their departing out of the land of Egypt. And the whole congregation of the children of Israel murmured against Moses and Aaron in the wilderness: And the children of Israel said unto them, Would to God we had died by the hand of the \divineName{Lord} in the land of Egypt, when we sat by the flesh pots, and when we did eat bread to the full; for ye have brought us forth into this wilderness, to kill this whole assembly with hunger. Then said the \divineName{Lord} unto Moses, Behold, I will rain bread from heaven for you; and the people shall go out and gather a certain rate every day, that I may prove them, whether they will walk in my law, or no. And it shall come to pass, that on the sixth day they shall prepare that which they bring in; and it shall be twice as much as they gather daily. And Moses and Aaron said unto all the children of Israel, At even, then ye shall know that the \divineName{Lord} hath brought you out from the land of Egypt: And in the morning, then ye shall see the glory of the \divineName{Lord}.

\gradual{The chief priests and the Pharisees gathered a council, and said: What do we, for this man doeth many miracles? If we let him thus alone, all men will believe on him: * And the Romans shall come, and take away both our place and nation. ℟. And one of them named Caiaphas, being the high priest that same year, prophesied saying: It is expedient for us, that one man should die for the people, and that the whole nation perish not. Then from that day forth they took counsel together to put him to death, saying. * And the Romans shall come, and take away both our place and nation. ℟. On the Mount of Olives he prayed to his Father; Father, if it be possible, let this cup pass from me. * The spirit indeed is willing, but the flesh is weak; thy will be done. ℣. Watch and pray, that ye enter not into temptation. * The spirit indeed is willing, but the flesh is weak; thy will be done.}

\begin{rubric}
    During the singing of the \blackrubric{℟.,} the Deacon places the Book of the Gospels on the Altar: and the Priest, the Deacon ministering the incense-boat, places incense in the censer. Then the Deacon says: \blackrubric{Cleanse my heart,} takes the Book from the Altar, and asks a blessing from the Priest: then, while the Subdeacon holds the Book, he stands between two Acolytes holding lighted candles, signs the book and censes it, and sings the Gospel, as below, in the usual way: after which the Subdeacon carries the Book to be kissed by the Priest, who is censed by the Deacon.
\end{rubric}

\readingcitation{Gospel}{Matthew 21:1}
\lett{A}{t that time:} When Jesus drew nigh unto Jerusalem, and was come to Bethphage, unto the mount of Olives, then sent he two disciples, Saying unto them, Go into the village over against you, and straightway ye shall find an ass tied, and a colt with her: loose them, and bring them unto me. And if any man say ought unto you, ye shall say, The Lord hath need of them; and straightway he will send them. All this was done, that it might be fulfilled which was spoken by the prophet, saying, Tell ye the daughter of Sion, Behold, thy King cometh unto thee, meek, and sitting upon an ass, and a colt the foal of an ass. And the disciples went, and did as Jesus commanded them, And brought the ass, and the colt, and put on them their clothes, and they set him thereon. And a very great multitude spread their garments in the way; others cut down branches from the trees, and strawed them in the way. And the multitudes that went before, and that followed, cried, saying, Hosanna to the Son of David: Blessed is he that cometh in the name of the Lord.

\begin{rubric}
    After this the branches are blessed. The Priest, standing at the same Epistle corner, says in the tone of the Ferial Collect:
\end{rubric}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{I}{ncrease,} O God, the faith of them that put their trust in thee, and graciously hear the prayers of thy humble servants; let thy manifold mercies descend upon us: and let these branches of palm or olive be ble {\ding{64}} ssed : and, as in a figure of the Church thou didst multiply Noah going forth from the ark, and Moses when he went out of Egypt with the children of Israel: so may we, bearing palms and branches of olive, go forth with good works to meet Christ: and through him enter into everlasting joy: Who liveth and reigneth with thee in the unity of the Holy Ghost, God: throughout all ages, world without end. \textit{Amen.}\\

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Lift up your hearts.}{℣. Sursum corda.}
\elcol{℟. We lift them up unto the Lord.}{℟. Habémus ad Dóminum.}
\elcol{℣. Let us give thanks unto our Lord God.}{℣. Grátias agámus Dómino, Deo nostro.}
\elcol{℟. It is meet and right so to do.}{℟. Dignum et justum est.}

\vspace{1\baselineskip}
\par\noindent
It is very meet, right, and our bounden duty, that we should at all times, and in all places, give thanks unto thee: O Lord holy Father almighty, everlasting God: Who art glorified in the council of thy Saints. For thee thy creatures serve: and acknowledge thee their only maker and their God, and thee all thy works do praise, and thee thy Saints do bless. For with fearless voice, before the kings and rulers of this world, they confess the great name of thine only begotten Son. Before whom Angels and Archangels, Thrones and Dominations stand: and with all the host of the heavenly army, do sing the hymn of thy glory, evermore saying:

\begin{rubric}
    And the \blackrubric{Sanctus} is sung by the Choir,
\end{rubric}

\elcol{\lett{H}{oly, Holy, Holy,} Lord God of hosts, Heaven and earth are full of thy glory: Glory be to thee, O Lord Most High. Blessed {\ding{64}} is he that cometh in the Name of the Lord. Hosanna in the Highest.}
{\lett{S}{anctus, Sanctus, Sanctus,} Dóminus, Deus Sábaoth. Pleni sunt c{\ae}li et terra glória tua. Glória tibi, Dómine altíssime. Benedíctus, {\ding{64}} qui venit in nómine Dómini. Hosánna in excélsis.}

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

\vspace{1\baselineskip}
\par\noindent
\letuspray
\lett{W}{e} beseech thee, O Lord holy Father almighty, everlasting God: that thou wouldest vouchsafe to bl~{\ding{64}}~ess and sanc~{\ding{64}}~tify this creature of olive, which thou hast bidden to spring from the wood of the tree, which likewise the dove, returning to the ark, did bear in her mouth: that all who shall receive thereof, may obtain for themselves protection both in body and soul: and let it be, O Lord, a remedy for our salvation, a sacrament
of thy grace. Through.
\letuspray
\lett{O}{God,} who dost gather together the things which are dispersed abroad, and dost preserve that which thou dost gather: who didst bless the people when they went forth bearing branches to meet Jesus: bl~{\ding{64}}~ess also these branches of palm and olive which thy servants receive with faith to the honour of thy name; that into whatsoever place they shall be brought, the dwellers in that place may obtain thy blessing: that all adversity being put to flight, thy right hand may protect those who are redeemed by Jesus Christ thy Son our Lord. Who liveth and reigneth with thee.
\letuspray
\lett{O}{God,} who by a wonderful order of thy providence wast pleased, even by means of things insensible, to shew forth the dispensation of our salvation: grant, we beseech thee; that the devout hearts of thy faithful people may profitably understand what is mystically signified by that deed when, on this day, the multitude, inspired by the heavenly light, went forth to meet the Redeemer, and strawed branches of palm and olive beneath his feet. The branches, then, of palm foreshadow his triumph over the prince of death; and the boughs of olive in a manner proclaim that the anointing of the Spirit is come. For the multitude rejoiced to know that even then it was prefigured: that our Redeemer, having compassion on the misery of mankind, was about to battle with the prince of death, and by his death to triumph. And therefore obediently they laid before him such things as should signify in him both the triumphs of his victory and the abundance of his mercy. Wherefore we likewise with sure faith, having in memory both the deed and the signification of the same, humbly beseech thee, O Lord holy Father almighty, everlasting God, through the same Jesus Christ our Lord: that in him and through him, whose members thou hast been pleased to make us, we may win the victory over the empire of death, and be found worthy to be partakers of his glorious resurrection. Who liveth and reigneth with thee.
\letuspray
\lett{O}{God,} who by an olive-branch didst command the dove to proclaim peace on earth: vouchsafe, we beseech thee; to sanctify with thy heavenly bless {\ding{64}} ing these branches of olive and other trees; that for all thy people they may be profitable unto salvation. Through Christ, our Lord. \textit{Amen.}
\letuspray
\lett{B}{l~{\ding{64}}~ess,} we beseech thee, O Lord, these branches of palm or olive: and grant; that as thy people on this day do outwardly perform these things to thy honour, so they, spiritually fulfilling the same with pure devotion, may win the victory over the enemy, and cleave steadfastly to every work of mercy. Through.
\begin{rubric}
    Then the Celebrant, after putting incense into the thurible, thrice sprinkles the boughs with holy water, saying the Antiphon, \blackrubric{Thou shalt purge me,} without note and without the Psalm: and then thrice incenses them. Then he says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}

%MANUAL ADJUSTMENT:
\vspace{1\baselineskip}

\letuspray
\lett{O}{God,} who didst send forth thy Son, Jesus Christ our Lord, into the world for our salvation, that he might humble himself to be made like unto us, and call us back to thee: before whom, at his coming to Jerusalem for the fulfilling of the Scriptures, the multitude of the people that believed with faithful devotion strawed their raiment and palm branches in the way: grant, we beseech thee; that we may so prepare for him the way of faith, that, every stone of stumbling and rock of offence being done away, our works may blossom before thee with branches of righteousness: and that we may be found worthy to follow in his footsteps. Who liveth and reigneth with thee.
\begin{rubric}
    The blessing being completed, the senior of the Clergy goes up to the Altar and presents a blessed branch to the Celebrant, who does not genuflect, nor kiss the hand of him who gives it. Then the Celebrant, standing before the Altar, turns to the people and distributes the branches, first to the senior from whom he received his, then to the Deacon and Subdeacon vested, and to the other clergy severally in order, last of all to the laity: all genuflecting and kissing the branch and the hand of the Celebrant, except Prelates, if they are present. And when he begins to distribute, the Choir sings the following Antiphons:
\end{rubric}\par
\antiphon{}{The children of the Hebrews bearing branches of olive, went out to meet the Lord, crying out and saying: Hosanna in the highest.}\\

\antiphon{Another}{The children of the Hebrews spread their garments in the way and cried out, saying: Hosanna to the Son of David; blessed is he that cometh in the name of the Lord.}
\begin{rubric}
    If these are not sufficient, they are repeated until the distribution is finished. Then the Priest says:
\end{rubric}
\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\letuspray
\lett{A}{lmighty} and everlasting God, who didst ordain that our Lord Jesus Christ should ride upon the foal of an ass, and didst teach the multitudes of the people to straw in his way their raiment and branches of trees, and likewise to sing Hosanna in his praise: grant, we beseech thee; that we may by thee be enabled to follow their innocency, and be partakers of their merit. Through the same Christ, our Lord. \textit{Amen.}
\begin{rubric}
    Then the Procession is made. And first the Celebrant puts incense in the censer: and the Deacon, turning to the people, says: \blackrubric{Let us proceed in peace,} and the Choir responds: \blackrubric{In the name of Christ. Amen.}
\end{rubric}
\begin{rubric}
    The Thurifer goes first with the smoking censer: then the Subdeacon vested, carrying the Cross, between two Acolytes with lighted candles: the Clergy follow in order, and last of all the Celebrant with the Deacon on his left, all with branches in their hands: and the following Antiphons are sung, either all, or some, so long as the Procession lasts.
\end{rubric}\par
\antiphon{}{When the Lord drew nigh unto Jerusalem, he sent two of his disciples, saying: Go into the village over against you: and ye shall find the colt of an ass tied, whereon never man sat: loose him, and bring him unto me. If any man say ought unto you, ye shall say: The Lord hath need of him. They loose him and bring him to Jesus: and put on him their clothes, and set him thereon: some spread their garments in the way: others strawed branches from the trees: and they that followed, cried: Hosanna, blessed is he that cometh in the name of the Lord: blessed is the kingdom of our father David: Hosanna in the highest: Son of David, have mercy upon us.}\par
\antiphon{Another}{When the people heard that Jesus was coming to Jerusalem, they took branches of palm-trees: and went forth to meet him, and the children cried saying: This is he that should come for the salvation of the people. This is our salvation and the redemption of Israel. How great is he whom Thrones and Dominations go forth to meet! Fear not, daughter of Sion: behold thy King cometh to thee, sitting upon an ass' colt, as it is written. Hail, O King, Creator of the world, who art come to redeem us.}\par
\antiphon{Another}{Six days before the solemnity of the Passover, when the Lord came into the city of Jerusalem, the children went forth to meet him: and in their hands carried branches of palm trees, and cried with a loud voice, saying: Hosanna in the highest: blessed art thou, that art come in the multitude of thy mercies: Hosanna in the highest.}\par
\antiphon{Another}{The multitudes with flowers and palms go forth to meet the Redeemer: and render worthy homage to the triumphant conqueror: the Gentiles with their lips proclaim the Son of God: and in the praise of Christ their voices thunder through the skies: Hosanna in the highest.}\par
\antiphon{Another}{With the Angels and the children may we be found faithful, crying unto the vanquisher of death: Hosanna in the highest.}\par
\antiphon{Another}{A great multitude, that were come together to the feast, cried unto the Lord: Blessed is he that cometh in the name of the Lord: Hosanna in the highest.}
\begin{rubric}
    As the Procession turns back, two or four Cantors enter the church, and the door being shut, with their faces toward the Procession, sing the first verse of the following hymn, \blackrubric{Gloria laus, etc.;} which verse is repeated by the Priest and others outside the church. Then those within sing the other verses which follow; either all or a part, as shall seem good: and those who are without respond to each verse: \blackrubric{Gloria laus,} as at the ℣. at the beginning.
\end{rubric}

\lett{G}{lory} and honour and praise be to thee, our King and Redeemer,\par
\secondline{Christ, to whom children of old loved their Hosannas to raise. ℟. Glory and honour.}
\thirdline{Israel's Monarch art thou, and the glorious Offspring of David,}
Thou that approachest, a King, blest in the Name of the Lord. ℟. Glory and honour.\par
Glory to thee in the highest, the heavenly armies are singing: Glory to thee upon earth, man and creation reply. ℟. Glory and honour.\par
Met thee with palms in their hands that day the folk of the Hebrews,\par
We with our prayers and our hymns now to thy presence approach. ℟. Glory and honour.\par
They to thee proffered their praise for to herald thy dolorous Passion;\par
We to the King on his throne utter the jubilant hymn. ℟. Glory and honour.\par
They were then pleasing to thee, unto thee our devotion be pleasing;\par
Merciful King, kind King, who in all goodness art pleased. ℟. Glory and honour.\par
\begin{inhead}
    or,
\end{inhead}
\lett{A}{ll} glory, laud, and honour To thee, Redeemer, King, To whom the lips of children Made sweet hosannas ring. ℟. All glory, laud.\par
\secondline{Thou art the King of Israel, Thou David's royal Son, Who in the Lord's name comest, The King and blessed One. ℟. All glory, laud.}
\thirdline{The company of Angels Are praising thee on high, And mortal men and all things Created make reply. ℟. All glory, laud.}
The people of the Hebrews With palms before thee went; Our praise and prayer and anthems Before thee we present. ℟. All glory, laud.\par
To thee before thy passion They sang their hymns of praise; To thee now high exalted Our melody we raise. ℟. All glory, laud.\par
Thou didst accept their praises, Accept the prayers we bring, Who in all good delightest, Thou good and gracious King. ℟. All glory, laud.\par
\begin{rubric}
    Then the Subdeacon with the foot of the Cross knocks at the door, which being straightway opened, the Procession enters the church singing:
\end{rubric}\par\noindent
℟. When the Lord entered the holy city, the children of the Hebrews, foretelling the resurrection of life, * With branches of palm cried: Hosanna in the highest. ℣. When the people heard that Jesus was coming to Jerusaiem, they went forth to meet him. With branches of palm cried: Hosanna in the highest.

\begin{rubric}
    Then Mass is celebrated, and branches are held in the hands, while the Passion and Gospel are sung only.
\end{rubric}

\introit
\lett{B}{e} not thou far from me, O Lord, thou art my succour, haste thee to help me: save me from the lion's mouth, thou hast heard me also from among the horns of the unicorns. \textit{Ps.} My God, my God, look upon me, why hast thou forsaken me: and art so far from my health, and from the words of my complaint? 
\collect\label{PalmSundayCollect}
\lett{A}{lmighty} and everlasting God, who of thy tender love towards mankind, hast sent thy Son, our Saviour Jesus Christ, to take upon him our flesh, and to suffer death upon the cross, that all mankind should follow the example of his great humility; Mercifully grant, that we may both follow the example of his patience, and also be made partakers of his resurrection. Through the same.
\begin{rubric}
    This Collect only is said, even if a Commemoration has been made in the Office.
\end{rubric}
\begin{rubric}
    This Collect is to be said every day, after the Collect appointed for the Day, until Good Friday.
\end{rubric}
\readingcitation{Epistle}{Philippians 2:5}
\lett{B}{rethren:} Let this mind be in you, which was also in Christ Jesus: Who, being in the form of God, thought it not robbery to be equal with God: But made himself of no reputation, and took upon him the form of a servant, and was made in the likeness of men: And being found in fashion as a man, he humbled himself, and became obedient unto death, even the death of the cross. Wherefore God also hath highly exalted him, and given him a name which is above every name: \inrub{Genuflect.} That at the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth; And that every tongue should confess that Jesus Christ is Lord, to the glory of God the Father.

\gradtr{Thou hast holden me by my right hand; thou shalt guide me with thy counsel: and after that receive me with glory. ℣. Truly God is loving unto Israel, even unto such as are of a clean heart; nevertheless, my feet were almost gone; my treadings had well nigh slipt: and why? I was grieved at the wicked, I do also see the ungodly in such prosperity.}{My God, my God, look upon me: why hast thou forsaken me? ℣. And art so far from my health, and from the words of my complaint? ℣. O my God, I cry in the daytime, but thou hearest not: and in the night season also I take no rest. ℣. And thou continuest holy, O thou worship of Israel. ℣. Our fathers hoped in thee: they trusted in thee, and thou didst deliver them. ℣. They called upon thee, and were holpen: they put their trust in thee, and were not confounded. ℣. But as for me, I am a worm, and no man: a very scorn of men, and the outcast of the people. ℣. All they that see me laugh me to scorn: they shoot out their lips, and shake their head, saying: ℣.He trusted in God that he would deliver him: let him deliver him, if he will have him. ℣. They stand staring and looking upon me: they part my garments among them, and cast lots upon my vesture. ℣. Save me from the lion's mouth: thou hast heard me also from among the horns of the unicorns. ℣. O praise the Lord, ye that fear him: magnify him, all ye of the seed of Jacob. ℣. They shall be counted unto the Lord for a generation: they shall come, and the heavens shall declare his righteousness. ℣. Unto a people that shall be born, whom the Lord hath made.}
\begin{rubric}
    The Passion begins at once: \blackrubric{Cleanse my heart} is not said, nor is a blessing asked, and lights and incense are not carried; neither is \blackrubric{The Lord be with you} said, nor the response, \blackrubric{Glory be to thee, O Lord,} and the Celebrant or Deacon, while be announces The Passion of our Lord, does not sign the book nor himself. Which is observed likewise on other days, when the Passion is read.
\end{rubric}

\readingcitation{Gospel}{Matthew 26:1}
\lett{A}{t that time:} Jesus said unto his disciples: {\ding{64}} Ye know that after two days is the feast of the passover, and the Son of man is betrayed to be crucified. \textit{C.} Then assembled together the chief priests, and the scribes, and the elders of the people, unto the palace of the high priest, who was called Caiaphas, and consulted that they might take Jesus by subtilty, and kill him. But they said, \textit{S.} Not on the feast day, lest there be an uproar among the people.

\textit{C.} Now when Jesus was in Bethany, in the house of Simon the leper, there came unto him a woman having an alabaster box of very precious ointment, and poured it on his head, as he sat at meat. But when his disciples saw it, they had indignation, saying, \textit{S.} To what purpose is this waste? For this ointment might have been sold for much, and given to the poor. \textit{C.} When Jesus understood it, he said unto them, {\ding{64}} Why trouble ye the woman? for she hath wrought a good work upon me. For ye have the poor always with you; but me ye have not always. For in that she hath poured this ointment on my body, she did it for my burial. Verily I say unto you, Wheresoever this gospel shall be preached in the whole world, there shall also this, that this woman hath done, be told for a memorial of her.

\textit{C.} Then one of the twelve, called Judas Iscariot, went unto the chief priests, and said unto them, \textit{S.} What will ye give me, and I will deliver him unto you? \textit{C.} And they covenanted with him for thirty pieces of silver. And from that time he sought opportunity to betray him.

Now the first day of the feast of unleavened bread the disciples came to Jesus, saying unto him, \textit{S.} Where wilt thou that we prepare for thee to eat the passover? \textit{C.} And he said, {\ding{64}} Go into the city to such a man, and say unto him, The Master saith, My time is at hand; I will keep the passover at thy house with my disciples. \textit{C.} And the disciples did as Jesus had appointed them; and they made ready the passover.

Now when the even was come, he sat down with the twelve. And as they did eat, he said, {\ding{64}} Verily I say unto you, that one of you shall betray me. \textit{C.} And they were exceeding sorrowful, and began every one of them to say unto him, \textit{S.} Lord, is it I? \textit{C.} And he answered and said, {\ding{64}} He that dippeth his hand with me in the dish, the same shall betray me. The Son of man goeth as it is written of him: but woe unto that man by whom the Son of man is betrayed! it had been good for that man if he had not been born. \textit{C.} Then Judas, which betrayed him, answered and said, \textit{S.} Master, is it I? \textit{C.} He said unto him, {\ding{64}} Thou hast said.

\textit{C.} And as they were eating, Jesus took bread, and blessed it, and brake it, and gave it to the disciples, and said, {\ding{64}} Take, eat; this is my body. \textit{C.} And he took the cup, and gave thanks, and gave it to them, saying, {\ding{64}} Drink ye all of it; for this is my blood of the new testament, which is shed for many for the remission of sins. But I say unto you, I will not drink henceforth of this fruit of the vine, until that day when I drink it new with you in my Father’s kingdom.

\textit{C.} And when they had sung an hymn, they went out into the mount of Olives. Then saith Jesus unto them, {\ding{64}} All ye shall be offended because of me this night: for it is written, I will smite the shepherd, and the sheep of the flock shall be scattered abroad. But after I am risen again, I will go before you into Galilee. \textit{C.} Peter answered and said unto him, \textit{S.} Though all men shall be offended because of thee, yet will I never be offended. \textit{C.} Jesus said unto him, {\ding{64}} Verily I say unto thee, That this night, before the cock crow, thou shalt deny me thrice. \textit{C.} Peter said unto him, \textit{S.} Though I should die with thee, yet will I not deny thee. \textit{C.} Likewise also said all the disciples.

Then cometh Jesus with them unto a place called Gethsemane, and saith unto the disciples, {\ding{64}} Sit ye here, while I go and pray yonder. \textit{C.} And he took with him Peter and the two sons of Zebedee, and began to be sorrowful and very heavy. Then saith he unto them, {\ding{64}} My soul is exceeding sorrowful, even unto death: tarry ye here, and watch with me. \textit{C.} And he went a little further, and fell on his face, and prayed, saying, {\ding{64}} O my Father, if it be possible, let this cup pass from me: nevertheless not as I will, but as thou wilt.

\textit{C.} And he cometh unto the disciples, and findeth them asleep, and saith unto Peter, {\ding{64}} What, could ye not watch with me one hour? Watch and pray, that ye enter not into temptation: the spirit indeed is willing, but the flesh is weak.

\textit{C.} He went away again the second time, and prayed, saying, {\ding{64}} O my Father, if this cup may not pass away from me, except I drink it, thy will be done. \textit{C.} And he came and found them asleep again: for their eyes were heavy. And he left them, and went away again, and prayed the third time, saying the same words.

Then cometh he to his disciples, and saith unto them, {\ding{64}} Sleep on now, and take your rest: behold, the hour is at hand, and the Son of man is betrayed into the hands of sinners. Rise, let us be going: behold, he is at hand that doth betray me.

\textit{C.} And while he yet spake, lo, Judas, one of the twelve, came, and with him a great multitude with swords and staves, from the chief priests and elders of the people. Now he that betrayed him gave them a sign, saying, \textit{S.} Whomsoever I shall kiss, that same is he: hold him fast. \textit{C.} And forthwith he came to Jesus, and said, \textit{S.} Hail, master; \textit{C.} and kissed him. And Jesus said unto him, {\ding{64}} Friend, wherefore art thou come? \textit{C.} Then came they, and laid hands on Jesus, and took him.

And, behold, one of them which were with Jesus stretched out his hand, and drew his sword, and struck a servant of the high priest's, and smote off his ear. Then said Jesus unto him, {\ding{64}} Put up again thy sword into his place: for all they that take the sword shall perish with the sword. Thinkest thou that I cannot now pray to my Father, and he shall presently give me more than twelve legions of angels? But how then shall the scriptures be fulfilled, that thus it must be?

\textit{C.} In that same hour said Jesus to the multitudes, {\ding{64}} Are ye come out as against a thief with swords and staves for to take me? I sat daily with you teaching in the temple, and ye laid no hold on me. But all this was done, that the scriptures of the prophets might be fulfilled. \textit{C.} Then all the disciples forsook him, and fled.

And they that had laid hold on Jesus led him away to Caiaphas the high priest, where the scribes and the elders were assembled. But Peter followed him afar off unto the high priest's palace, and went in, and sat with the servants, to see the end. Now the chief priests, and elders, and all the council, sought false witness against Jesus, to put him to death; but found none: yea, though many false witnesses came, yet found they none. At the last came two false witnesses, and said, \textit{S.} This fellow said, I am able to destroy the temple of God, and to build it in three days. \textit{C.} And the high priest arose, and said unto him, \textit{S.} Answerest thou nothing? what is it which these witness against thee? \textit{C.} But Jesus held his peace. And the high priest answered and said unto him, \textit{S.} I adjure thee by the living God, that thou tell us whether thou be the Christ, the Son of God. \textit{C.} Jesus saith unto him, {\ding{64}} Thou hast said: nevertheless I say unto you, Hereafter shall ye see the Son of man sitting on the right hand of power, and coming in the clouds of heaven. \textit{C.} Then the high priest rent his clothes, saying, \textit{S.} He hath spoken blasphemy; what further need have we of witnesses? behold, now ye have heard his blasphemy. What think ye? \textit{C.} They answered and said, \textit{S.} He is guilty of death.

\textit{C.} Then did they spit in his face, and buffeted him; and others smote him with the palms of their hands, saying, \textit{S.} Prophesy unto us, thou Christ, Who is he that smote thee?

\textit{C.} Now Peter sat without in the palace: and a damsel came unto him, saying, \textit{S.} Thou also wast with Jesus of Galilee. \textit{C.} But he denied before them all, saying, \textit{S.} I know not what thou sayest. \textit{C.} And when he was gone out into the porch, another maid saw him, and said unto them that were there, \textit{S.} This fellow was also with Jesus of Nazareth. \textit{C.} And again he denied with an oath, \textit{S.} I do not know the man. \textit{C.} And after a while came unto him they that stood by, and said to Peter, \textit{S.} Surely thou also art one of them; for thy speech bewrayeth thee. \textit{C.} Then began he to curse and to swear, saying, \textit{S.} I know not the man. \textit{C.} And immediately the cock crew. And Peter remembered the word of Jesus, which said unto him, {\ding{64}} Before the cock crow, thou shalt deny me thrice. \textit{C.} And he went out, and wept bitterly.

When the morning was come, all the chief priests and elders of the people took counsel against Jesus to put him to death: and when they had bound him, they led him away, and delivered him to Pontius Pilate the governor.

Then Judas, which had betrayed him, when he saw that he was condemned, repented himself, and brought again the thirty pieces of silver to the chief priests and elders, saying, \textit{S.} I have sinned in that I have betrayed the innocent blood. \textit{C.} And they said, \textit{S.} What is that to us? see thou to that. \textit{C.} And he cast down the pieces of silver in the temple, and departed, and went and hanged himself. And the chief priests took the silver pieces, and said, \textit{S.} It is not lawful for to put them into the treasury, because it is the price of blood. \textit{C.} And they took counsel, and bought with them the potter's field, to bury strangers in. Wherefore that field was called, The field of blood, unto this day. Then was fulfilled that which was spoken by Jeremy the prophet, saying, And they took the thirty pieces of silver, the price of him that was valued, whom they of the children of Israel did value; and gave them for the potter's field, as the Lord appointed me.

And Jesus stood before the governor: and the governor asked him, saying, \textit{S.} Art thou the King of the Jews? \textit{C.} And Jesus said unto him, {\ding{64}} Thou sayest. \textit{C.} And when he was accused of the chief priests and elders, he answered nothing. Then said Pilate unto him, \textit{S.} Hearest thou not how many things they witness against thee? \textit{C.} And he answered him to never a word; insomuch that the governor marvelled greatly.

Now at that feast the governor was wont to release unto the people a prisoner, whom they would. And they had then a notable prisoner, called Barabbas. Therefore when they were gathered together, Pilate said unto them, \textit{S.} Whom will ye that I release unto you? Barabbas, or Jesus which is called Christ? \textit{C.} For he knew that for envy they had delivered him.

When he was set down on the judgment seat, his wife sent unto him, saying, \textit{S.} Have thou nothing to do with that just man: for I have suffered many things this day in a dream because of him. \textit{C.} But the chief priests and elders persuaded the multitude that they should ask Barabbas, and destroy Jesus. The governor answered and said unto them, \textit{S.} Whether of the twain will ye that I release unto you? \textit{C.} They said, \textit{S.} Barabbas. \textit{C.} Pilate saith unto them, \textit{S.} What shall I do then with Jesus which is called Christ? \textit{C.} They all say unto him, \textit{S.} Let him be crucified. \textit{C.} And the governor said, \textit{S.} Why, what evil hath he done? \textit{C.} But they cried out the more, saying, \textit{S.} Let him be crucified.

\textit{C.} When Pilate saw that he could prevail nothing, but that rather a tumult was made, he took water, and washed his hands before the multitude, saying, \textit{S.} I am innocent of the blood of this just person: see ye to it. \textit{C.} Then answered all the people, and said, \textit{S.} His blood be on us, and on our children. \textit{C.} Then released he Barabbas unto them: and when he had scourged Jesus, he delivered him to be crucified.

Then the soldiers of the governor took Jesus into the common hall, and gathered unto him the whole band of soldiers. And they stripped him, and put on him a scarlet robe. And when they had platted a crown of thorns, they put it upon his head, and a reed in his right hand: and they bowed the knee before him, and mocked him, saying, \textit{S.} Hail, King of the Jews! \textit{C.} And they spit upon him, and took the reed, and smote him on the head. And after that they had mocked him, they took the robe off from him, and put his own raiment on him, and led him away to crucify him. And as they came out, they found a man of Cyrene, Simon by name: him they compelled to bear his cross.

And when they were come unto a place called Golgotha, that is to say, a place of a skull, they gave him vinegar to drink mingled with gall: and when he had tasted thereof, he would not drink. And they crucified him, and parted his garments, casting lots: that it might be fulfilled which was spoken by the prophet, They parted my garments among them, and upon my vesture did they cast lots. And sitting down they watched him there; and set up over his head his accusation written, \textsc{This is Jesus the King of the Jews}. Then were there two thieves crucified with him, one on the right hand, and another on the left.

And they that passed by reviled him, wagging their heads, and saying, \textit{S.} Thou that destroyest the temple, and buildest it in three days, save thyself. If thou be the Son of God, come down from the cross. \textit{C.} Likewise also the chief priests mocking him, with the scribes and elders, said, \textit{S.} He saved others; himself he cannot save. If he be the King of Israel, let him now come down from the cross, and we will believe him. He trusted in God; let him deliver him now, if he will have him: for he said, I am the Son of God. \textit{C.} The thieves also, which were crucified with him, cast the same in his teeth.

Now from the sixth hour there was darkness over all the land unto the ninth hour. And about the ninth hour Jesus cried with a loud voice, saying, {\ding{64}} Eli, Eli, lama sabachthani? \textit{C.} that is to say, {\ding{64}} My God, my God, why hast thou forsaken me? \textit{C.} Some of them that stood there, when they heard that, said, \textit{S.} This man calleth for Elias. \textit{C.} And straightway one of them ran, and took a spunge, and filled it with vinegar, and put it on a reed, and gave him to drink. The rest said, \textit{S.} Let be, let us see whether Elias will come to save him.

\textit{C.} Jesus, when he had cried again with a loud voice, yielded up the ghost. \inrub{Here genuflect and pause awhile.} And, behold, the veil of the temple was rent in twain from the top to the bottom; and the earth did quake, and the rocks rent; and the graves were opened; and many bodies of the saints which slept arose, and came out of the graves after his resurrection, and went into the holy city, and appeared unto many. Now when the centurion, and they that were with him, watching Jesus, saw the earthquake, and those things that were done, they feared greatly, saying, \textit{S.} Truly this was the Son of God.

\textit{C.} And many women were there beholding afar off, which followed Jesus from Galilee, ministering unto him: among which was Mary Magdalene, and Mary the mother of James and Joses, and the mother of Zebedee's children.

When the even was come, there came a rich man of Arimath{\ae}a, named Joseph, who also himself was Jesus' disciple: he went to Pilate, and begged the body of Jesus. Then Pilate commanded the body to be delivered. And when Joseph had taken the body, he wrapped it in a clean linen cloth, and laid it in his own new tomb, which he had hewn out in the rock: and he rolled a great stone to the door of the sepulchre, and departed. And there was Mary Magdalene, and the other Mary, sitting over against the sepulchre.
\begin{rubric}
    Here \blackrubric{Cleanse my heart} is said, a blessing is asked, incense is carried without lights, and the book is censed: \blackrubric{The Lord be with you} is not said, and the Celebrant or Deacon does not sign the Book nor himself: and what follows is sung in the tone of the Gospel: at the conclusion of which the Celebrant kisses the Book and is censed. All of which is observed in the other Passions except on Good Friday.
\end{rubric}
Now the next day, that followed the day of the preparation, the chief priests and Pharisees came together unto Pilate, saying, Sir, we remember that that deceiver said, while he was yet alive, After three days I will rise again. Command therefore that the sepulchre be made sure until the third day, lest his disciples come by night, and steal him away, and say unto the people, He is risen from the dead: so the last error shall be worse than the first. Pilate said unto them, Ye have a watch: go your way, make it as sure as ye can. So they went, and made the sepulchre sure, sealing the stone, and setting a watch.
\offertory{Thy rebuke hath broken my heart; I am full of heaviness: I looked for some to have pity on me, but there was no man, neither found I any to comfort me: they gave me gall to eat: and when I was thirsty they gave me vinegar to drink.}

\secret\label{PalmSundaySecret}
\lett{G}{rant,} we beseech thee, O God: that the gift which we offer in the sight of thy Majesty may obtain for us grace to serve thee with devotion, and bring us in the end to everlasting felicity. Through
\communion{Father, if this cup may not pass away from me, except I drink it: thy will be done.}
\postcommunion\label{PalmSundayPostcommunion}
\lett{M}{ay} the operation of this mystery, O Lord: avail for the cleansing of our sins, and for the fulfilment of our godly desires. Through.

\lettspace



%MANUAL ADJUSTMENT:
\clearpage
\subsection{Holy Monday}
\fancyhead[RO,LE]{\textit{Holy Monday}}
\fancyhead[RE,LO]{}


   \introit
   \lett{P}{lead} thou my cause, O Lord, with them that strive with me, and fight thou against them that fight against me: lay hand upon the shield and buckler, and stand up to help me, O Lord, thou strength of my salvation. \textit{Ps.} Bring forth the spear, and stop the way against them that persecute me: say unto my soul: I am thy salvation.
   
   \collect
	\lett{A}{lmighty} God, whose most dear Son went not up to joy but first he suffered pain, and entered not into glory before he was crucified; Mercifully grant that we, walking in the way of the cross, may find it none other than the way of life and peace. Through the same.
	
\begin{rubric}
    \nth{2} Collect \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayCollect}).
\end{rubric}
	
\readingcitation{Epistle}{Isaiah 63:1}
\lett{I}{n those days:} Said Isaiah: Who is this that cometh from Edom, with dyed garments from Bozrah? this that is glorious in his apparel, travelling in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winefat? I have trodden the winepress alone; and of the people there was none with me: for I will tread them in mine anger, and trample them in my fury; and their blood shall be sprinkled upon my garments, and I will stain all my raiment. For the day of vengeance is in mine heart, and the year of my redeemed is come. And I looked, and there was none, to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my fury it upheld me. And I will tread down the people in mine anger, and make them drunk in my fury, and I will bring down their strength to the earth. I will mention the lovingkindnesses of the Lord, and the praises of the Lord according to all that the Lord hath bestowed on us, and the great goodness toward the house of Israel, which he hath bestowed on them according to his mercies, and according to the multitude of his lovingkindnesses. For he said, Surely they are my people, children that will not lie: so he was their Saviour. In all their affliction he was afflicted, and the angel of his presence saved them: in his love and in his pity he redeemed them; and he bare them, and carried them all the days of old. But they rebelled, and vexed his holy Spirit; therefore he was turned to be their enemy, and he fought against them. Then he remembered the days of old, Moses, and his people, saying, Where is he that brought them up out of the sea with the shepherd of his flock? where is he that put his holy Spirit within him? that led them by the right hand of Moses with his glorious arm, dividing the water before them, to make himself an everlasting name? that led them through the deep, as an horse in the wilderness., that they should not stumble? As a beast goeth down into the valley, the spirit of the Lord caused him to rest: so didst thou lead thy people, to make thyself a glorious name. Look down from heaven, and behold from the habitation of thy holiness and of thy glory: where is thy zeal and thy strength, the sounding of thy bowels and of thy mercies toward me? are they restrained? Doubtless thou art our father, though Abraham be ignorant of us, and Israel acknowledge us not, thou, O Lord, art our father our redeemer; thy name is from everlasting. O Lord, why hast thou made us to err from thy ways, and hardened our heart from thy fear? Return for thy servants' sake, the tribes of thine inheritance. The people of thy holiness have possessed it but a little while: our adversaries have trodden down thy sanctuary. We are thine: thou never barest rule over them; they were not called by thy name.

\gradtr{Awake, O Lord, and stand up to judge my quarrel, avenge thou my cause, my God, and my Lord. ℣. Bring forth the spear, and stop the way against them that persecute me.}{O Lord, deal not with us after our sins which we have committed: nor reward us according to our wickednesses. ℣. Lord, remember not our old sins, but have mercy upon us, and that soon: for we are come to great misery. \inrub{Here genuflect.} ℣. Help us, O God of our salvation: and for the glory of thy name, O Lord, deliver us: and be merciful unto our sins, for thy name's sake. \inrub{Arise.}}

\readingcitation{Gospel}{Mark 14:1}
\lett{A}{t that time:} After two days was the feast of the passover, and of unleavened bread: and the chief priests and the scribes sought how they might take him by craft, and put him to death. But they said, Not on the feast day, lest there be an uproar of the people. And being in Bethany in the house of Simon the leper, as he sat at meat, there came a woman having an alabaster box of ointment of spikenard very precious; and she brake the box, and poured it on his head. And there were some that had indignation within themselves, and said, Why was this waste of the ointment made? for it might have been sold for more than three hundred pence, and have been given to the poor. And they murmured against her. And Jesus said, Let her alone; why trouble ye her? she hath wrought a good work on me. For ye have the poor with you always, and whensoever ye will ye may do them good: but me ye have not always. She hath done what she could: she is come aforehand to anoint my body to the burying. Verily I say unto you, Wheresoever this gospel shall be preached throughout the whole world, this also that she hath done shall be spoken of for a memorial of her. And Judas Iscariot, one of the twelve, went unto the chief priests, to betray him unto them. And when they heard it, they were glad, and promised to give him money. And he sought how he might conveniently betray him. And the first day of unleavened bread, when they killed the passover, his disciples said unto him, Where wilt thou that we go and prepare that thou mayest eat the passover? And he sendeth forth two of his disciples, and saith unto them, Go ye into the city, and there shall meet you a man bearing a pitcher of water: follow him. And wheresoever he shall go in, say ye to the good-man of the house, The Master saith, Where is the guestchamber, where I shall eat the passover with my disciples? And he will shew you a large upper room furnished and prepared: there make ready for us. And his disciples went forth, and came into the city, and found as he had said unto them: and they made ready the passover. And in the evening he cometh with the twelve. And as they sat and did eat, Jesus said, Verily I say unto you, One of you which eateth with me shall betray me. And they began to be sorrowful, and to say unto him one by one, Is it I? and another said, Is it I? And he answered and said unto them, It is one of the twelve, that dippeth with me in the dish. The Son of man indeed goeth, as it is written of him: but woe to that man by whom the Son of man is betrayed! good were it for that man if he had never been born. And as they did eat, Jesus took bread, and blessed, and brake it, and gave to them, and said, Take, eat; this is my body. And he took the cup, and when he had given thanks, he gave it to them: and they all drank of it. And he said unto them, This is my blood of the new testament, which is shed for many. Verily I say unto you, I will drink no more of the fruit of the vine, until that day that I drink it new in the kingdom of God. And when they had sung an hymn, they went out into the mount of Olives. And Jesus saith unto them, All ye shall be offended because of me this night: for it is written, I will smite the shepherd, and the sheep shall be scattered. But after that I am risen I will go before you into Galilee. But Peter said unto him, Although all shall be offended, yet will not I. And Jesus saith unto him, Verily I say unto thee, that this day even in this night, before the cock crow twice, thou shalt deny me thrice. But he spake the more vehemently, If I should die with thee, I will not deny thee in any wise. Likewise also said they all. And they came to a place which was named Gethsemane: and he saith to his disciples, Sit ye here, while I shall pray. And he taketh with him Peter and James and John, and began to be sore amazed, and to be very heavy; and saith unto them, My soul is exceeding sorrowful unto death: tarry ye here, and watch. And he went forward a little, and fell on the ground, and prayed that, if it were possible, the hour might pass from him. And he said, Abba, Father, all things are possible unto thee; take away this cup from me: nevertheless not what I will, but what thou wilt. And he cometh. and findeth them sleeping, and saith unto Peter, Simon, sleepest thou? couldest not thou watch one hour? Watch ye and pray, lest ye enter into temptation. The spirit truly is ready, but the flesh is weak. And again lie went away, and prayed, and spake the same words. And when ho returned, he found them asleep again, (for their eyes were heavy,) neither wist they what to answer him. And he cometh the third time, and saith unto them, Sleep on now, and take your rest: it is enough, the hour is come; behold, the Son of man is betrayed into the hands of sinners. Rise up, let us go; lo, he that betrayeth me is at hand. And immediately, while he yet spake, cometh Judas, one of the twelve, and with him a great multitude with swords and staves, from the chief priests and the scribes and the elders. And he that betrayed him had given them a token, saying, Whomsoever I shall kiss, that same is he; take him, and lead him away safely. And as soon as he was come, he goeth straightway to him, and saith, Master, master; and kissed him. And they laid their hands on him, and took him. And one of them that stood by drew a sword, and smote a servant of the high priest, and cut off his ear. And Jesus answered and said unto them, Are ye come out, as against a thief, with swords and with staves, to take me? I was daily with you in the temple teaching, and ye took me not: but the scriptures must be fulfilled. And they all forsook him, and fled. And there followed him a certain young man, having a linen cloth cast about his naked body; and the young men laid hold on him: and he left the linen cloth, and fled from them naked. And they led Jesus away to the high priest: and with him were assembled all the chief priests and the elders and the scribes. And Peter followed him afar off, even into the palace of the high priest: and he sat with the servants, and warmed himself at the fire. And the chief priests and all the council sought for witness against Jesus to put him to death; and found none. For many bare false witness against him, but their witness agreed not together. And there arose certain, and bare false witness against him, saying, We heard him say, I will destroy this temple that is made with hands, and within three days I will build another made without hands. But neither so did their witness agree together. And the high priest stood up in the midst, and asked Jesus, saying, Answerest thou nothing? what is it which these witness against thee? But he held his peace, and answered nothing. Again the high priest asked him, and said unto him, Art thou the Christ, the Son of the Blessed? And Jesus said, I am: and ye shall see the Son of man sitting on the right hand of power, and coming in the clouds of heaven. Then the high priest rent his clothes, and saith, What need we any further witnesses? ye have heard the blasphemy: what think ye? And they all condemned him to be guilty of death. And some began to spit on him, and to cover his face, and to buffet him, and to say unto him, Prophesy: and the servants did strike him with the palms of their hands. And as Peter was beneath in the palace, there cometh one of the maids of the high priest: and when she saw Peter warming himself, she looked upon him, and said, And thou also wast with Jesus of Nazareth. But he denied, saying, I know not, neither understand I what thou sayest. And he went out into the porch; and the cock crew. And a maid saw him again, and began to say to them that stood by, This is one of them. And he denied it again. And a little after, they that stood by said again to Peter, Surely thou art one of them: for thou art a Galilean, and thy speech agreeth thereto. But he began to curse and to swear, saying, I know not this man of whom ye speak. And the second time the cock crew. And Peter called to mind the word that Jesus said unto him, Before the cock crow twice, thou shalt deny me thrice. And when he thought thereon, he wept.

\offertory{Deliver me, O Lord, from mine enemies: for I flee unto thee to hide me, teach me to do the thing that pleaseth thee: for thou art my God.}

\secret
\lett{A}{lmighty} God, let these sacrifices cleanse us by their effectual power, that we may thereby attain in purity unto the author of the same. Through.

\begin{rubric}
    \nth{2} Secret \blackrubric{of Palm Sunday} (p. \pageref{PalmSundaySecret}).
\end{rubric}

\vspace{-1ex}

\communion{Let them be put to confusion and shame together, that rejoice at my trouble: let them be clothed with rebuke and dishonour that boast themselves against me.}

\vspace{-1ex}

\postcommunion
\lett{M}{ay} thy holy mysteries, O Lord, impart unto us heavenly fervour: that we may delight in their performance, and likewise in bringing forth the fruit thereof. Through.

\begin{rubric}
    \nth{2} Postcommunion \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayPostcommunion}).
\end{rubric}

\textsc{Priest.} Let us pray.\par
\textsc{Deacon.} Humble your heads before God.\par
\lett{H}{elp} us, O God of our salvation: and grant, that we may come with joy to the celebration of those benefits by which thou hast vouchsafed to redeem us. Through.


%MANUAL ADJUSTMENT:
\clearpage
\subsection{Holy Tuesday}
\fancyhead[RO,LE]{\textit{Holy Tuesday}}
\fancyhead[RE,LO]{}

\introit
\lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life, and resurrection: by whom we are saved and set free. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.

\collect
	\lett{O}{Lord} God, whose blessed Son, our Saviour, gave his back to the smiters and hid not his face from shame; Grant us grace to take joyfully the sufferings of the present time, in full assurance of the glory that shall be revealed. Through the same.
	
	\begin{rubric}
    \nth{2} Collect \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayCollect}).
	\end{rubric}
	
	\readingcitation{Epistle}{Isaiah 50:5}
	\lett{I}{n those days:} Said Isaiah: The Lord \divineName{God} hath opened mine ear, and I was not rebellious, neither turned away back. I gave my back to the smiters, and my cheeks to them that plucked off the hair: I hid not my face from shame and spitting. For the Lord God will help me; therefore shall I not be confounded: therefore have I set my face like a flint and I know that I shall not be ashamed. He is near that justifieth me; who will contend with me? let us stand together: who is mine adversary? let him come near to me. Behold, the Lord God will help me; who is he that shall condemn me? lo, they all shall wax old as a garment; the moth shall eat them up. Who is among you that feareth the Lord that obeyeth the voice of his servant, that walketh in darkness, and hath no light? let him trust in the name of the Lord, and stay upon his God. Behold, all ye that kindle a fire, that compass yourselves about with sparks: walk in the light of your fire, and in the sparks that ye have kindled. This shall ye have of mine hand; ye shall lie down in sorrow.
	
	\gradual{Nevertheless, when they were sick, I put on sackcloth, and humbled my soul with fasting: and my prayer shall turn into mine own bosom. ℣. Plead thou my cause, O Lord, with them that strive with me, and fight thou against them that fight against me: lay hand upon the shield and buckler, and stand up to help me.}
	
	\readingcitation{Gospel}{Mark 15:1}
	\lett{A}{t that time:} Straightway in the morning the chief priests held a consultation with the elders and scribes and the whole council, and bound Jesus, and carried him away, and delivered him to Pilate. And Pilate asked him, Art thou the King of the Jews? And he answering said unto him, Thou sayest it. And the chief priests accused him of many things: but he answered nothing. And Pilate asked him again, saying, Answerest thou nothing? behold how many things they witness against thee. But Jesus yet answered nothing; so that Pilate marvelled. Now at that feast he released unto them one prisoner, whomsoever they desired. And there was one named Barabbas, which lay bound with them that had made insurrection with him, who had committed murder in the insurrection. And the multitude crying aloud began to desire him to do as he had ever done unto them. But Pilate answered them, saying, Will ye that I release unto you the King of the Jews? For he knew that the chief priests had delivered him for envy. But the chief priests moved the people, that he should rather release Barabbas unto them. And Pilate answered and said again unto them, What will ye then that I shall do unto him whom ye call the King of the Jews? And they cried out again, Crucify him. Then Pilate said unto them, Why, what evil hath he done? And they cried out the more exceedingly, Crucify him, And so Pilate, willing to content the people, released Barabbas unto them, and delivered Jesus, when he had scourged him, to be crucified. And the soldiers led him away into the hall, called Pr{\ae}torium; and. they called together the whole band. And they clothed him with purple, and platted a crown of thorns, and put it about his head, and began to salute him, Hail, King of the Jews! And they smote him on the head with a reed, and. did spit upon him, and bowing their knees worshipped him. And when they had mocked. him, they took off the purple from him, and put his own clothes on him, and led him out to. crucify him. And they compel one Simon a Cyrenian, who passed by, coming out of the country, the father of Alexander and Rufus, to bear his cross. And they bring him unto the place Golgotha, which is, being interpreted, The place of a skull. And they gave him to drink wine mingled with myrrh: but he received it not. And when they had crucified him, they parted his garments, casting lots upon them, what every man should take. And it was the third hour, and they crucified him. And the superscription of his accusation was written over, \divineName{the King of the Jews}. And with him they crucify two thieves; the one on his right hand, and the other on his left. And the scripture was fulfilled, which saith, And he was numbered with the transgressors. And they that passed by railed on him, wagging their heads, and saying, Ah, thou that destroyest the temple, and buildest it in three days, save thy self, and come down from the cross. Likewise also the chief priests mocking said among them selves with the scribes, He saved others; himself he cannot save. Let Christ the King of Israel descend now from the cross, that we may see and believe. And they that were crucified with him reviled him. And when the sixth hour was come, there was darkness over the whole land until the ninth hour. And at the ninth hour Jesus cried with a loud voice, saying, Eloi, Eloi, lama sabachthani? which is, being interpreted, My God, my God, why hast thou forsaken me? And some of them that stood by, when they heard it, said, Behold, he calleth Elias. And one ran and filled a spunge full of vinegar, and put it on a reed, and gave him to drink, saying, Let alone; let us see whether Elias will come to take him down. And Jesus cried with a loud voice, and gave up the ghost. \inrub{Here genuflect, and pause awhile.} And the veil of the temple was rent in twain from the top to the bottom. And when the centurion, which stood over against him, saw that he so cried out, and gave up the ghost, he said, Truly this man was the Son of God.
	
	\offertory{Keep me, O Lord, from the hands of the ungodly: preserve me from the wicked men.}
		
	\secret
	\lett{W}{e} beseech thee, O Lord: that these sacrifices which we offer unto thee in this time of fasting, ordained for the healing of our souls, may effectually renew us. Through.

	\begin{rubric}
    \nth{2} Secret \blackrubric{of Palm Sunday} (p. \pageref{PalmSundaySecret}).
	\end{rubric}
	
	\communion{They that sit in the gate speak against me: and the drunkards make songs upon me: but, Lord, I make my prayer unto thee: in an acceptable time, O God, in the multitude of thy mercy.}
	
	\postcommunion
	\lett{A}{lmighty} God, let thy holy mysteries both heal our vices and bestow upon us everlasting remedies. Through.
	
	\lettspace
	
	\begin{rubric}
    \nth{2} Postcommunion \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayPostcommunion}).
	\end{rubric}
	
	\textsc{Priest.} Let us pray.\par
	\textsc{Deacon.} Humble your heads before God.\par
	\lett{L}{et} thy merciful kindness, O God, cleanse us throughly from all the corruption of our former nature, and make us to be effectually renewed unto holiness of life. Through.


%MANUAL ADJUSTMENT:
\clearpage
\subsection{Spy Wednesday}
\fancyhead[RO,LE]{\textit{Spy Wednesday}}
\fancyhead[RE,LO]{}

\introit

%MANUAL ADJUSTMENT:
\vspace{-0.25\baselineskip}

\lett{A}{t} the name of Jesus every knee should bow, of things in heaven, and things in earth, and things under the earth: for the Lord became obedient unto death, even the death of the cross: wherefore Jesus Christ is Lord, to the glory of God the Father. \textit{Ps.} Hear my prayer, O Lord: and let my crying come unto thee.

\begin{rubric}
    After the Kyrie,
\end{rubric}
\elcol{℣. Let us pray.}{℣. Orémus.}
\elcol{℣. Let us bow the knee.}{℣. Flectámus génua.}
\elcol{℣. Arise.}{℣. Leváte.}

%MANUAL ADJUSTMENT:
\vspace{-0.25\baselineskip}

\collect

%MANUAL ADJUSTMENT:
\vspace{-0.15\baselineskip}

\lett{G}{rant,} we beseech thee, almighty God: that we, who are continually afflicted by reason of our transgressions, may be delivered by the passion of thine only-begotten Son. Who liveth.


%MANUAL ADJUSTMENT:
\vspace{-0.25\baselineskip}

\readingcitation{Epistle}{Isaiah 62:11}

%MANUAL ADJUSTMENT:
\vspace{-0.1\baselineskip}

\lett{T}{hus saith the Lord God:} Say ye to the daughter of Zion, Behold, thy salvation cometh; behold, his reward is with him. Who is this that cometh from Edom, with dyed garments from Bozrah? this that is glorious in his apparel, travelling in the greatness of his strength? I that speak in righteousness, mighty to save. Wherefore art thou red in thine apparel, and thy garments like him that treadeth in the winefat? I have trodden the winepress alone; and of the people there was none with me: for I will tread them in mine anger, and trample them in my fury; and their blood shall be sprinkled upon my garments, and I will stain all my raiment. For the day of vengeance is in mine heart, and the year of my redeemed is come. And I looked, and there was none to help; and I wondered that there was none to uphold: therefore mine own arm brought salvation unto me; and my fury, it upheld me. And I will tread down the people in mine anger, and make them drunk in my fury, and I will bring down their strength to the earth. I will mention the lovingkindnesses of the \divineName{Lord}, and the praises of the \divineName{Lord}, according to all that the \divineName{Lord} hath bestowed on us.

\gradual{Hide not thy face from thy servant, for I am in trouble: O haste thee, and hear me. ℣. Save me, O God, for the waters are come in, even unto my soul: I stick fast in the deep mire, where no ground is.}\\

\elcol{℣. The Lord be with you.}{℣. Dóminus vobíscum.}
\elcol{℟. And with thy spirit.}{℟. Et cum spíritu tuo.}
\elcol{℣. Let us pray.}{℣. Orémus.}

\collect
\lett{A}{ssist} us mercifully with thy help, O Lord God of our salvation; that we may enter with joy upon the meditation of those mighty acts, whereby thou hast given unto us life and immortality. Through.
	
	\begin{rubric}
    \nth{2} Collect \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayCollect}).
	\end{rubric}

\readingcitation{Epistle}{Hebrews 9:16}
\lett{B}{rethren:} Where a testament is, there must also of necessity be the death of the testator. For a testament is of force after men are dead: otherwise it is of no strength at all while the testator liveth. Whereupon neither the first testament was dedicated without blood. For when Moses had spoken every precept to all the people according to the law, he took the blood of calves and of goats, with water, and. scarlet wool, and hyssop, and sprinkled both the book and all the people, saying, This is the blood of the testament which God hath enjoined unto you. Moreover, he sprinkled likewise with blood both the tabernacle, and all the vessels of the ministry. And almost all things are by the law purged with blood; and without shedding of blood is no remission. It was therefore necessary that the patterns of things in the heavens should be purified with these; but the heavenly things themselves with better sacrifices than these. For Christ is not entered into the holy places made with hands, which are the figures of the true; but into heaven itself, now to appear in the presence of God for us: nor yet that he should offer himself often, as the high priest entereth into the holy place every year with blood of others; for then must he often have suffered since the foundation of the world: but now once in the end of the world hath he appeared to put away sin by the sacrifice of himself. And as it is appointed unto men once to die, but after this the judgment: so Christ was once offered to bear the sins of many; and unto them that look for him shall he appear the second time without sin unto salvation.

\tract{Hear my prayer, O Lord, and let my crying come unto thee. ℣. Hide not thy face from me in the time of my trouble. ℣. Incline thine ear unto me when I call, O hear me, and that right soon. ℣. For my days are consumed away like smoke: and my bones are burnt up as it were a fire-brand. ℣. My heart is smitten down, and withered like grass: so that I forget to eat my bread. ℣. Thou shalt arise, O Lord, and have mercy upon Sion: for it is time that thou have mercy upon her.}

\readingcitation{Gospel}{Luke 22:1}
\lett{A}{t that time:} The feast of unleavened bread drew nigh, which is called the Passover. And the chief priests and scribes sought how they might kill him; for they feared the people. Then entered Satan into Judas surnamed Iscariot, being of the number of the twelve. And he went his way, and communed with the chief priests and captains, how he might betray him unto them. And they were glad, and covenanted to give him money. And he promised, and sought opportunity to betray him unto them in the absence of the multitude. Then came the day of unleavened bread, when the passover must be killed. And he sent Peter and John, saying, Go and prepare us the passover, that we may eat. And they said unto him, Where wilt thou that we prepare? And he said unto them, Behold, when ye are entered into the city, there shall a man meet you, bearing a pitcher of water; follow him into the house where he entereth in. And ye shall say unto the goodman of the house, The Master saith unto thee, Where is the guest-chamber, where I shall eat the passover with my disciples? And he shall shew you a large upper room furnished: there make ready. And they went, and found as he had said unto them: and they made ready the passover. And when the hour was come, he sat down, and the twelve apostles with him. And he said unto them, With desire I have desired to eat this passover with you before I suffer: for I say unto you, I will not any more eat thereof, until it be fulfilled in the kingdom of God. And he took the cup, and gave thanks, and said, Take this, and divide it among yourselves: for I say unto you I will not drink of the fruit of the vine, until the kingdom of God shall come. And he took bread, and gave thanks, and brake it, and gave unto them, saying, This is my body which is given for you: this do in remembrance of me. Likewise also the cup after supper, saying, This cup is the new testament in my blood, which is shed for you. But, behold, the hand of him that betrayeth me is with me on the table. And truly the Son of man goeth, as it was determined: but woe unto that man by whom he is betrayed! And they began to enquire among themselves, which of them it was that should do this thing. And there was also a strife among them, which of them should be accounted the greatest And he said unto them, The kings of the Gentiles exercise lordship over them; and they that exercise authority upon them are called benefactors. But ye shall not be so: but he that is greatest among you, let him be as the younger; and he that is chief, as he that doth serve. For whether is greater, he that sitteth at meat, or he that serveth? is not he that sitteth at meat? but I am among you as he that serveth. Ye are they which have continued with me in my temptations. And I appoint unto you a kingdom, as my Father hath appointed unto me; that ye may eat and drink at my table in my kingdom. and sit on thrones, judging the twelve tribes of Israel. And the Lord said, Simon, Simon, behold, Satan hath desired to have you, that he may sift you as wheat: but I have prayed for thee, that thy faith fail not; and when thou art converted, strengthen thy brethren. And he said unto him, Lord, I am ready to go with thee, both into prison, and to death. And he said, I tell thee, Peter, the cock shall not crow this day, before that thou shalt thrice deny that thou knowest me. And he said unto them, When I sent you without purse, and scrip, and shoes, lacked ye any thing? And they said, Nothing. Then said he unto them, But now, he that hath a purse, let him take it, and likewise his scrip: and he that hath no sword, let him sell his garment, and buy one. For I say unto you, that this that is written must yet be accomplished in me, And he was reckoned among the transgressors: for the things concerning me have an end. And they said, Lord, behold, here are two swords. And he said unto them, It is enough. And he came out, and went, as he was wont, to the mount of Olives; and his disciples also followed him. And when he was at the place, he said unto them, Pray that ye enter not into temptation. And he was withdrawn from them about a stone's cast, and kneeled down, and prayed, saying, Father, if thou be willing, remove this cup from me: nevertheless not my will, but thine, be done. And there appeared an angel unto him from heaven, strengthening him. And being in an agony he prayed more earnestly: and his sweat was as it were great drops of blood falling down to the ground. And when he rose up from prayer, and was come to his disciples, he found them sleeping for sorrow, and said unto them, Why sleep ye? rise and pray, lest ye enter into temptation. And while he yet spake behold a multitude, and he that was called Judas, one of the twelve, went before them, and drew near unto Jesus to kiss him. But Jesus said unto him. Judas, betrayest thou the Son of man with a kiss? When they which were about him saw what would follow, they said unto him, Lord, shall we smite with the sword? And one of them smote the servant of the high priest, and cut off his right ear. And Jesus answered and said, Suffer ye thus far. And he touched his ear, and healed him. Then Jesus said unto the chief priests, and captains of the temple, and the elders, which were come to him, Be ye come out, as against a thief, with swords and staves? When I was daily with you in the temple, ye stretched forth no hands against me: but this is your hour, and the power of darkness. Then took they him, and led him, and brought him into the high priest's house. And Peter followed afar off. And when they had kindled a fire in the midst of the hall, and were set down together, Peter sat down among them. But a certain maid beheld him as he sat by the fire, and earnestly looked upon him, and said, This man was also with him. And he denied him, saying, Woman, I know him not. And after a little while another saw him, and said, Thou art also of them. And Peter said, Man, I am not. And about the space of one hour after another confidently affirmed, saying, Of a truth this fellow also was with him: for he is a Galilean. And Peter said, Man, I know not what thou sayest. And immediately, while he yet spake, the cock crew. And the Lord turned, and looked upon Peter. And Peter remembered the word of the Lord, how he had said unto him, Before the cock crow, thou shalt deny me thrice. And Peter went out, and wept bitterly. And the men that held Jesus mocked him, and smote him. And when they had blindfolded him, they struck him on the face, and asked him, saying, Prophesy, who is it that smote thee? And many other things blasphemously spake they against him. And as soon as it was day the elders of the people and the chief priests and the scribes came together, and led him into their council, saying, Art thou the Christ? tell us. And he said unto them, If I tell you, ye will not believe: and if I also ask you, ye will not answer me, nor let me go. Hereafter shall the Son of man sit on the right hand of the power of God. Then said they all, Art thou then the Son of God? And he said unto them, Ye say that I am. And they said, What need we any further witness? for we ourselves have heard of his own mouth.

\offertory{Hear my prayer, O Lord, and let my crying come unto thee: hide not thy face from me.}

%MANUAL ADJUSTMENT:
\vspace{-0.4\baselineskip}

\secret

%MANUAL ADJUSTMENT:
\vspace{-0.1\baselineskip}

\lett{A}{ccept,} we beseech thee, O Lord, the gift which we offer, and vouchsafe so to work in us: that as we set forth in a mystery the passion of thy Son our Lord, so by devout affection we may attain unto the same. Through the same.

	\begin{rubric}
    \nth{2} Secret \blackrubric{of Palm Sunday} (p. \pageref{PalmSundaySecret}).
	\end{rubric}

\communion{I have mingled my drink with weeping: for thou hast taken me up, and cast me down: and I am withered like grass: but thou, O Lord, shalt endure for ever: thou shalt arise, and have mercy upon Sion, for it is time that thou have mercy upon her.}

%MANUAL ADJUSTMENT:
\vspace{-0.4\baselineskip}

\postcommunion

%MANUAL ADJUSTMENT:
\vspace{-0.1\baselineskip}

\lett{A}{lmighty} God, enlighten our understanding: that, as these wondrous mysteries do testify to the temporal death of thy Son, so thereby we may have assurance that thou hast bestowed upon us life eternal. Through the same.

	\begin{rubric}
    \nth{2} Postcommunion \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayPostcommunion}).
	\end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
	\textsc{Priest.} Let us pray.\par
	\textsc{Deacon.} Humble your heads before God.\par
	\lett{A}{lmighty} God, we beseech thee graciously to behold this thy family, for which our Lord Jesus Christ was contented to be betrayed, and given up into the hands of wicked men, and to suffer death upon the Cross. Who liveth.
	
	
\section{Sacred Triduum}

\subsection{Maundy Thursday}
\fancyhead[RO,LE]{\textit{Maundy Thursday}}
\fancyhead[RE,LO]{}
\begin{inhead}
    {First Class Double}
\end{inhead}
    
    \introit
    \lett{B}{ut} it behoveth us to glory in the Cross of our Lord Jesus Christ: in whom is our salvation, life, and resurrection: by whom we are saved, and set free. \textit{Ps.} God be merciful unto us, and bless us: and shew us the light of his countenance, and be merciful unto us.
    
    \begin{rubric}
    	\blackrubric{Gloria in excelsis} is said, and after the intonation thereof the bells are rung: and, the Hymn being ended, they are silent until Easter Even, as is there noted.
    \end{rubric}
    
    
    \collect
    \lett{A}{lmighty} Father, whose dear Son, on the night before he suffered, did institute the Sacrament of his Body and Blood; Mercifully grant that we may thankfully receive the same in remembrance of him, who in these holy mysteries giveth us a pledge of life eternal; through the same thy Son Jesus Christ our Lord, who now liveth and reigneth with thee and the Holy Spirit ever, one God, world without end. \textit{Amen.}
    
    \begin{rubric}
    	\nth{2} Collect \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayCollect}).
    \end{rubric}
    
    \readingcitation{Epistle}{1 Corinthians 11:17}
    \lett{B}{rethren:} In this that I declare unto you I praise you not, that ye come together not for the better, but for the worse. For first of all, when ye come together in the church, I hear that there be divisions among you; and I partly believe it. For there must be also heresies among you, that they which are approved may be made manifest among you. When ye come together therefore into one place, this is not to eat the Lord's supper. For in eating every one taketh before other his own supper: and one is hungry, and another is drunken. What? have ye not houses to eat and to drink in? or despise ye the church of God, and shame them that have not? What shall I say to you? shall I praise you in this? I praise you not. For I have received of the Lord that which also I delivered unto you, That the Lord Jesus the same night in which he was betrayed took bread: and when he had given thanks, he brake it, and said, Take, eat: this is my body, which is broken for you: this do in remembrance of me After the same manner also he took the cup, when he had supped, saying, This cup is the new testament in my blood: this do ye, as oft as ye drink it, in remembrance of me. For as often as ye eat this bread, and drink this cup, ye do shew the Lord's death till he come. Wherefore whosoever shall eat this bread, and drink this cup of the Lord, unworthily, shall be guilty of the body and blood of the Lord. But let a man examine himself, and so let him eat of that bread, and drink of that cup. For he that eateth and drinketh unworthily, eateth and drinketh damnation to himself, not discerning the Lord's body. For this cause many are weak and sickly among you, and many sleep. For if we would judge ourselves, we should not be judged. But when we are judged, we are chastened of the Lord, that we should not be condemned with the world. Wherefore, my brethren, when ye come together to eat, tarry one for another. And if any man hunger, let him eat at home; that ye come not together unto condemnation. And the rest will I set in order when I come.
    
    \gradual{Christ for us became obedient unto death, even the death of the cross. ℣. Wherefore God also hath highly exalted him: and given him a name which is above every name.}
    
    \readingcitation{Gospel}{Luke 23:1}
    \lett{A}{t that time:} The whole multitude of them arose, and led him unto Pilate. And they began to accuse him, saying, We found this fellow perverting the nation, and forbidding to give tribute to Cæsar, saying that he himself is Christ a King. And Pilate asked him, saying, Art thou the King of the Jews? And he answered him and said, Thou sayest it. Then said Pilate to the chief priests and to the people, I find no fault in this man. And they were the more fierce, saying, He stirreth up the people, teaching through-out all Jewry, beginning from Galilee to this place. When Pilate heard of Galilee, he asked whether the man were a Galilean. And as soon as he knew that he belonged unto Herod's jurisdiction, he sent him to Herod, who himself also was at Jerusalem at that time. And when Herod saw Jesus, he was exceeding glad: for he was desirous to see him of a long season, because he had heard many things of him; and he hoped to have seen some miracle done by him. Then he questioned with him in many words; but he answered him nothing. And the chief priests and scribes stood and vehemently accused him. And Herod with his men of war set him at nought, and mocked him, and arrayed him in a gorgeous robe, and sent him again to Pilate. And the same day Pilate and Herod were made friends together: for before they were at enmity between themselves. And Pilate, when he had called together the chief priests and the rulers and the people, said unto them, Ye have brought this man unto me, as one that perverted the people: and, behold, I, having examined him before you, have found no fault in this man touching those things whereof ye accuse him: no, nor yet Herod: for I sent you to him; and, lo, nothing worthy of death is done unto him. I will therefore chastise him, and release him. (For of necessity he must release one unto them at the feast.) And they cried out all at once, saying, Away with this man, and release unto us Barabbas: (who for a certain sedition made in the city, and for murder, was cast into prison.) Pilate therefore, willing to release Jesus, spake again to them. But they cried, saying, Crucify him, crucify him. And he said unto them the third time, Why, what evil hath he done? I have found no cause of death in him: I will therefore chastise him, and let him go. And they were instant with loud voices, requiring that he might be crucified. And the voices of them and of the chief priests prevailed. And Pilate gave sentence that it should be as they required. And he released unto them him that for sedition and murder was cast into prison, whom they had desired; but he delivered Jesus to their will. And as they led him away, they laid hold upon one Simon, a Cyrenian, coming out of the country, and on him they laid the cross, that he might bear it after Jesus. And there followed him a great company of people, and of women, which also bewailed and lamented him. But Jesus turning unto them said, Daughters of Jerusalem, weep not for me, but weep for yourselves, and for your children. For, behold, the days are coming, in the which they shall say, Blessed are the barren, and the wombs that never bare, and the paps which never gave suck. Then shall they begin to say to the mountains, Fall on us; and to the hills, Cover us. For if they do these things in a green tree, what shall be done in the dry? And there were also two other, malefactors, led with him to be put to death. And when they were come to the place which is called Calvary, there they crucified him, and the malefactors, one on the right hand, and the other on the left. Then said Jesus, Father, forgive them; for they know not what they do. And they parted his raiment, and cast lots. And the people stood beholding. And the rulers also with them derided him, saying, He saved others; let him save himself, if he be Christ, the chosen of God. And the soldiers also mocked him, coming to him, and offering him vinegar, and saying, If thou be the king of the Jews, save thyself. And a superscription also was written over him, in letters of Greek, and Latin, and Hebrew, \textsc{This is the King of the Jews}. And one of the malefactors which were hanged railed on him, saying, If thou be Christ, save thyself and us. But the other answering rebuked him, saying, Dost not thou fear God, seeing thou art in the same condemnation? And we indeed justly; for we receive the due reward of our deeds: but this man hath done nothing amiss. And he said unto Jesus, Lord, remember me when thou comest into thy kingdom. And Jesus said unto him, Verily I say unto thee, To day shalt thou be with me in paradise. And it was about the sixth hour, and there was a darkness over all the earth until the ninth hour. And the sun was darkened, and the veil of the temple was rent in the midst. And when Jesus had cried with a loud voice, he said, Father, into thy hands I commend my spirit: and having said thus, he gave up the ghost. \inrub{Here genuflect, and pause awhile.} Now when the centurion saw what was done, he glorified God, saying, Certainly this was a righteous man. And all the people that came together to that sight, beholding the things which were done, smote their breasts, and returned. And all his acquaintance, and the women that followed him from Galilee, stood afar off, beholding these things.
    
    \offertory{The right hand of the Lord bringeth mighty things to pass, the right of the Lord hath exalted me: I shall not die, but live, and declare the works of the Lord.}
    
    \secret
    \lett{W}{e} beseech thee, O Lord holy, Father almighty, everlasting God: that he may render our sacrifice acceptable unto thee, who on this day commanded and taught his disciples to do this in remembrance of him, even Jesus Christ thy Son our Lord. Who liveth.
    
    \begin{rubric}
    	\nth{2} Secret \blackrubric{of Palm Sunday} (p. \pageref{PalmSundaySecret}).
    \end{rubric}
    
    \begin{rubric}
    	After the Comfortable Words, the \blackrubric{Pax} is not given.
    \end{rubric}
    
    \begin{rubric}
    	\textsc{Note,} To-day the Priest consecrates two Hosts, of which he consumes one, and reserves the other (touched with the Blood) for the following day, on which the Sacrament is not consecrated; he also reserves some consecrated particles, in case of need, for the sick.\par
    	But he consumes the whole of the Blood, and before washing his fingers he places the reserved Host in another Chalice, which the Deacon covers with the Pall and the Paten inverted, and spreads a Veil over it, and places it in the midst of the Altar.\par
    	Then the Communion is said and the Mass finished. The Priest genuflects, however, whenever he approaches or departs from the middle of the Altar, or crosses before the Sacrament reserved in the Chalice: and when he is to say \blackrubric{The Lord be with you,} he does not turn to the people in the middle of the Altar, lest be turn his back on the Sacrament, but from the Gospel side: and at the end he gives the blessing from the same place and does not complete the circle.
    \end{rubric}
    
    \communion{The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do.}
    
    \postcommunion
    \lett{O}{Lord} our God, who hast refreshed us with life-giving sustenance, we beseech thee: that we who observe this institution in the time of our mortal life, may obtain the benefit thereof in thy gift of immortality. Through.
    
    \begin{rubric}
    	\nth{2} Postcommunion \blackrubric{of Palm Sunday} (p. \pageref{PalmSundayPostcommunion}).
    \end{rubric}
    
    \begin{rubric}
    	\blackrubric{Ite, Missa est} is said, and the Blessing is given, and the Gospel of St. John is read, at the beginning of which the Priest does not sign the Altar, but himself only.
    \end{rubric}
    
    \subsubsection{Altar of Repose \& Stripping of the Altar}
    
    \begin{rubric}
    	To-day a suitable place is to be prepared in some Chapel of the Church, or Altar: and adorned as decently as may be, with veils and lights: where the Chalice with the Host, reserved as above, shall be set. When the Mass is over, torches are lighted, and a Procession is formed in the usual manner, but with another Subdeacon vested carrying the Cross. The Celebrant, vested in a white Cope, standing before the Altar, places incense in two thuribles without a blessing: then kneeling in the middle, with one he censes the Sacrament thrice: and receiving the Chalice with the Sacrament from the hand of the Deacon, who stands, and having it covered with the ends of the Veil with which his shoulders are covered, he proceeds between the same Deacon on his right, and the Subdeacon on his left, under a baldachino with two Acolytes continually censing the Sacrament, to the place prepared, where it is to be kept for the morrow. Meanwhile, during the Procession, the following hymn is sung.
    \end{rubric}
    
    %?\subsection{Pange Lingua}
    
    \gregorioscore{resources/gabc/Propers/HolyFerias/PangeLingua.gabc}
    
    
    \begin{rubric}
    	When they arrive at the place prepared, the Deacon kneels down and receives from the Priest, who stands, the Chalice with the Sacrament and places it first upon the Altar, where it is censed by the Priest kneeling, as above: then be places it in the receptacle.
    \end{rubric}

%MANUAL ADJUSTMENT:
\clearpage
    \subsubsection{Psalm 22. \textit{Deus, Deus meus}}
    
    \begin{rubric}
    	Afterwards Evensong is said in the Choir without singing. Which being finished, the Priest, vested in an Alb and violet stole, with the Ministers shall lay bare the Altars, reading the following,
    \end{rubric}
    
        \antiphon{}{They parted my garments among them: and upon my vesture did they cast lots.}

\lett{M}{y} God, my God, look upon me; why hast thou forsaken me : and art so far from my health, and from the words of my complaint?\par
\secondline{\psanum{2}O my God, I cry in the day-time, but thou hearest not : and in the night-season also I take no rest.}
\thirdline{\psanum{3}And thou continuest holy : O thou worship of Israel.}
\psanum{4}Our fathers hoped in thee : they trusted in thee, and thou didst deliver them.\par
\psanum{5}They called upon thee, and were holpen : they put their trust in thee, and were not confounded.\par
\psanum{6}But as for me, I am a worm, and no man : a very scorn of men, and the outcast of the people.\par
\psanum{7}All they that see me laugh me to scorn : they shoot our their lips, and shake their heads, saying,\par
\psanum{8}He trusted in God, that he would deliver him : let him deliver him, if he will have him.\par
\psanum{9}But thou art he that took me out of my mother's womb : thou wast my hope, when I hanged yet upon my mother's breasts.\par
\psanum{10}I have been left unto thee ever since I was born : thou art my God, even from my mother's womb.\par
\psanum{11}O go not from me, for trouble is hard at hand : and there is none to help me.\par
\psanum{12}Many oxen are come about me : fat bulls of Basan close me in on every side.\par
\psanum{13}They gape upon me with their mouths : as it were a ramping and a roaring lion.\par
\psanum{14}I am poured out like water, and all my bones are out of joint : my heart also in the midst of my body is even like melting wax.\par
\psanum{15}My strength is dried up like a potsherd, and my tongue cleaveth to my gums : and thou shalt bring me into the dust of death.\par
\psanum{16}For many dogs are come about me : and the council of the wicked layeth siege against me.\par
\psanum{17}They pierced my hands and my feet; I may tell all my bones : they stand staring and looking upon me.\par
\psanum{18}They part my garments among them : and casts lots upon my vesture.\par
\psanum{19}But be not thou far from me, O Lord : thou art my succour, haste thee to help me.\par
\psanum{20}Deliver my soul from the sword : my darling from the power of the dog.\par
\psanum{21}Save me from the lion's mouth : thou hast heard me also from among the horns of the unicorns.\par
\psanum{22}I will declare thy Name unto my brethren : in the midst of the congregation will I praise thee.\par
\psanum{23}O praise the Lord, ye that fear him : magnify him, all ye of the seed of Jacob, and fear him, all ye seed of Israel.\par
\psanum{24}For he hath not despised, nor abhorred, the low estate of the poor : he hath not hid his face from him, but when he called unto him he heard him.\par
\psanum{25}My praise is of thee in the great congregation : my vows will I perform in the sight of them that fear him.\par
\psanum{26}The poor shall eat and be satisfied : they that seek after the Lord shall praise him; your heart shall live for ever.\par
\psanum{27}All the ends of the world shall remember themselves, and be turned unto the Lord : and all the kindreds of the nations shall worship before him.\par
\psanum{28}For the kingdom is the Lord's : and he is the Governor among the people.\par
\psanum{29}All such as be fat upon earth : have eaten and worshipped.\par
\psanum{30}All they that go down into the dust shall kneel before him : and no man hath quickened his own soul.\par
\psanum{31}My seed shall serve him : they shall be counted unto the Lord for a generation.\par
\psanum{32}They shall come, and the heavens shall declare his righteousness : unto a people that shall be born, whom the Lord hath made.\par

    \antiphon{}{They parted my garments among them: and upon my vesture did they cast lots.}
    
    \subsubsection{Washing of Feet}
    
    \begin{rubric}
    	After the stripping of the Altars, at the appointed hour a signal is given with a board, and the Clergy come together to perform the Maundy.\par
    	The Prelate or Superior is vested with Stole and violet Cope over Amice and Alb, and in the appointed place with the assistance of the Deacon (who with the Subdeacon assists him in white vestments, as at Mass) puts incense in the thurible: then the Deacon holding the book of the Gospels before his breast, kneels before the Superior and asks a blessing: on receiving this, while two Acolytes with lighted candles stand by, and the Subdeacon holds the book, he signs it and censes it, and sings as usual the following Gospel.
    	\end{rubric}
    

    \readingcitation{Gospel}{John 13:1}
    \lett{B}{efore} the feast of the passover, when Jesus knew that his hour was come that he should depart out of this world unto the Father, having loved his own which were in the world, he loved them unto the end. And supper being ended, the devil having now put into the heart of Judas Iscariot, Simon's son, to betray him; Jesus knowing that the Father had given all things into his hands, and that he was come from God, and went to God; he riseth from supper, and laid aside his garments; and took a towel, and girded himself. After that he poureth water into a bason, and began to wash the disciples' feet, and to wipe them with the towel wherewith he was girded. Then cometh he to Simon Peter: and Peter saith unto him, Lord, dost thou wash my feet? Jesus answered and said unto him, What I do thou knowest not now; but thou shalt know hereafter. Peter saith unto him, Thou shalt never wash my feet. Jesus answered him, If I wash thee not, thou hast no part with me. Simon Peter saith unto him, Lord, not my feet only, but also my hands and my head. Jesus saith to him, He that is washed needeth not save to wash his feet, but is clean every whit: and ye are clean, but not all. For he knew who should betray him; therefore said he, Ye are not all clean. So after he had washed their feet, and had taken his garments, and was set down again, he said unto them, Know ye what I have done to you? Ye call me Master and Lord: and ye say well; for so I am. If I then, your Lord and Master, have washed your feet; ye also ought to wash one another's feet. For I have given you an example, that ye should do as I have done to you.
    
    \begin{rubric}
    	When this is finished, the Subdeacon carries the open book to be kissed by the Superior, and the Deacon censes him as usual.\par
    	Then the Superior puts off the Cope, and is girded by the Deacon and Subdeacon with a linen cloth, and thus girded, the Deacon and Subdeacon assisting, proceeds to the washing of feet. Those who are to be washed are arranged in order, and while Clerks minister a bason and water, and the Subdeacon holds the right foot of each, the Superior kneels before each, washes his foot, wipes and kisses it, the Deacon giving the towel to wipe them. Meanwhile the following are sung:
    \end{rubric}
    
    \antiphon{}{A new commandment I give unto you: that ye love one another, as I have loved you, saith the Lord. \textit{Ps. 119.} Blessed are those that are undefiled in the way: and walk in the law of the Lord. ℣. A new commandment I give unto you: that ye love one another, as I have loved you, saith the Lord.}
    
    \antiphon{}{After the Lord had risen from supper, he poured water into a bason, and began to wash his disciples' feet: this example left he unto them. \textit{Ps. 48.} Great is the Lord, and highly to be praised: in the city of our God, even upon his holy hill. ℣. After the Lord had risen from supper, he poured water into a bason, and began to wash his disciples' feet: this example left he unto them.}
    
    \antiphon{}{The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do. \textit{Ps. 85.} Lord, thou art become gracious unto thy land: thou hast turned away the captivity of Jacob. ℣. The Lord Jesus, after he had supped with his disciples, washed their feet, and said to them: Know ye what I, your Lord and Master, have done to you? I have given you an example, that ye also should so do.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. Then cometh he to Simon Peter, and Peter saith unto him.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. What I do, thou knowest not now: but thou shalt know hereafter.}
    
    \antiphon{}{Lord, dost thou wash my feet? Jesus answered, and said unto him: If I wash not thy feet, thou hast no part with me. ℣. If I, your Lord and Master, have washed your feet: how much more ought ye to wash one another's feet? \textit{Ps. 49.} O hear ye this, all ye people: ponder it with your ears, all ye that dwell in the world. ℣. If I, your Lord and Master, have washed your feet: how much more ought ye to wash one another's feet?}
    
    \antiphon{}{By this shall all men know that ye are my disciples, if ye have love one to another. ℣. Jesus said to his disciples. ℣. By this shall all men know that ye are my disciples, if ye have love one to another.}
    
    \antiphon{}{Let there abide in you faith, hope, charity, these three: but the greatest of these is charity. ℣. And now abideth faith, hope, charity, these three, but the greatest of these is charity. ℣. Let there abide in you faith, hope, charity, these three: but the greatest of these is charity.}
    
    \antiphon{}{Blessed be the holy Trinity and the undivided Unity: we will give thanks unto him, because he hath shewed his mercy upon us. ℣. Let us bless the Father and the Son with the Holy Ghost. \textit{Ps. 84.} O how amiable are thy dwellings, thou Lord of hosts! My soul hath a desire and longing to enter into the courts of the Lord. ℣. Blessed be the holy Trinity and the undivided Unity: we will give thanks unto him, because he hath shewed his mercy upon us.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. The love of Christ hath gathered us together in one. ℣. Let us rejoice and be glad in him. ℣. Let us fear and love the living God. ℣. And love one another in sincerity of heart.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. When therefore we are gathered together in one: ℣. Take we heed that we be not divided in mind. ℣. Let malicious quarrels, let contentions cease. ℣. And let Christ our God be in the midst of us.}
    
    \antiphon{}{Where charity and love are, there is God. ℣. So may we also with the Blessed see ℣. In glory thy countenance, O Christ our God: ℣. Joy, that is infinite and undefiled, ℣. For ever and for evermore. Amen.}
    
    \begin{rubric}
    	After the washing, the Superior, or he that hath washed the feet, washes his hands, and wipes then with another towel: then returning to the place where he was at first he puts on the Cope, and standing with head uncovered says:
    \end{rubric}
    
    \begin{rubric}
    	\blackrubric{Our Father,} secretly until,
    \end{rubric}
    
    ℣. And lead us not into temptation.
    
    ℟. But deliver us from evil.
    
    ℣. Thou hast charged, O Lord.
    
    ℟. That we shall diligently keep thy commandments.
    
    ℣. Thou didst wash the feet of thy disciples.
    
    ℟. Despise not thou the works of thine hands.
    
    ℣. O Lord, hear my prayer.
    
    ℟. And let my cry come unto thee.
    
    ℣. The Lord be with you.
    
    ℟. And with thy spirit.
    
    \letuspray
    
    \lett{A}{ssist} us, we beseech thee, O Lord, in this our bounden service: and since thou wast pleased to wash the feet of thy disciples, despise not the works of thine own hands which thou hast bidden us to follow: that as here outward stains by us and from us are washed away; so also the secret sins of us all may be blotted out by thee. Which do thou vouchsafe to grant, who livest and reignest God: throughout all ages, world without end. \textit{Amen.}
