\fancyhead[RO,LE]{\textit{Prefaces}}
\fancyhead[RE,LO]{}
\section{Prefaces for the Mass}\label{prefaces}

\subsection{Proper Prefaces without Notation}

\begin{multicols}{2}

\raggedcolumns	

\centerline{\textit{Advent}}%From the Ritual of the Altar
\begin{rubric}
	Every day in \blackrubric{Advent,} including \blackrubric{Christmas Eve.}
\end{rubric}
\lett{T}{hrough} Jesus Christ our Lord; whom thou, the merciful and faithful God, didst promise as a Saviour to lost mankind; by whose truth the unlearned should be instructed, by whose holiness the ungodly should be justified, by whose power the weak should be holpen. Now, therefore, when the time in which thou shouldest send him draweth near, and now that the day of our deliverance already shineth; trusting in these thy promises, with holy joy we do rejoice.\\

\centerline{\textit{Christmas}}
%\subsubsection{Christmas}%1928 BCP:
\begin{rubric}
	Upon \blackrubric{Christmas Day} and its Octave, until \blackrubric{Epiphany Eve}, inclusive, and on the Feast of the \blackrubric{Most Holy Body of Christ} and its Octave.
\end{rubric}
\lett{B}{ecause} thou didst give Jesus Christ, thine only Son, to be born (as at this time) for us; who, by the operation of the Holy Ghost, was made very man, of the substance of the Virgin Mary his mother; and that without spot of sin, to make us clean from all sin.%\\

%MANUAL ADJUSTMENT:
\newcolumn

\centerline{\textit{Epiphany}}
%\subsubsection{Epiphany}%1928 BCP:
\begin{rubric}
	Upon the \blackrubric{Epiphany} and its Octave.
\end{rubric}
\lett{T}{hrough} Jesus Christ our Lord, who, in substance of our mortal flesh, manifested forth his glory; that he might bring us out of darkness into his own glorious light.\\

\centerline{\textit{Septuagesima Sunday}}
%\subsubsection{Septuagesima Sunday}
\begin{rubric}
	Optionally upon \blackrubric{Septuagesima Sunday.}
\end{rubric}
\lett{F}{or} by things seen, we are instructed how we should pursue those things unseen. And so are reminded, as the year leadeth on its course, to pass, from things past unto things to come, from the old man into the newness of life; that, being freed from reliance on earthly goods, we may lay hold upon the excelling richness of the heavenly gift; and by that food, which is preserved for us in every age, we may arrive unto that victorious sustenance which remaineth for ever, even Jesus Christ our Lord.\\

%Quia per ea quæ conspiciuntur instruimur quibus modis ad invisibilia tendere debeamus. Denique commonemur, anni ducente successu, de præteritis ad futura, de vetustate in novitatem vitæ transire, Ut terrenis sustentationibus expediti, cælestis doni capiamus desiderabilius ubertatem, Et per eum cibum qui beneficiis prærogatur alternis, perveniamus ad victum sine fine mansurum, Jesum Christum Dominum nostrum.

\centerline{\textit{Sexagesima Sunday}}
%\subsubsection{Sexagesima Sunday}
\begin{rubric}
	Optionally upon \blackrubric{Sexagesima Sunday.}
\end{rubric}
\lett{W}{ho} deigned to instruct the rational creature by such stewardship, lest he be given over to temporal things and strive not for the everlasting reward. That he may neither withdraw from correction nor become haughty in approbation, but rather overcome all adversities, this being his glorious devotion: Jesus Christ our Lord.\\

%Qui rationabilem creaturam, ne temporalibus dedita bonis, ad praemia sempiterna non tendat, ea dispensatione dignaris erudire, ut nec castigatione deficiat, nec prosperitatibus insolescat; sed hoc potius fiat ejus gloriosa devotio, quo nullis adversitatibus obruta superetur. Per Christum Dominum nostrum.

\centerline{\textit{Quinquagesima Sunday}}
%\subsubsection{Quinquagesima Sunday}
\begin{rubric}
	Optionally on \blackrubric{Quinquagesima Sunday.}
\end{rubric}
\lett{A}{nd} to entreat thy Majesty by lowly devotion, that thou, beholding the weakness of our mortal nature, accuse us not of our depravity in thy wrath, but rather cleanse, cultivate, and console us by thine immeasurable gentleness. And since we can do no good thing without thee, bestow upon us thy grace alone that we may live in wholesome conversation.\\

%Et majestatem tuam cernua devotione exorare, ut modulum terrenae fragilitatis aspiciens, non in ira tua pro nostra pravitate nos arguas, sed immensa clementia purifices, erudias, consoleris. Qui quum sine te nihil possumus facere, quod tibi sit placitum, tua nobis gratia sola praestabit, ut salubri conversatione vivamus.

\vspace{-1ex}

\centerline{\textit{Lent}}
%\subsubsection{Lent}%1933 English Missal:
\begin{rubric}
	From \blackrubric{Ash Wednesday} until the Saturday before \blackrubric{Passion Sunday,} inclusive.
\end{rubric}
\lett{W}{ho} by bodily fasting dost overcome vice, dost raise the mind, and dost bestow on us virtue and heavenly rewards: through Christ, our Lord.\\

\vspace{-1ex}

\centerline{\textit{Holy Rood}}
%\subsubsection{Holy Rood}%1933 English Missal:
\begin{rubric}
	From \blackrubric{Passion Sunday} until \blackrubric{Maundy Thursday,} inclusive. And all Masses of the \blackrubric{Holy Rood,} the \blackrubric{Passion of the Lord,} or the \blackrubric{Most Precious Blood,} and within their Octaves, even when celebrated within the Octave of \blackrubric{Christmas.}
\end{rubric}
\lett{W}{ho} by the tree of the Cross didst give salvation unto mankind: that whence death arose, thence life might rise again: and that he, who by a tree overcame, might also by a tree be overcome: through Christ our Lord.

%MANUAL ADJUSTMENT:
\vspace{0.5\baselineskip}

\centerline{\textit{Easter}}
%\subsubsection{Easter} %1928 BCP:
\begin{rubric}
	From \blackrubric{Holy Saturday} until \blackrubric{Ascension Eve,} inclusive.
\end{rubric}
\lett{B}{ut} chiefly are we bound to praise thee for the glorious Resurrection of thy Son Jesus Christ our Lord: for he is the very Paschal Lamb, which was offered for us, and hath taken away the sin of the world; who by his death hath destroyed death, and by his rising to life again hath restored to us everlasting life.

%MANUAL ADJUSTMENT:
\vspace{0.5\baselineskip}

\centerline{\textit{Ascension}}
%\subsubsection{Ascension}%1928 BCP:
\begin{rubric}
	From \blackrubric{Ascension Thursday} until the Friday after its Octave, inclusive. It is also said, even without a Commemoration, on the Friday after the Octave.
\end{rubric}
\lett{T}{hrough} thy most dearly beloved Son Jesus Christ our Lord; who, after his most glorious Resurrection, manifestly appeared to all his Apostles, and in their sight ascended up into heaven, to prepare a place for us; that where he is, thither we might also ascend, and reign with him in glory.\\

\centerline{\textit{Holy Ghost}}
%\subsubsection{Holy Ghost}%1928 BCP Whitsun Preface:
\begin{rubric}
	From \blackrubric{Whitsunday Eve} until \blackrubric{Trinity Sunday,} exclusive, in Votive Masses of the Holy Ghost, even when celebrated within the Octave of \blackrubric{Christmas.}
\end{rubric}
\lett{T}{hrough} Jesus Christ our Lord; according to whose most true promise, the Holy Ghost came down (as at this time) from heaven, lighting upon the Apostles, to teach them, and to lead them into all truth; giving them boldness with fervent zeal constantly to preach the Gospel unto all nations; whereby we have been brought out of darkness and error into the clear light and true knowledge of thee, and of thy Son Jesus Christ.\\

\centerline{\textit{Most Holy Trinity}}
%\subsubsection{Most Holy Trinity}%1928 BCP:
\begin{rubric}
	Upon Sundays, from the \nth{3} Sunday after \blackrubric{Trinity} until the \blackrubric{Sunday Next before Advent,} inclusive. And also in Masses of the \blackrubric{Most Holy Trinity,} even when celebrated within the Octave of \blackrubric{Christmas.} And on Sundays in \blackrubric{Septuagesimatide.}
\end{rubric}
\lett{W}{ho,} with thine only-begotten Son, and the Holy Ghost, art one God, one Lord, in Trinity of Persons and in Unity of Substance. For that which we believe of thy glory, O Father, the same we believe of the Son, and of the Holy Ghost, without any difference of inequality.\\

\centerline{\textit{Compassion of Our Lord}}
%\subsubsection{Compassion of Our Lord}%1933 English Missal:
\begin{rubric}
	Upon the Feast of the of the \blackrubric{Compassion of Our Lord} and its Octave Day.
\end{rubric}
\lett{W}{ho} didst will that thine only-begotten Son, as he hung upon the Cross, should be pierced by the soldier's spear: that his Heart thus opened, the shrine of the divine bounty, might pour forth upon us streams of mercy and of grace: that as it never ceased to burn with love for us, so it might be a rest for the godly, and an open refuge of salvation for the penitent.\\

\centerline{\textit{Christ the King}}
%\subsubsection{Christ the King}%1933 English Missal:
\begin{rubric}
	Masses of \blackrubric{Our Lord Jesus Christ the King,} even when they are celebrated within the Octave of \blackrubric{Christmas.}
\end{rubric}
\lett{W}{ho} didst anoint thine only-begotten Son, our Lord Jesus Christ, with the oil of gladness, to be a Priest for ever and the King of all the world: that, offering himself an unspotted sacrifice of peace upon the altar of the cross, he might accomplish the sacrament of the redemption of mankind: and making all creatures subject to his governance, might deliver up to thine infinite Majesty an eternal and universal kingdom. A kingdom of truth and life: a kingdom of sanctity and grace: a kingdom of justice, love, and peace.\\

\centerline{\textit{All Hallows'}}
%\subsubsection{All Hallows'}%1928 BCP:
\begin{rubric}
	Upon the Feast of \blackrubric{All Hallows} and its Octave.
\end{rubric}
\lett{W}{ho,} in the multitude of thy saints, hast compassed us about with so great a cloud of witnesses that we, rejoicing in their fellowship, may run with patience the race that is set before us, and together with them may receive the crown of glory that fadeth not away.\\

%MANUAL ADJUSTMENT:
\newcolumn

\centerline{\textit{Of Our Lord}}
%\subsubsection{Of Our Lord}
\begin{rubric}
	Upon the Feasts of the \blackrubric{Purification,} \blackrubric{Annunciation,} and \blackrubric{Transfiguration.}
\end{rubric}
\lett{B}{ecause} in the Mystery of the Word made flesh, thou hast caused a new light to shine in our hearts, to give the knowledge of thy glory in the face of thy Son, Jesus Christ our Lord.\\

\centerline{\textit{Blessed Virgin Mary}}
%\subsubsection{Blessed Virgin Mary}
\begin{rubric}
	In Masses of the \blackrubric{Blessed Virgin Mary,} and their Octaves.
\end{rubric}
\lett{A}{nd} that in the \textit{N.} of Blessed Mary ever Virgin we should praise, bless, and magnify thee. In that by the overshadowing of the Holy Ghost she conceived thine only-begotten Son: and, the glory of her maidenhood yet abiding, shed forth upon the world the light eternal, Jesus Christ, our Lord.\\

\centerline{\textit{St. Joseph, Spouse of the Blessed Virgin}}
%\subsubsection{St. Joseph, Spouse of the Blessed Virgin}
\begin{rubric}
	In Masses of \blackrubric{St. Joseph,} and their Octaves.
\end{rubric}
\lett{A}{nd} that in the Festivity (Veneration) of blessed Joseph, we should magnify, bless, and glorify thee with worthy praises. Who being a just man was given by thee for a Spouse to the Virgin Mother of God: and, being a faithful and wise servant, was made ruler over thy Household: that in the office of a father he should guard thine only-begotten Son, conceived by the overshadowing of the Holy Ghost, even Jesus Christ our Lord.\\

\centerline{\textit{Apostles \& Evangelists}}
%\subsubsection{Apostles \& Evangelists}
\begin{rubric}
	In Masses of the \blackrubric{Apostles} or \blackrubric{Evangelists} and in a Mass of the Election and Enthronement of the Patriarch and their anniversaries.
\end{rubric}
\lett{T}{hat} we, O Lord, should humbly entreat thee, that thou, the everlasting Shepherd, wouldst not forsake thy flock: but through thy blessed Apostles keep it by thy continual protection. That it may be governed by those same rulers, whom in thy stead thou hast appointed for thy work, as shepherds of thy people.\\

\centerline{\textit{Requiem}}
%\subsubsection{Requiem}
\begin{rubric}
	In all \blackrubric{Masses for the Dead.}
\end{rubric}
\lett{T}{hrough} Christ, our Lord. In whom hath shone forth unto us the hope of a blessed resurrection, that they who bewail the certain condition of their mortality may be consoled by the promise of immortality to come. For to thy faithful people, O Lord, life is changed not taken away: and at the dissolution of the tabernacle of this earthly sojourning a dwelling-place eternal is made ready in the heavens.

\end{multicols}

\clearpage

\subsection{Proper Prefaces with Solemn Chant}
\fancyhead[RO,LE]{\textit{Solemn Prefaces}}

\subsubsection{Introduction}
\fancyhead[RE,LO]{Introduction}

\gregorioscore{resources/gabc/Mass/Prefaces/FestalIntroduction.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Advent}%From the Ritual of the Altar
\fancyhead[RE,LO]{Advent}

\gregorioscore{resources/gabc/Mass/Prefaces/Advent.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Christmas}%1928 BCP:
\fancyhead[RE,LO]{Christmas}

\gregorioscore{resources/gabc/Mass/Prefaces/Christmas.gabc}


\subsubsection{Epiphany}%1928 BCP:
\fancyhead[RE,LO]{Epiphany}

\gregorioscore{resources/gabc/Mass/Prefaces/Epiphany.gabc}

%\subsubsection{Septuagesima Sunday}
%\fancyhead[RE,LO]{Septuagesima}


%\subsubsection{Sexagesima Sunday}
%\fancyhead[RE,LO]{Sexagesima}


%\subsubsection{Quinquagesima Sunday}
%\fancyhead[RE,LO]{Quinquagesima}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Lent}%1933 English Missal:
\fancyhead[RE,LO]{Lent}

\gregorioscore{resources/gabc/Mass/Prefaces/Lent.gabc}


\subsubsection{Holy Rood}%1933 English Missal:
\fancyhead[RE,LO]{Holy Rood}

\gregorioscore{resources/gabc/Mass/Prefaces/HolyRood.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Easter} %1928 BCP:
\fancyhead[RE,LO]{Easter}

\gregorioscore{resources/gabc/Mass/Prefaces/Easter.gabc}

\subsubsection{Ascension}%1928 BCP:
\fancyhead[RE,LO]{Ascension}

\gregorioscore{resources/gabc/Mass/Prefaces/Ascension.gabc}

\subsubsection{Holy Ghost}%1928 BCP Whitsun Preface:
\fancyhead[RE,LO]{Holy Ghost}

\gregorioscore{resources/gabc/Mass/Prefaces/HolyGhost.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Most Holy Trinity}%1928 BCP:
\fancyhead[RE,LO]{Most Holy Trinity}

\gregorioscore{resources/gabc/Mass/Prefaces/Trinity.gabc}


\subsubsection{Compassion of Our Lord}%1933 English Missal:
\fancyhead[RE,LO]{Divine Compassion}

\gregorioscore{resources/gabc/Mass/Prefaces/CompassionFestal.gabc}

\subsubsection{Christ the King}%1933 English Missal:
\fancyhead[RE,LO]{Christ the King}

\gregorioscore{resources/gabc/Mass/Prefaces/CTK.gabc}

\subsubsection{All Hallows'}%1928 BCP:
\fancyhead[RE,LO]{All Hallows'}

\gregorioscore{resources/gabc/Mass/Prefaces/AllHallows.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Of Our Lord}
\fancyhead[RE,LO]{Of Our Lord}

\gregorioscore{resources/gabc/Mass/Prefaces/OurLord.gabc}

\subsubsection{Blessed Virgin Mary}
\fancyhead[RE,LO]{Blessed Virgin Mary}

\gregorioscore{resources/gabc/Mass/Prefaces/BVM.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{St. Joseph, Spouse of the Blessed Virgin}
\fancyhead[RE,LO]{St. Joseph}

\gregorioscore{resources/gabc/Mass/Prefaces/Joseph.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Apostles \& Evangelists}
\fancyhead[RE,LO]{Apostles}

\gregorioscore{resources/gabc/Mass/Prefaces/Apostles.gabc}


\subsubsection{Conclusion}
\fancyhead[RE,LO]{Conclusion}

\gregorioscore{resources/gabc/Mass/Prefaces/FestalClosing.gabc}

%\subsubsection{Common}
%\fancyhead[RE,LO]{Common}

%\gregorioscore{resources/gabc/Mass/Prefaces/Common.gabc}

\clearpage

\subsection{Proper Prefaces with Ferial Chant}
\fancyhead[RO,LE]{\textit{Ferial Prefaces}}

\subsubsection{Introduction}
\fancyhead[RE,LO]{Introduction}

\gregorioscore{resources/gabc/Mass/Prefaces/FerialIntroduction.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Lent}
\fancyhead[RE,LO]{Lent}

\gregorioscore{resources/gabc/Mass/Prefaces/LentFerial.gabc}

\subsubsection{Holy Rood}%1933 English Missal:
\fancyhead[RE,LO]{Holy Rood}

\gregorioscore{resources/gabc/Mass/Prefaces/HolyRoodFerial.gabc}

%\subsubsection{Compassion of Our Lord}%1933 English Missal:
%\fancyhead[RE,LO]{Divine Compassion}

%\gregorioscore{resources/gabc/Mass/Prefaces/CompassionFerial.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Requiem}
\fancyhead[RE,LO]{Requiem}

\gregorioscore{resources/gabc/Mass/Prefaces/Requiem.gabc}

%\subsubsection{Common}
%\fancyhead[RE,LO]{Common}

%\gregorioscore{resources/gabc/Mass/Prefaces/CommonFerial.gabc}

%MANUAL ADJUSTMENT:
\clearpage
\subsubsection{Conclusion}
\fancyhead[RE,LO]{Conclusion}

\gregorioscore{resources/gabc/Mass/Prefaces/FerialClosing.gabc}